### 2025-02-07 Download_data.qmd
R code to download data

---
title: "Weather data"
format: html
editor: visual
echo: false
---

## Set up

```{r setup, include=FALSE}

# libraries
library(tigris)
library(dplyr)
library(exactextractr)
library(data.table)
library(raster)
library(stringr)
library(lubridate)
library(terra)
library(dplyr)
library(purrr)
library(sf)
library(future.apply)
library(soilDB)

# no warnings in the output
options(warn=-1)
```

# Weather data

## Function

```{r}
process_gridMET_data <- function(states, var_name, years) {

  if (!dir.exists("Data")) dir.create("Data")

  counties <- purrr::map_dfr(states, ~ tigris::counties(state = .x)) %>%
    dplyr::select(STATEFP, COUNTYFP) %>%
    mutate(rowid = row_number())

  all_data <- list()

  for (year in years) {

    target_url <- paste0("http://www.northwestknowledge.net/metdata/data/", var_name, "_", year, ".nc")
    file_name <- paste0("Data/", var_name, "_", year, ".nc")

    tryCatch({
      downloader::download(url = target_url, destfile = file_name, mode = "wb", timeout = 600)

      temp_rast <- rast(file_name)

      temp_data <- exact_extract(temp_rast, counties, progress = FALSE) %>%
        rbindlist(idcol = "rowid") %>%
        melt(id.var = c("rowid", "coverage_fraction")) %>%
        .[!is.na(value), ] %>%
        .[, variable := str_sub(variable, -5, -1) %>% as.numeric()] %>%
        .[, date := variable + ymd("1900-01-01")] %>%
        .[, .(value = sum(value * coverage_fraction) / sum(coverage_fraction)), by = .(rowid, date)] %>%
        .[, var := var_name]

      all_data[[as.character(year)]] <- temp_data

    }, error = function(e) {
      message(paste("Failed to download or process data for", var_name, "in year", year, ":", e))
    })
  }

  final_data <- rbindlist(all_data, idcol = "year")

  final_data <- counties %>%
    left_join(final_data, by = "rowid")

  return(final_data)
}

```

```{r}

states <- c("IA", "IL", "IN", "MN", "MO", "OH", "KS")

# states <- c("IA", "IL")   
var_names <- c("pr", "tmmx", "tmmn") # precipitation, max temperature, min temperature
#years <- 2019:2020
years <- 1980:2023

results <- lapply(var_names, function(var_name) {
  process_gridMET_data(states = states, var_name = var_name, years = years)
})

final_results <- rbindlist(results, idcol = "variable")

saveRDS(final_results, "all_states.rds")
```

# Precipitation

```{r}

pr_data <- final_results[final_results$var == "pr", ]
head(pr_data)

library(data.table)Precipitation
library(lubridate)

setDT(pr_data)

pr_data[, `:=`(year = year(date), month = month(date))]

# sum rain
monthly_pr_data <- pr_data[, .(monthly_pr = sum(value, na.rm = TRUE)), by = .(year, month, STATEFP, COUNTYFP, rowid)]

head(monthly_pr_data)

```

# Monthly weather data

```{r}

# Monthly data

setDT(final_results)

final_results[, `:=`(year = year(date), month = month(date))]

monthly_pr <- final_results[var == "pr", .(monthly_pr = sum(value, na.rm = TRUE)), 
                            by = .(year, month, STATEFP, COUNTYFP, rowid)]

monthly_temp <- final_results[var %in% c("tmmx", "tmmn"), 
                              .(monthly_temp = mean(value, na.rm = TRUE)), 
                              by = .(year, month, STATEFP, COUNTYFP, rowid, var)]

monthly_temp_wide <- dcast(monthly_temp, year + month + STATEFP + COUNTYFP + rowid ~ var, value.var = "monthly_temp")

monthly_data <- merge(monthly_pr, monthly_temp_wide, by = c("year", "month", "STATEFP", "COUNTYFP", "rowid"), all = TRUE)

```

# Yearly weather data

```{r}

setDT(final_results)

final_results[, year := year(date)]

yearly_pr <- final_results[var == "pr", .(yearly_pr = sum(value, na.rm = TRUE)), 
                           by = .(year, STATEFP, COUNTYFP, rowid)]

yearly_temp <- final_results[var %in% c("tmmx", "tmmn"), 
                             .(yearly_temp = mean(value, na.rm = TRUE)), 
                             by = .(year, STATEFP, COUNTYFP, rowid, var)]

yearly_temp_wide <- dcast(yearly_temp, year + STATEFP + COUNTYFP + rowid ~ var, value.var = "yearly_temp")

yearly_data <- merge(yearly_pr, yearly_temp_wide, by = c("year", "STATEFP", "COUNTYFP", "rowid"), all = TRUE)

head(yearly_data)
```

# Yield data

```{r}

# Define crops, states, and years
crops <- c("CORN", "WHEAT", "SOYBEANS")
states <- c("IOWA", "ILLINOIS", "INDIANA", "MINNESOTA", "MISSOURI", "OHIO", "KANSAS")
years <- paste(2000:2023)

# Function to get data for a specific crop, state, and year
get_crop_data <- function(crop, state, year) {
  tryCatch({
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = crop,
      geographic_level = "COUNTY",
      state = state,
      year = as.character(year),
      geometry = TRUE
    ) %>%
      # Keep only selected variables
      dplyr::select(
        year, county_name, county_code, state_name,
        state_fips_code, short_desc, Value
      )
  }, error = function(e) {
    # Return NULL for combinations with no data 
    NULL
  })
}

# Create all combinations of crops, states, and years
combinations <- expand.grid(crop = crops, state = states, year = years)

crop_state_year_data <- combinations %>%
  pmap_dfr(~ get_crop_data(..1, ..2, ..3))

head(crop_state_year_data)

```

Download yield data

```{r}

(
  IL_CO_NE_corn <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "CORN",
      geographic_level = "COUNTY",
      state = c("IOWA", "ILLINOIS", "INDIANA", "MINNESOTA", "MISSOURI", "OHIO", "KANSAS"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

(
  IL_CO_NE_corn <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "CORN",
      geographic_level = "COUNTY",
      state = c("ILLINOIS", "IOWA", "MISSOURI", "KANSAS"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

##$$

# works: "ILLINOIS", "IOWA", "MISSOURI", "KANSAS"

(
  IL_CO_NE_corn_two_more <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "CORN",
      geographic_level = "COUNTY",
      state = c("INDIANA", "OHIO"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

corn_1 <-  IL_CO_NE_corn
corn_2 <-  IL_CO_NE_corn_two_more

## Wheat

(
  WHEAT_1 <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "WHEAT",
      geographic_level = "COUNTY",
      state = c("ILLINOIS"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

# c("IOWA", "ILLINOIS", "INDIANA", "MINNESOTA", "MISSOURI", "OHIO", "KANSAS")

(
  WHEAT_iowa <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "WHEAT",
      geographic_level = "COUNTY",
      state = c("IOWA"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

(
  WHEAT_rest <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "WHEAT",
      geographic_level = "COUNTY",
      state = c("INDIANA", "MINNESOTA", "MISSOURI", "OHIO"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

(
  WHEAT_kan <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "WHEAT",
      geographic_level = "COUNTY",
      state = c("KANSAS"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

wheat_1 <- WHEAT_1
wheat_2 <-  WHEAT_iowa
wheat_3 <- WHEAT_rest
wheat_4 <-  WHEAT_kan
# SOYBEANS

(
  SOYBEANS <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "SOYBEANS",
      geographic_level = "COUNTY",
      state = c("ILLINOIS", "IOWA", "MISSOURI", "KANSAS"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

(
  SOYBEANS_2 <-
    tidyUSDA::getQuickstat(
      key = "169CA3FD-76B8-38E5-8DD7-F2CB7BCEC122",
      program = "SURVEY",
      commodity = "SOYBEANS",
      geographic_level = "COUNTY",
      state = c("INDIANA", "OHIO"),
      year = paste(2000:2023),
      geometry = TRUE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(
      year, county_name, county_code, state_name,
      state_fips_code, short_desc, Value
    )
)

soybeans_1 <-  SOYBEANS
soybeans_2 <-  SOYBEANS_2

##

combined_all <- bind_rows(
  corn_1,
  corn_2,
  wheat_1,
  wheat_2,
  wheat_3,
  wheat_4,
  soybeans_1,
  soybeans_2
)

saveRDS(combined_all , "combined_all_yields_all_non_filtered.rds")
library(dplyr)

all_non_irr <- combined_all %>%
  filter(short_desc %in% c(
    "CORN, GRAIN - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "WHEAT - YIELD, MEASURED IN BU / ACRE" ,
    "WHEAT, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS - YIELD, MEASURED IN BU / ACRE" ,  
    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE"
  ))

saveRDS(all_non_irr, "all_non_irr_yields_for_all_states.rds")

library(dplyr)

selected_short_desc <- c(
  "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
  "WHEAT, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
  "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE"
)

test_it <- all_non_irr %>%
  filter(
    !(state_name == "KANSAS" & !short_desc %in% selected_short_desc)
  )

saveRDS(test_it, "all_final_yields_for_all_states.rds")

###

com <- as.data.frame(test_it) %>%
  full_join(
    as.data.frame(Weather_after_2000),
    by = c("year" = "year", "county_code" = "COUNTYFP", "state_fips_code" = "STATEFP")
  ) %>%
  filter(complete.cases(across(-geometry)))

saveRDS(com, "com.rds")

com <- com[order(com$year), ]

# Load the dplyr package
library(dplyr)

# Create the new column 'crop' by extracting the crop name from 'short_desc'
com_2 <- com %>%
  mutate(crop = case_when(
    grepl("CORN", short_desc) ~ "CORN",
    grepl("WHEAT", short_desc) ~ "WHEAT",
    grepl("SOYBEANS", short_desc) ~ "SOYBEANS",
    TRUE ~ NA_character_  # For any unmatched rows
  ))

com_2$short_desc <- NULL

com_2 <- dplyr::select(com_2, year, county_name, county_code, state_name, state_fips_code, 
                       Value, crop, everything())

saveRDS(com_2, "com_2.rds")

```

# **SSURGO**

```{r}

states <- c("IA", "IL", "IN", "MN", "MO", "OH", "KS")

all_data_states <- st_as_sf(all_states)
class(all_data_states)

counties <- com_2
counties <- st_as_sf(counties)

# Transform CRS to EPSG:4326
counties <- st_transform(counties, crs = 4326)
st_is_valid(counties)

```

# Get SC code for counties

```{r}

#  function to fetch sc_code for a single county
fetch_sc_code <- function(county_row) {
  tryCatch({

    result <- SDA_spatialQuery(county_row, what = "areasymbol", db = "SSURGO")
    if (!is.null(result)) {
      return(result$areasymbol[1])
    } else {
      return(NA)
    }
  }, error = function(e) {

    return(NA)
  })
}

# Apply function to each row
counties$sc_code <- purrr::map_chr(1:nrow(counties), function(i) {
  print(paste0("Fetching sc_code for county ", i, " of ", nrow(counties)))
  fetch_sc_code(counties[i, ])
})

# change to EPSG:4326 
counties <- st_transform(counties, crs = 4326)

saveRDS( counties, "com_3.rds")
```

# Download soil characteristics for counties

```{r}

dir.create("SSURGO", recursive = TRUE)

plan("multisession", workers = parallel::detectCores() - 1)

# Download and save SSURGO geometry (mukey) data for counties
files_ls <- lapply(1:nrow(counties), \(x) {
  file.exists(paste0("SSURGO/ssurgo_sf_", x, "_", counties$county_code[x], ".rds"))
}) %>% unlist()

future_lapply(
  which(!files_ls),
  \(x) {
    print(paste0(x, "/", nrow(counties)))

    ssurgo_geom_file_name <- paste0("SSURGO/ssurgo_sf_", x, "_", counties$county_code[x], ".rds")
    if (!file.exists(ssurgo_geom_file_name)) {
      print(paste0("(geom) working on ", x, "/", nrow(counties)))
      temp <- get_ssurgo_geom(county = counties[x, ])
      saveRDS(temp, ssurgo_geom_file_name)
    }
  }
)
###########################

# setwd("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/lt")

files_ls <-
  lapply(
    1:nrow(counties),
    \(x) {
      file.exists(paste0("SSURGO/ssurgo_sf_", x, "_", counties[x, ]$sc_code, ".rds"))
    }
  ) %>%
  unlist()

future_lapply(
  which(!files_ls),
  \(x) {
    print(paste0(x, "/", nrow(counties)))

    ssurgo_geom_file_name <- paste0("SSURGO/ssurgo_sf_", x, "_", counties[x, ]$sc_code, ".rds")
    if (!file.exists(ssurgo_geom_file_name)) {
      print(paste0("(geom) working on ", x, "/", nrow(counties)))
      temp <- get_ssurgo_geom(county = counties[x, ])
      saveRDS(temp, ssurgo_geom_file_name)
    }
  }
)

future_lapply(
  1:nrow(counties),
  \(x) {
    print(paste0(x, "/", nrow(counties)))

    ssurgo_geom_file_name <- paste0("SSURGO/ssurgo_sf_", x, "_", counties[x, ]$sc_code, ".rds")

    ssurgo_data_file_name <- paste0("SSURGO/ssurgo_data_", x, "_", counties[x, ]$sc_code, ".rds")

    if (!file.exists(ssurgo_data_file_name) & file.exists(ssurgo_geom_file_name)) {
      ssurgo_geom <- readRDS(ssurgo_geom_file_name)
      print(paste0("(data) Working on ", x, "/", nrow(counties)))
      temp <- get_ssurgo_raw_data(ssurgo_geom, vars = "awc_r")
      saveRDS(temp, ssurgo_data_file_name)
    }
  }
)

future_lapply(
  1:nrow(counties),
  \(x) {
    print(paste0(x, "/", nrow(counties)))

    ssurgo_data_file_name <- paste0("SSURGO/ssurgo_data_", x, "_", counties[x, ]$sc_code, ".rds")

    whc_file_name <- paste0("SSURGO/whc_", x, "_", counties[x, ]$sc_code, ".rds")

    if (!file.exists(whc_file_name) & file.exists(ssurgo_data_file_name)) {
      ssurgo_data <- readRDS(ssurgo_data_file_name)
      print(paste0("(whc) Working on ", x, "/", nrow(counties)))
      temp <- get_wch(ssurgo_data)
      saveRDS(temp, whc_file_name)
    }
  }
)

# ch
ssurgo_soil <-
  fs::dir_ls("SSURGO") %>%
  .[str_detect(., "whc")] %>%
  purrr::map(readRDS) %>%
  rbindlist()

saveRDS(ssurgo_soil, "ssurgo_soil_all.rds")

```

## Combine all data (weather, ssurgo, yield)

```{r}

Combined_yield_Weather_whc <- com_3 %>%
  left_join(ssurgo_soil_all, by = "sc_code")

saveRDS(Combined_yield_Weather_whc, "Combined_yield_Weather_whc_all_states.rds")

# Convert tmmn and tmmx from Kelvin to Celsius
Combined_yield_Weather_whc_all_states <- Combined_yield_Weather_whc_all_states %>%
  mutate(
    tmmn = tmmn - 273.15,
    tmmx = tmmx - 273.15
  )

saveRDS(Combined_yield_Weather_whc_all_states, "Combined_yield_Weather_whc_all_states_degree.rds")

```

### 2025-02-07 Rlearner_python_in_R.qmd
Python in R torch

---
title: "Rlearner_in_R"
format: html
editor: visual
echo: false
---

## Set up

```{r setup, include=FALSE}

# libraries
library(knitr)
library(readr)
library(torch)
library(reshape2)
library(dplyr)
library(GGally)
library(ggplot2)
library(purrr)

# directory 
knitr::opts_knit$set(root.dir = "/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Takashi_code/Write_python_in_R/Data")

# no warnings in the output
options(warn=-1)
```

# Rann_regression_20230413 copy_d2

## Import data

```{r}
# Import data

# setwd("/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Takashi_code/Write_python_in_R/Data")
data_2nd_stage <- read_csv("./DG2_ss.csv")
evall_N_seq <- read_csv("./DG2_eval_data.csv")

```

```{r}

plot_loss <- function(history) {
  loss <- as_array(history$loss)
  val_loss <- as_array(history$val_loss)

  # data frame
  loss_data <- data.frame(
    epoch = seq_along(loss),
    loss = loss,
    val_loss = val_loss
  )

  loss_data_long <- melt(loss_data, id.vars = "epoch")

  ggplot(loss_data_long, aes(x = epoch, y = value, color = variable)) +
    geom_line(size = 1) +
    labs(x = "Epoch", y = "MAE", color = "Loss Type") +
    theme_minimal() +
    theme(legend.position = "top") +
    geom_point()
}

```

# Multiple input DNN

```{r}

dataset <- data_2nd_stage %>%
  select(y_tilde, x1, x2, T_1_tilde, T_2_tilde, T_3_tilde)

# Split data into train (80%) and test (20%)

train_indices <- sample(1:nrow(dataset), size = 0.8 * nrow(dataset))  
train_dataset <- dataset[train_indices, ] 
test_dataset <- dataset[-train_indices, ]

# Convert data to torch tensors
train_tensor <- torch_tensor(as.matrix(train_dataset))
test_tensor <- torch_tensor(as.matrix(test_dataset))

```

```{r}

# Pair plot 
ggpairs(train_dataset, diag = list(continuous = "density")) +
  theme_minimal()

```

```{r}

# features
train_features_tensor <- train_tensor[, 2:6, drop = FALSE]
test_features_tensor <- test_tensor[, 2:6, drop = FALSE]

# labels
train_labels_tensor <- train_tensor[, 1, drop = FALSE]   # First column (y_tilde) 
test_labels_tensor <- test_tensor[, 1, drop = FALSE]   # First column (y_tilde)

# compute mean and Std (from train data) for standardization
train_means <- torch_mean(train_features_tensor, dim = 1)
train_sds <- torch_std(train_features_tensor, dim = 1, unbiased = TRUE)

# standardize features (use training stats for both train and test data)
scaled_train_features_tensor <- (train_features_tensor - train_means) / train_sds
scaled_test_features_tensor <- (test_features_tensor - train_means) / train_sds

```

```{r}

#  R torch neural network model
RModel <- nn_module(
  initialize = function() {

    self$dense1_x <- nn_linear(2, 64)
    self$dense2_x <- nn_linear(64, 64)
    self$dense3_x <- nn_linear(64, 1)

    self$dense1_T <- nn_linear(2, 64)
    self$dense2_T <- nn_linear(64, 64)
    self$dense3_T <- nn_linear(64, 1)

    self$dense1_T3 <- nn_linear(2, 64)
    self$dense2_T3 <- nn_linear(64, 64)
    self$dense3_T3 <- nn_linear(64, 1)
  },

  forward = function(input) {

    x <- input[, 1:2]   # First two columns
    T <- input[, 3:5]   # Last three columns

    T1 <- T[, 1, drop=FALSE]   # First column
    T2 <- T[, 2, drop=FALSE]   # Second column
    T3 <- T[, 3, drop=FALSE]   # Third column

    # Pass x to dense layers
    dense1 <- x %>%
      self$dense1_x() %>%
      nnf_relu() %>%
      self$dense2_x() %>%
      nnf_relu() %>%
      self$dense3_x()

    dense1 <- dense1 * T1

    dense2 <- x %>%
      self$dense1_T() %>%
      nnf_relu() %>%
      self$dense2_T() %>%
      nnf_relu() %>%
      self$dense3_T()

    dense2 <- dense2 * T2

    dense3 <- x %>%
      self$dense1_T3() %>%
      nnf_relu() %>%
      self$dense2_T3() %>%
      nnf_relu() %>%
      self$dense3_T3()

    dense3 <- dense3 * T3

    # Sum the outputs
    output <- dense1 + dense2 + dense3
    return(output)
  }
)

# model 
model <- RModel()

# optimizer and loss function
optimizer <- optim_adam(model$parameters, lr = 0.001)
loss_fn <- nnf_l1_loss  # MAE (Mean Absolute Error)

print(model)

```

```{r}

# Hyperparameters
epochs <- 500
batch_size <- 128
patience <- 10  # early stopping 
min_delta <- 0.0  # minimum improvement

# Optimizer & loss function
optimizer <- optim_adam(model$parameters, lr = 0.001)
loss_fn <- nnf_l1_loss  # MAE

train_dataset <- tensor_dataset(scaled_train_features_tensor, train_labels_tensor)
test_dataset <- tensor_dataset(scaled_test_features_tensor, test_labels_tensor)

train_dataloader <- dataloader(train_dataset, batch_size = batch_size, shuffle = TRUE)
test_dataloader <- dataloader(test_dataset, batch_size = batch_size, shuffle = FALSE)

# early stopping parameters
best_val_loss <- Inf
epochs_without_improvement <- 0

# history list
history <- list(train_loss = c(), val_loss = c())

# training Loop
for (epoch in 1:epochs) {
  model$train()
  train_loss <- 0

  coro::loop(for (batch in train_dataloader) {
    optimizer$zero_grad()

    inputs <- batch[[1]]
    targets <- batch[[2]]

    outputs <- model(inputs)
    loss <- loss_fn(outputs, targets)

    loss$backward()
    optimizer$step()

    train_loss <- train_loss + loss$item()
  })

  # average training loss
  train_loss <- train_loss / length(train_dataloader)

  # validation
  model$eval()
  val_loss <- 0

  coro::loop(for (batch in test_dataloader) {
    inputs <- batch[[1]]
    targets <- batch[[2]]

    outputs <- model(inputs)
    loss <- loss_fn(outputs, targets)

    val_loss <- val_loss + loss$item()
  })

  # average validation loss
  val_loss <- val_loss / length(test_dataloader)

  history$train_loss <- c(history$train_loss, train_loss)
  history$val_loss <- c(history$val_loss, val_loss)

  cat(sprintf("Epoch %d: Train Loss = %.4f | Val Loss = %.4f\n", epoch, train_loss, val_loss))

  if (val_loss < best_val_loss - min_delta) {
    best_val_loss <- val_loss
    epochs_without_improvement <- 0  # Reset counter
  } else {
    epochs_without_improvement <- epochs_without_improvement + 1
  }

  if (epochs_without_improvement >= patience) {
    cat("Early stopping triggered. Training stopped.\n")
    break
  }
}
```

```{r}

library(ggplot2)
library(tidyr)

plot_loss <- function(history) {
  # Convert history list to a data frame
  df <- data.frame(
    epoch = seq_along(history$train_loss),
    train_loss = history$train_loss,
    val_loss = history$val_loss
  ) %>%
    pivot_longer(cols = c(train_loss, val_loss), 
                 names_to = "Loss_Type", 
                 values_to = "Loss")

  ggplot(df, aes(x = epoch, y = Loss, color = Loss_Type)) +
    geom_line(linewidth = 1) +  # ✅ Fix applied
    geom_point() +
    labs(x = "Epoch", y = "Mean Absolute Error (MAE)", color = "Loss Type") +
    theme_minimal() +
    theme(legend.position = "top") +
    ggtitle("Training and Validation Loss")
}

plot_loss(history)

```

```{r}
# predictions (on test set)
model$eval()  
test_predictions_tensor <- model(scaled_test_features_tensor)  
test_predictions <- as_array(test_predictions_tensor)  # convert tensor to numeric vector
test_labels_vec <- as_array(test_labels_tensor)  # convert true labels to numeric vector

df <- data.frame(
  True_Values = test_labels_vec,
  Predictions = test_predictions
)

ggplot(df, aes(x = True_Values, y = Predictions)) +
  geom_point(color = "blue", alpha = 0.6) +  
  labs(x = "True Values", y = "Predictions", title = "model predictions vs true values") +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")  
```

```{r}

# predict using the entire data

features <- dataset  
labels <- features$y_tilde  
features$y_tilde <- NULL

#  standardize features (use the same mean and std as training)
scaled_features <- scale(features, center = train_means, scale = train_sds)

# convert to torch tensor
scaled_features_tensor <- torch_tensor(as.matrix(scaled_features))
labels_tensor <- torch_tensor(as.numeric(labels))

#  Predictions
model$eval()  
predictions_tensor <- model(scaled_features_tensor)  
predictions <- as_array(predictions_tensor)  
true_values <- as_array(labels_tensor)

df <- data.frame(
  True_Values = true_values,
  Predictions = predictions
)

ggplot(df, aes(x = True_Values, y = Predictions)) +
  geom_point(color = "blue", alpha = 0.6) +  # Scatter plot
  labs(x = "True Values", y = "Predictions", title = "model predictions vs true values") +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")

```

```{r}
# loops over rows (repeats the row 100 times)

test_dataset <- data_2nd_stage
features <- test_dataset %>% select(x1, x2)
Tseq <- evall_N_seq$T

RMSE_vector <- c()

# iterate over 6 rows (like `for i in tqdm([0,1,2,3,4,5])` in takashi's)
walk(1:6, function(i) {
  feature_1 <- features[i, ]

  rep_feature_1 <- feature_1[rep(1, 100), ]  # repeat row
  rep_feature_1 <- cbind(rep_feature_1, evall_N_seq[, 1:3])  # T_1 and T_2 and T_3
  colnames(rep_feature_1) <- c("x1", "x2", "T_1_tilde", "T_2_tilde", "T_3_tilde")  # Rename columns

  # standardize features using the training mean and std
  scaled_feature_1 <- scale(rep_feature_1, center = train_means, scale = train_sds)
  scaled_feature_1_tensor <- torch_tensor(as.matrix(scaled_feature_1))  # Convert to tensor

  #  model predictions
  model$eval()  
  prediction_tensor <- model(scaled_feature_1_tensor)  
  prediction <- as_array(prediction_tensor)  # convert tensor to numeric vector

  # true_Y_tilde
  true_Y_tilde <- (feature_1$x1 + feature_1$x2^2) * log(Tseq)  # formula

  # predictions vs true values
  df <- data.frame(Tseq = Tseq, Prediction = prediction, True_Y = true_Y_tilde)

  p <- ggplot(df, aes(x = Tseq)) +
    geom_line(aes(y = Prediction, color = "Prediction")) +
    geom_line(aes(y = True_Y, color = "True_Y"), linetype = "dashed") +
    labs(x = "T", y = "Y_tilde", title = paste("Row", i, "Prediction vs True")) +
    theme_minimal() +
    scale_color_manual(values = c("Prediction" = "blue", "True_Y" = "black"))

   print(p)
  # compute RMSE
  MSE <- mean((true_Y_tilde - mean(true_Y_tilde) - (prediction - mean(prediction)))^2)
  RMSE <- sqrt(MSE)
  RMSE_vector <<- c(RMSE_vector, RMSE)  
})

print(RMSE_vector) # for 6 rows
```

# Rann_regression_MultiSimulation_train5field_20230504

## Import data

```{r}

data_2nd_stage <- readRDS("./data_20230504/data_2nd_stage.rds")
evall_N_seq = readRDS("./data_20230504/evall_N_seq.rds")
df_train_sim_id <- read_csv("./Output_ann_Train_5field/df_train_sim_id.csv")
```

```{r}

sim_id_train <- unlist(df_train_sim_id[2, ])   # row 1 in python is row 2 in R
sim_id_train <- sim_id_train[-1]

# convert to torch tensor and move to GPU (MPS)
sim_id_train_tensor <- torch_tensor(sim_id_train, device = torch_device("mps"))

# check it's on GPU
print(sim_id_train_tensor)

```

# check if GPU is working

```{r}
# check if GPU is working

mps_is_available <- function() {
  tryCatch({
    tmp <- torch_tensor(1, device = torch_device("mps"))
    TRUE
  }, error = function(e) {
    FALSE
  })
}

# --- Select device: CUDA > MPS > CPU ---
if (cuda_is_available()) {
  device <- torch_device("cuda")
} else if (mps_is_available()) {
  device <- torch_device("mps")
} else {
  device <- torch_device("cpu")
}
cat("Using device:", device$type, "\n")
```

# Multiple input DNN

```{r}

# loop for each simulation 
sim_ids <- unique(data_2nd_stage$sim)
num <- length(sim_ids)

for (j in seq_len(num)) {

  sim_ids_ <- sim_ids[-j]

  sim_id_train <- as.numeric(df_train_sim_id[2, ])
  sim_id_train <- sim_id_train[-1]  # remove first element
  train_sim_ids <- sim_id_train[1:2]

  sim_id_test <- sim_ids[j]
  cat("Processing sim_id_test =", sim_id_test, "\n")

  dataset <- data_2nd_stage %>%
    filter(sim %in% train_sim_ids) %>%
    select(y_tilde, Nk, plateau, b0, T_1_tilde, T_2_tilde, T_3_tilde)

  n <- nrow(dataset)
  train_idx <- sample(seq_len(n), size = floor(0.8 * n))
  train_dataset <- dataset[train_idx, ]
  val_dataset   <- dataset[-train_idx, ]

  train_labels <- train_dataset$y_tilde
  train_features <- train_dataset %>% select(-y_tilde)
  val_labels <- val_dataset$y_tilde
  val_features <- val_dataset %>% select(-y_tilde)

  # standardize

  col_means <- sapply(train_features, mean)
  col_sds   <- sapply(train_features, sd)

  scale_data <- function(df) {
    as.data.frame(scale(df, center = col_means, scale = col_sds))
  }

  scaled_train_features <- scale_data(train_features)
  scaled_val_features   <- scale_data(val_features)

  # convert data to torch tensors ---
  x_train <- torch_tensor(as.matrix(scaled_train_features), dtype = torch_float(), device = device)
  y_train <- torch_tensor(matrix(train_labels, ncol = 1), dtype = torch_float(), device = device)
  x_val   <- torch_tensor(as.matrix(scaled_val_features), dtype = torch_float(), device = device)
  y_val   <- torch_tensor(matrix(val_labels, ncol = 1), dtype = torch_float(), device = device)

  #  neural network model 
  Net <- nn_module(
    "Net",
    initialize = function() {
      # Branch 1
      self$branch1_fc1 <- nn_linear(3, 64)
      self$branch1_fc2 <- nn_linear(64, 64)
      self$branch1_fc3 <- nn_linear(64, 1)
      # Branch 2
      self$branch2_fc1 <- nn_linear(3, 64)
      self$branch2_fc2 <- nn_linear(64, 64)
      self$branch2_fc3 <- nn_linear(64, 1)
      # Branch 3
      self$branch3_fc1 <- nn_linear(3, 64)
      self$branch3_fc2 <- nn_linear(64, 64)
      self$branch3_fc3 <- nn_linear(64, 1)
    },
    forward = function(x) {
      x_main <- x[ , 1:3]
      T <- x[ , 4:6]

      # Branch 1
      b1 <- self$branch1_fc1(x_main) %>% nnf_relu()
      b1 <- self$branch1_fc2(b1) %>% nnf_relu()
      b1 <- self$branch1_fc3(b1)
      b1 <- b1 * T[ , 1, drop = FALSE]

      # Branch 2
      b2 <- self$branch2_fc1(x_main) %>% nnf_relu()
      b2 <- self$branch2_fc2(b2) %>% nnf_relu()
      b2 <- self$branch2_fc3(b2)
      b2 <- b2 * T[ , 2, drop = FALSE]

      # Branch 3
      b3 <- self$branch3_fc1(x_main) %>% nnf_relu()
      b3 <- self$branch3_fc2(b3) %>% nnf_relu()
      b3 <- self$branch3_fc3(b3)
      b3 <- b3 * T[ , 3, drop = FALSE]

      out <- b1 + b2 + b3
      out
    }
  )

  model <- Net()$to(device = device)

  optimizer <- optim_adam(model$parameters, lr = 0.001)

  # training loop 
  epochs <- 50     #500
  batch_size <- 512  
  patience <- 10
  best_val_loss <- Inf
  epochs_no_improve <- 0

  # when running the following code, I get this Warning:
  ### [W MPSFallback.mm:11] Warning: The operator 'aten::sgn.out' is not currently supported on the MPS backend and will fall back to run on the CPU. This may have performance implications. (function operator())

  for (epoch in 1:epochs) {
    model$train()
    optimizer$zero_grad()

    output <- model(x_train)
    loss <- nnf_l1_loss(output, y_train)

    loss$backward()
    optimizer$step()

    model$eval()
    with_no_grad({
      val_output <- model(x_val)
      val_loss <- nnf_l1_loss(val_output, y_val)$item()
    })

    # early stopping 
    if (val_loss < best_val_loss - 1e-6) {
      best_val_loss <- val_loss
      epochs_no_improve <- 0
    } else {
      epochs_no_improve <- epochs_no_improve + 1
    }

    if (epochs_no_improve >= patience) {
      cat("Early stopping at epoch", epoch, "\n")
      break
    }
  }

  ####$$%^^&&&***  End of where I get warning

  # predictions on val set 
  model$eval()
  with_no_grad({
    val_preds <- model(x_val)$to(device = "cpu")$numpy()
  })
  val_preds <- as.vector(val_preds)

  outcome_val <- data.frame(pred = val_preds, true = val_labels)
  outfile_val <- sprintf("./Output_Rann_Train_5field/validation_%s.csv", sim_id_test)
  write.csv(outcome_val, outfile_val, row.names = FALSE)

  # === EONR Estimation ===

  p_corn <- 6.25 / 25.4   
  p_N    <- 1 / 0.453592

  test_dataset <- data_2nd_stage %>% filter(sim == sim_id_test)

  test_features <- test_dataset %>% select(Nk, plateau, b0)

  test_evall_N_seq <- evall_N_seq %>%
    filter(sim == sim_id_test) %>%
    arrange(N)
  Nseq <- test_evall_N_seq$N
  n_row <- nrow(test_dataset)

  estEONR_vector <- numeric(n_row)

  # for each row (field) in the test data
  for (i in 1:n_row) {

    feature_1 <- test_features[i, , drop = FALSE]
    # repeat the row 100 times
    rep_feature_1 <- feature_1[rep(1, 100), ]

    temp <- test_evall_N_seq %>% select(T_1, T_2, T_3)
    names(temp) <- c("T_1_tilde", "T_2_tilde", "T_3_tilde")
    rep_feature_1 <- cbind(rep_feature_1, temp)

    rep_feature_1_scaled <- as.data.frame(mapply(function(col, mean_val, sd_val) {
      (col - mean_val) / sd_val
    }, rep_feature_1, col_means, col_sds, SIMPLIFY = FALSE))

    # convert the scaled data to a torch tensor.
    rep_feature_tensor <- torch_tensor(as.matrix(rep_feature_1_scaled), dtype = torch_float(), device = device)

    model$eval()
    with_no_grad({
      prediction <- model(rep_feature_tensor)$to(device = "cpu")$numpy()
    })
    prediction <- as.vector(prediction)

    MP <- prediction * p_corn - Nseq * p_N
    #  N value that maximizes MP
    estEONR <- Nseq[which.max(MP)]
    estEONR_vector[i] <- estEONR
  }

  outcome <- data.frame(pred = estEONR_vector, true = test_dataset$opt_N)
  outfile_eonr <- sprintf("./Output_Rann_Train_5field/EONR_%s.csv", sim_id_test)
  write.csv(outcome, outfile_eonr, row.names = FALSE)

}

```

### 2025-07-07 Writing/manuscript_CF_continuous.qmd
Writing

# Introduction

# Materials and Method

## Data generation

## Models and estimation procedures

We employ three modeling approaches to estimate the yield response function to nitrogen application rates and to determine the site-specific optimal nitrogen rate. First, we implement artificial neural networks (ANNs) using two variations of the R-learner framework. In the first variation (RANN), both the outcome variable (yield) and the treatment variable (nitrogen rate) are orthogonalized. In the second variation, only the outcome variable is orthogonalized (half_RANN). Next, we apply an S-learner ANN, in which yield is modeled as a function of the nitrogen rate and a set of observed covariates (X), without orthogonalization. In addition to these ANN-based methods, we also use a causal forest (CF), implemented as a double machine learning approach following the R-learner framework. Finally, we include a random forest (RF) model using the R-learner approach with orthogonalization of the outcome variable only (half-RRF).

The econometric model to estimate the R-learner CF is as following:

$$
Y=\sum_{k=1}^K \theta_k(X) \cdot S_k(N)+g(X)+v
$$
where $Y$ is corn yield, $N$ is nitrogen rate (treatment variable), and $X$ is a collection of observed variables, $\alpha, \beta$, and $\gamma$. In this equation, $g(X)$ is the nuisance function.
Here the treatment variable (nitrogen rate) is split into $K$ basis variables. Then, each of the basis variables are orthogonalized by cross-fitting using $X$ as covariates via boosted regression forest. Let the treatment basis variables be denoted as $S_k \tilde{(T)}$ for $k=1, \ldots, K$. Then, the second stage solves the following problem using the CF algorithm:

$$
\operatorname{Min} \sum_{i=1}^N\left[{Y}_i-\sum_{k=1}^K \theta_k\left(X_i\right) \tilde{S}_k\left(N_i\right)\right]
$$

Once $\theta_k\left(X_i\right)$ are estimated, then the conditional treatment effect of moving from one rate $\left(N_0\right)$ to another $\left(N_1\right)$ for observation $i$ can be obtained by $\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_1\right)-\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_0\right)$.

Profit impact of moving from $N_0$ to $N_1$ is

$$
P_{\text {corn }} \sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_1\right)-P_{\text {corn }} \sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_0\right)-P_N\left(N_1-N_0\right)
$$

A grid search over a sequence of $N$ values is used to identify the profit-maximizing N rate.

The S-learner ANN approach estimates the following model non-parametrically without imposing any functional forms on the relationships between the variables.

$$
Y=f(X, N)+v
$$
Once $f(X, N)$ is estimated (denote it as $\hat{f}(X, N)$ ), then the conditional treatment effect of moving from one rate ( $N_0$ ) to another ( $N_1$ ) can be obtained by $\hat{f}\left(X, N_1\right)-\hat{f}\left(X, N_0\right)$. The associated change in profit is $P_{\text {corn }} \hat{f}\left(X, N_1\right)-P_{\text {corn }} \hat{f}\left(X, N_0\right)-P_N\left(N_1-N_0\right)$. Similar to the CF approach, a grid search is used to identify numerically the optimal nitrogen rate.

The R-learner ANN (RANN) approach estimates the site-specific nitrogen response function by first orthogonalizing the outcome variable and then modeling the residuals using a neural network. Specifically, the outcome residual $y_i-\hat{g}\left(X_i\right)$ is regressed on orthogonalized treatment basis variables $\tilde{T}_1, \tilde{T}_2, \tilde{T}_3$, interacted with covariates $X$, using a structured artificial neural network.
The model architecture includes three separate branches, each corresponding to one of the orthogonalized basis variables for the nitrogen rate. Each branch processes the covariates $X$ and multiplies the output by one of the treatment basis variables. The final yield prediction is the sum of the outputs from all three branches:

$$
\hat{Y}_i=f_1\left(X_i\right) \cdot \tilde{T}_{1 i}+f_2\left(X_i\right) \cdot \tilde{T}_{2 i}+f_3\left(X_i\right) \cdot \tilde{T}_{3 i}
$$
where $f_k(\cdot)$ are the outputs from each neural network branch, and $\tilde{T}_{k i}$ are the orthogonalized treatment basis variables for observation $i$. These basis variables are obtained by cross-fitting regression forests of each treatment basis on $X$, removing the component of the treatment that is predictable from the covariates.

Once the model is trained, we use it to predict yield for a grid of nitrogen rates. These predictions are then used to compute the marginal profit function:

$$
\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
$$
The site specific nitrogen rate is the value of $N$ that maximizes this marginal profit function. This procedure is repeated for each observation in the test set to obtain field-specific EONR estimates. As with the other approaches, the profit-maximizing nitrogen rate is identified by conducting a grid search over a sequence of nitrogen values.

The half R-learner artificial neural network (Half-RANN) estimates the site-specific nitrogen response function by orthogonalizing only the outcome variable while keeping the original (non-orthogonalized) nitrogen rate as the treatment input. Specifically, the orthogonalized outcome residual $\tilde{Y}_i=Y_i-\hat{g}\left(X_i\right)$, where $\hat{g}\left(X_i\right)$ is a nuisance function estimated from covariates $X$, is modeled using a neural network that takes the original treatment $N_i$ and the covariates $X_i$ as inputs. The model learns a function $f(X, N)$ using a fully connected neural network:

$$
\tilde{Y}_i=f\left(X_i, N_i\right)+\varepsilon_i
$$
Here, $X$ includes three observed variables that characterize the yield response function: the nitrogen level at which yield plateaus, the plateau yield level, and the intercept yield level at zero nitrogen. The nitrogen rate $N$ is included directly as one of the inputs to the network without basis transformation or orthogonalization. 
Once the network is trained, predicted yield responses $\hat{Y}(X, N)$ are obtained across a grid of nitrogen rates. These predictions are used to compute marginal profit:

$$
\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
$$
As with other models, a grid search over candidate $N$ values is used to identify the profit-maximizing rate.

The half R-learner random forest (Half-RRF) approach estimates the site-specific yield response to nitrogen by orthogonalizing the outcome variable and modeling the resulting residuals using a random forest. In this approach, only the outcome $Y$ is orthogonalized with respect to observed covariates $X$, while the nitrogen treatment variable $N$ enters the model in its original form. The goal is to capture treatment effect heterogeneity while preserving flexibility in the treatment-response relationship. 
The model estimates the following equation:

$$
\tilde{Y}_i=f\left(X_i, N_i\right)+\varepsilon_i
$$
where $\tilde{Y}_i=Y_i-\hat{g}\left(X_i\right)$ is the residualized outcome, and $f(\cdot)$ is a random forest fitted to the nitrogen rate $N$ and covariates $X$.
The model is trained using repeated K-fold cross-validation and hyperparameter tuning through grid search. Once trained, yield predictions $\hat{Y}(X, N)$ are generated for a grid of nitrogen rates. These predictions are used to compute the marginal profit function:

$$
\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
$$
This yields site-specific recommendations for nitrogen application.

## Performance Measurement

# Results and Discussions

# Conclusion

### 2025-07-15 Writing/manuscript_CF_continuous.qmd
working on writing
Simukation Scenarios added

@@
 
+For each simulation round, we evaluate model performance by calculating the root mean squared error (RMSE) of EONR predictions and the associated profit loss for each method.
+The RMSE of EONR predictions in a given simulation round is computed as follows:
+
+$$
+R M S E_{E O N R}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{E O N R}_i-E O N R_i\right)^2}
+
+$$
+To assess the predictive accuracy and economic relevance of these modeling approaches, we design a set of simulation experiments that reflect realistic agronomic conditions.
+These simulations serve as a controlled environment for evaluating model performance in estimating yield response functions and recommending optimal nitrogen rates.
+
+## Simulation Scenarios
+
+To evaluate the performance of the proposed models we design Monte Carlo simulations that mimic a typical on-farm precision experimentation (OFPE) setting.
+These simulations examine the ability of each method to accurately estimate the site-specific yield response to nitrogen application and the EONR under varying data availability conditions.
+
+Our simulation consists of 500 independently generated fields, each representing a distinct agronomic environment with field-specific parameters that shape the crop’s response to nitrogen application.
+These parameters vary systematically across fields to introduce heterogeneity in yield responses.
+The true yield response in each field follows a quadratic-plateau functional form, a widely used structure for modeling nitrogen response curves in agronomic studies.
+
+For each simulation round, one field is selected as the test field.
+The remaining fields are used to construct the training set in three scenarios: (1) one-field training, (2) five-field training, and (3) ten-field training.
+These scenarios allow us to assess how model performance changes with the size and diversity of training data.
+Within each round, the model is trained on the designated number of training fields and then used to estimate the yield response function and the site-specific EONR for the test field.
+
+The experimental design within each field consists of randomized nitrogen treatment rates applied across plots.
+The nitrogen rates vary between fields and are drawn to reflect realistic trial setups that span a range of agronomic optima.
+For evaluation purposes, we use a grid of candidate nitrogen rates to compute predicted yield and associated marginal profit for each observation.
+The predicted EONR is the nitrogen rate that maximizes predicted profit.
+
 # Results and Discussions

### 2025-07-15 codes/R/CF_Continuous_Loop.Rmd
some restructuring

# Objectives

+ Check the CF performance by training on one, five and ten fields

# Preparation

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
#library(here)
library(reticulate)

#--- R functions ---#
here("Shared_codes/R") %>%
  fs::dir_ls(., full.names = TRUE) %>%
  lapply(source)

#--- python functions ---#
use_condaenv("cf_conda311", required = TRUE)
py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
source_python("Shared_codes/Python/run_CF_c.py")

```

## Data

```{r}
# turn off multi-threading 
Sys.setenv(OMP_NUM_THREADS = 1, MKL_NUM_THREADS = 1)

# load data only once, then drop
raw_sim_data <- readRDS("Data/raw_sim_data.rds")
rm(list = setdiff(ls(), c("raw_sim_data","np","run_CF_c_py","gm_wrapper_functions_you_need")))
gc()
```

# Train CF (Train on five fields. We can simply change 5 to other numbers)

```{r}

x_vars <- c("Nk", "plateau", "b0")
T_formula <- yield ~ s(N, k = 4)

output_dir <- "Results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

res_list_one <- lapply(1:500, function(test_sim_id) {
  result_file <- file.path(output_dir, paste0("sim_", test_sim_id, ".rds"))

  if (file.exists(result_file)) {
    message(sprintf("Skipping sim %d (already saved)", test_sim_id))
    return(NULL)
  }

  tryCatch({
    # Train on five fields
    train_sim_ids <- sample(setdiff(1:500, test_sim_id), 5)
    sim_id_ls <- c(train_sim_ids, test_sim_id)

    data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% sim_id_ls, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    train_data <- data[sim != test_sim_id, ]

    test_data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    T_info <- prepare_T_mat(T_formula, train_data)
    Y <- train_data[, yield]
    X <- train_data[, ..x_vars] %>% as.matrix()
    W <- X
    te_hat_cf <- run_CF_c_py(Y, T_info$T_sp, X, W, n_estimators = 2000)

    te_info     <- get_te(te_hat_cf, test_data, x_vars)
    T_seq       <- train_data[, seq(min(N), max(N), length.out = 200)]
    response_dt <- find_response_semi(T_seq, T_info, te_info)

    pCorn <- 6.25 / 25.4
    pN    <- 1 / 0.453592

    ss_eonr <- response_dt[
      , profit := est * pCorn - pN * T
    ][
      , .SD[which.max(profit)], by = aunit_id
    ][
      , .(aunit_id, opt_N_hat = T)
    ]

    true_eonr <- raw_sim_data$reg_data[[1]] %>%
      .[sim == test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table() %>%
      .[, .(aunit_id, b0, b1, b2, Nk)] %>%
      .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
      .[, opt_N := pmin(Nk, pmax(0, opt_N))] %>%
      .[, .(aunit_id, opt_N)]

    combined <- ss_eonr[true_eonr, on = "aunit_id"]
    combined[, sim := test_sim_id]

    saveRDS(combined, file = result_file)

    combined
  }, error = function(e) {
    message(sprintf("Error with sim %d: %s", test_sim_id, e$message))
    return(data.table(sim = test_sim_id, aunit_id = integer(), opt_N_hat = numeric(), opt_N = numeric()))
  })
})
```

### 2025-07-15 presentation/iisa-2025-TM.qmd
some restructuring

---
title: "Single-orthogonalized Neural Network for Site-specific Yield Response Function Estimation"
author:
  - name: Taro Mieno
    affiliations:
      - name: University of Nebraska Lincoln
  - name: Mona Mousavi
    affiliations:
      - name: University of Nebraska Lincoln
  - name: David Bullock
    affiliations:
      - name: University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: AnnArbor
format: 
  revealjs: 
    theme: [default, custom.scss]
    embed-resources: true
    html-math-method: mathjax
    fontsize: 1.4em
    callout-icon: false
    scrollable: true
    echo: true
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

## Background

+ Nitrogen rate management for agricultural production is an important issue (economic, environmental, health)

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather)

+ On-farm field experiment (a way)
  + run trials and collect data
  + <span style = "color: blue;"> estimate how yield responds to nitrogen based on observed characteristics </span>
  + identify site-specific optimal nitrogen rates based on the estimated model

+ Existing prediction oriented approaches (simple application of random forest, neural network) tend to fail for identifying the heterogeneous causal impact of a treatment

+ Existing causal machine learning approaches  (e.g., r-learner) specialize in identifying the impact of a binary treatment (but, we want yield response <span style = "color: blue;"> curve </span>)

```{=tex}
\begin{align*}
\Delta Yield = f(N_1| X) - f(N_0| X)\;\; \mbox{, where } N_1 > N_0
\end{align*}
```

## Research Objective

+ Propose an ML approach that can identify heterogenous treatment effects well when the treatment variable is continuous

+ Test the performance of the approach against existing approaches that are commonly used via Monte Carlo simulation

## Double machine learning and our approach

::: {.panel-tabset}

### Recap on Double-ML

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

::: {.fragment}
:::{.callout-note title="Steps"}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $T$ on $X$ and $W$ and predict $T[Y|X,W]$ (denoted as $h(X,W)$)
3. Regress $Y-f(X,W)$ on $T-h(X,W)$
:::
:::
<!--end of the fragment-->

### Single-orthogonalized (our approach)

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

::: {.fragment}
:::{.callout-note title="Steps"}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
3. Regress $Y-f(X,W)$ on $T$
:::
:::
<!--end of the fragment-->

::: {.fragment}
:::{.callout-important title="Reasoning"}
+ The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$. 
+ This is particularly effective if the share of $g(X,W)$ in the total variation of $Y$ is greater relative to $\theta(X, T)$.
:::
:::
<!--end of the fragment-->

### Double-orthogonalized?

```{=tex}
\begin{align*}
\mbox{Model 1: } Y & = \theta(X)\cdot T + g(X, W) + v \\
\mbox{Model 2: } Y & = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ In model 1, Double-orthogonalization is fine because the model is linear in $T$.
+ In model 2, Orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.
:::
<!--end of panel-->

## Simulation: Data Generation

::: {.panel-tabset}

### Data generating process

Corn yield is assumed to follow the quadratic plateau functional form as follows:

```{=tex}
\begin{align*}
Y_{j, i}=f\left(N, \alpha_{j, i}, \alpha_{j, i}, \gamma{j, i}, \tau{j, i}\right)= \begin{cases}\alpha_{j, i}+\beta_{j, i} N+\gamma_{j, i} N^2+\varepsilon_{j, i}, & N<\tau_{j, i} \\ \alpha_{j, i}+\beta_{j, i} \tau_{j, i}+\gamma_{j, i} \tau_{j, i}^2+\varepsilon_{j, i}, & N \geq \tau_{j, i}\end{cases}
\end{align*}
```

+ $Y$: Corn yield
+ $N$: Nitrogen rate
+ $\alpha, \beta, \gamma, \tau$: parameters (consider them observable soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$

### On-farm trial (nitrogen rate)

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_trial.png") 
```

### A parameter ($\tau$)

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_par.png") 
```

### Corn yield

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_yield.png") 
```

```{r}
#| echo: false
#| eval: false
data <- 
  readRDS(here::here("presentation/sim_data_1.rds"))$data[[1]] %>%
  sf::st_as_sf()

g_yield <- 
  ggplot(data = data) +
  geom_sf(aes(fill = yield)) +
  scale_fill_viridis_c(name = "Corn Yield (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

ggsave("presentation/g_yield.png", g_yield, dpi = 300, height = 4)

g_trial <-
  ggplot(data = data) +
  geom_sf(aes(fill = factor(N_tgt))) +
  scale_fill_viridis_d(name = "N rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14)
  )

ggsave("presentation/g_trial.png", g_trial, dpi = 300, height = 4)

# library(latex2exp)
# plot(latex2exp::TeX("\\eta"))

g_par <-
  ggplot(data = data) +
  geom_sf(aes(fill = Nk)) +
  scale_fill_viridis_c(name = "") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

ggsave("presentation/g_par.png", g_par, dpi = 300, height = 4)
```

:::
<!--end of panel-->

## Simulation: Other Components

::: {.panel-tabset}

### Models

```{=tex}
\begin{align*}
Y_{i,k} = f(N_{i,k}, X_{i,k}) + v_{i,k}
\end{align*}
```

where $i$ and $k$ indicate a cell and field, respectively.

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

<!-- ### Estimation approaches

<span style = "color: blue;"> Single-orthogonalized </span>

1. Regress $Y$ on $X$ ($N$ not included) using random forest, predict $Y$ conditional on $X$, and subtract it from $Y$ (orthogonalization)
2. Regress the orthogonalized $Y$ on $N$ and $X$ using neural network

<br>

<span style = "color: blue;">Non-orthogonalized  </span>

Regress $Y$ on $N$ and $X$ using neural network

<br>

<span style = "color: blue;">Double-orthogonalized </span>

1. Regress $Y$ on $X$ ($N$ not included) using random forest, predict $Y$ conditional on $X$, and subtract it from $Y$ (orthogonalization)
2. Regress $T$ on $X$ ($N$ not included) using random forest, predict $T$ conditional on $X$, and subtract it from $T$ (orthogonalization)
3. Regress the orthogonalized $Y$ on orthogonalized $N$ and $X$ using neural network -->

### Scenarios

+ Estimation approaches:

```{r}
#| echo: false
data.table(
  Approach = c("Double-orthogonalized", "Single-orthogonalized", "Non-orthogonalized"),
  "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "NA"),
  "T-orthogonalization" = c("Randome Forest", "NA", "NA"),
  "Final stage" = c("Neural Network", "Neural Network", "Neural Network")
) %>%
flextable() %>%
fontsize(size = 16, part = "all") %>%
autofit()
```

+ Number of fields from which data is combined: 1, 5, and 10

### Site-specific EONR estimation

For a given price of corn and nitrogen (denoted $P_{Corn}$ and $P_{N}$, respectively), economically optimal nitrogen rate (EONR) is found for each cell of all the fields by solving the following profit maximization problem:

```{=tex}
\begin{align*}
\widehat{EONR}_{i,k} = Argmax_{N} \;\; P_{Corn} \times \widehat{f}(N, X_{i,k}) - P_{N} \times N
\end{align*}
```

+ $\widehat{f}$: estimated yield response curve

### Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

```{=tex}
\begin{align*}
RMSE = \sqrt{\sum_i^N \sum_k^K(EONR_{i,k} - \widehat{EONR_{i,k}})^2/NK}
\end{align*}
```

:::
<!--end of panel-->

## Results

## Conclusion

### 2025-07-15 presentation/iisa-2025-beamer-TM.qmd
some restructuring

---
title: "Single-orthogonalized Neural Network for Site-specific Yield Response Function Estimation"
author:
  - name: Taro Mieno$^*$
  - name: Mona Mousavi$^*$
  - name: David Bullock$^\dagger$
institute:
  - $^*$ University of Nebraska Lincoln
  - $^\dagger$ University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
header-includes: 
- \logo{\ifnum\thepage>1\includegraphics[width=1cm]{nebraska-n.jpg}\fi}
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

# Background and Research Objectives

## Background

+ Nitrogen rate management for agricultural production

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather)

+ On-farm field experiment (a way)
  + Run on-farm agronomic trials and collect data
  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
  + Identify site-specific optimal nitrogen rates

+ An increasing amount of data from on-farm field trials has been accumulated over the years.

+ Estimating yield response functions using data pooled from multiple fields can provide significant agronomic and economic insights.

## Motivation

+ When on-farm trial data is pooled, the treatment variable becomes "continuous" rather than discrete.

+ Existing approaches—such as direct applications of random forests or neural networks—often tend to fail to capture heterogeneous treatment effects of a continuous variable (Kakimoto et al., 2022).

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.

## Research Objective

+ Find a causal machine learning approach capable of effectively identifying heterogeneous treatment effects tailored for yield response function estimation using data from on-farm trials, where the treatment variable is continuous and randomized.

+ Evaluate the performance of the proposed method through Monte Carlo simulations, comparing it against commonly used existing approaches.

# A New Approach

## Recap on Double-ML

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

. . .

:::{.callout-note title="Steps" icon=false}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $T$ on $X$ and $W$ and predict $T[Y|X,W]$ (denoted as $h(X,W)$)
3. Regress $Y-f(X,W)$ on $T-h(X,W)$
:::

## Single-orthogonalized (our approach)

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

. . .

:::{.callout-note title="Steps" icon=false}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
3. Regress $Y-f(X,W)$ on $T$
:::

. . .

:::{.callout-important title="Reasoning" icon=false}
The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$. 
:::

## Double-orthogonalized?

```{=tex}
\begin{align*}
\mbox{Model 1: } Y & = \theta(X)\cdot T + g(X, W) + v \\
\mbox{Model 2: } Y & = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ In model 1, Double-orthogonalization is fine because the model is linear in $T$.
+ In model 2, Orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.

# Simulation

## Corn yield

Corn yield is assumed to follow the quadratic plateau functional form as follows:

```{=tex}
\begin{align*}
Y=f\left(N, \alpha, \beta{i, k}, \gamma, \tau\right)= \begin{cases}\alpha+\beta N+\gamma N^2+\varepsilon, & N<\tau \\ \alpha+\beta \tau+\gamma \tau^2+\varepsilon, & N \geq \tau\end{cases}
\end{align*}
```

+ $Y$: Corn yield
+ $N$: Nitrogen rate
+ $\alpha, \beta, \gamma, \tau$: parameters (consider them observable soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$

## On-farm trial (nitrogen rate)

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_trial.png")
```

## A parameter ($\tau$)

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_par.png")
```

## Corn yield

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_yield.png")
```

```{r}
#| echo: false
#| eval: false
data <-
  readRDS(here::here("presentation/sim_data_1.rds"))$data[[1]] %>%
  sf::st_as_sf()

g_yield <-
  ggplot(data = data) +
  geom_sf(aes(fill = yield)) +
  scale_fill_viridis_c(name = "Corn Yield (kg/ha)  ") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2, "cm"),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  )

ggsave("presentation/g_yield.png", g_yield, dpi = 300, width = 10, height = 6)

g_trial <-
  ggplot(data = data) +
  geom_sf(aes(fill = factor(N_tgt))) +
  scale_fill_viridis_d(name = "N rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  )

ggsave("presentation/g_trial.png", g_trial, dpi = 300, width = 10, height = 6)

# library(latex2exp)
# plot(latex2exp::TeX("\\eta"))

g_par <-
  ggplot(data = data) +
  geom_sf(aes(fill = Nk)) +
  scale_fill_viridis_c(name = "") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  )

ggsave("presentation/g_par.png", g_par, dpi = 300, width = 10, height = 6)
```

## Models

```{=tex}
\begin{align*}
Y_{i,k} = f(N_{i,k}, X_{i,k}) + v_{i,k}
\end{align*}
```

where $i$ and $k$ indicate a cell and field, respectively.

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

## Scenarios

+ Estimation approaches:

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("f_est.png")
```

```{r}
#| echo: false
#| eval: false
f_est <-
  data.table(
    Approach = c("Double-orthogonalized", "Single-orthogonalized", "Non-orthogonalized"),
    "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "NA"),
    "T-orthogonalization" = c("Randome Forest", "NA", "NA"),
    "Final stage" = c("Neural Network", "Neural Network", "Neural Network")
  ) %>%
  flextable() %>%
  fontsize(size = 16, part = "all") %>%
  autofit()

flextable::save_as_image(f_est, here::here("presentation/f_est.png"), res = 300)
```

+ Number of fields from which data is combined: 1, 5, and 10

## Site-specific EONR estimation

For a given price of corn and nitrogen, economically optimal nitrogen rate (EONR) is found for each cell of all the fields by solving the following profit maximization problem:

```{=tex}
\begin{align*}
\widehat{EONR}_{i,k} = Argmax_{N} \;\; P_{Corn} \times \widehat{f}(N, X_{i,k}) - P_{N} \times N
\end{align*}
```

+ $P_{Corn}$: corn price
+ $P_N$: nitrogen price
+ $\widehat{f}$: yield response curve estimated by the model under evaluation

## Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

```{=tex}
\begin{align*}
RMSE = \sqrt{\sum_i^N \sum_k^K(EONR_{i,k} - \widehat{EONR_{i,k}})^2/NK}
\end{align*}
```

# Results and Conclusions

## Simulation Results

```{r}
#| eval: false
single_orth <-
  readRDS(here::here("results/Half_RANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Single"]

double_orth <-
  readRDS(here::here("results/RANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Double"]

no_orth <-
  readRDS(here::here("results/simple_ANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Non"]

all_results <- rbind(single_orth, double_orth, no_orth)

rmse_data <-
  all_results[, .(rmse = sqrt(sum((true - pred)^2) / .N)), by = .(est_type, field, sim)] %>%
  .[, field_text := case_when(
    field == "1_field" ~ "Number of fields: 1",
    field == "5_fields" ~ "Number of fields: 5",
    field == "10_fields" ~ "Number of fields: 10",
  )] %>%
  .[, field_text := factor(field_text, levels = c("Number of fields: 1", "Number of fields: 5", "Number of fields: 10"))] %>%
  .[, est_type := factor(est_type, levels = c("Single", "Double", "Non"))]

g_results <-
  ggplot(data = rmse_data) +
  geom_histogram(aes(x = rmse), binwidth = 3) +
  facet_grid(est_type ~ field_text) +
  xlab("RMSE by iteration") +
  ylab("Count") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    strip.text = element_text(size = 20)
  )

ggsave("presentation/g_results.png", g_results, dpi = 300, height = 7.5)
```

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("g_results.png")
```

## Conclusion

+ The single-orthogonalized approach performed better than the non-orthogonalized (the common approach taken by researchers for yield response function estimation)
+ Double-orthogonalized approach performs poorly and its performance does not improve even if the number of observations increases (as expected)

## Future extensions

+ More realistic crop yield simulations using crop growth models
+ Other specifications of the approaches
  + Models other than NN for the final stage (e.g., random forest, boosted regression forest, etc)
  + Models other than RF for Y-orthogonalization
+ Is there any way to make the estimation more efficient taking into account the spatial correlation of the error term? (like spatial error model)

### 2025-07-15 codes/R/CF_Continuous_Loop.Rmd
minor edits

@@
 library(mgcv)
-#library(here)
+# library(here)
 library(reticulate)
@@
 #--- R functions ---#
-here("Shared_codes/R") %>%
+here("codes/R") %>%
   fs::dir_ls(., full.names = TRUE) %>%
@@
 py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
-source_python("Shared_codes/Python/run_CF_c.py")
-
+source_python("codes/Python/run_CF_c.py")
 ```
@@
 # load data only once, then drop
-raw_sim_data <- readRDS("Data/raw_sim_data.rds")
-rm(list = setdiff(ls(), c("raw_sim_data","np","run_CF_c_py","gm_wrapper_functions_you_need")))
+raw_sim_data <- readRDS("data/raw_sim_data.rds")
+rm(list = setdiff(ls(), c("raw_sim_data", "np", "run_CF_c_py", "gm_wrapper_functions_you_need")))
 gc()
@@
 ```{r}
-
 x_vars <- c("Nk", "plateau", "b0")
@@
 
-output_dir <- "Results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
+output_dir <- "results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
+output_dir <- "results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
 if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
@@
 
-  tryCatch({
+  tryCatch(
+    {
       # Train on five fields
@@
       ][
-      , .SD[which.max(profit)], by = aunit_id
+        , .SD[which.max(profit)],
+        by = aunit_id
       ][
@@
       combined
-  }, error = function(e) {
+    },
+    error = function(e) {
       message(sprintf("Error with sim %d: %s", test_sim_id, e$message))
       return(data.table(sim = test_sim_id, aunit_id = integer(), opt_N_hat = numeric(), opt_N = numeric()))
-  })
+    }
+  )
 })

### 2025-07-17 Writing/manuscript_CF_continuous.qmd
comments added to the manuscript

@@
+---
+title: "temp"
+format:
+  html: default
+execute:
+  echo: false
+---
+
+```{r}
+#| include: false
+library(data.table) 
+library(dplyr) 
+library(flextable) 
+library(officer) 
+```
+
 # Introduction
 
-# Materials and Method
+# Method
+
+We run Monte Carlo simulation to test the performance of ...
 
@@
 ## Models and estimation procedures
+
+<span style = "color: blue;"> Explain the types of estimation you do first in general.
+You have double-orthogonalized, single-orthogonalized, and non-orthogonalized.
+The difference among the models we try are in how we do the orthogonalization and final stage estimation.
+The table below shows all the cases you consider.
+</span>
+
+<span style = "color: blue;"> Use tex and align environment for equations like below instead of using $$. </span>
+
+```{=tex}
+\begin{align*}
+Y = f(X)
+\end{align*}
+```
+
+```{r}
+#| include: false
+f_est <-
+  data.table(
+    "Approach Label" = c("DO-NN", "SO-RFF", "SO-NN", "DO-CF","NO-NN", "NO-RF"),
+    "Approach Type" = c("Double-orthogonalized", "Single-orthogonalized", "Single-orthogonalized", "Double-orthogonalized", "Non-orthogonalized", "Non-orthogonalized"),
+    "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "Randome Forest", "Randome Forest", "NA", "NA"),
+    "T-orthogonalization" = c("Randome Forest", "NA", "NA", "Randome Forest", "NA", "NA"),
+    "Final stage" = c("Neural Network", "Random Forest", "Neural Network", "Causal Forest", "Neural Network", "Random Forest")
+  ) %>%
+  flextable() %>%
+  fontsize(size = 16, part = "all") %>%
+  autofit()
+
+flextable::save_as_image(f_est, here::here("writing/f_est.png"), res = 300)
+```
+
+```{r}
+#| echo: false
+#| fig-width: 4
+#| out-width: 4in
+knitr::include_graphics(here::here("writing/f_est.png"))
+```
 
@@
 Finally, we include a random forest (RF) model using the R-learner approach with orthogonalization of the outcome variable only (half-RRF).
+
+<span style = "color: red;"> Right now, you are explaining how EONR is calculated by method.
+But, all the estimation cases follow the same procedure.
+For all cases, you have $\hat{y} = \hat{f}(N:X)$.
+So, you can just this function in explaining how to get EONR for all the cases.
+You are repeating the same equations of profit three times.
+You should recognized that problem.
+Think carefully how to structure the section.
+Don't just write.
+</span>
+
+### Non-orthogonalized ANN (NO-NN) and RF (NO-RF)
+
+<span style = "color: blue;"> I am adding NO-RF as one of the method to consider. Write code for this and get the results. It should be easy. </span>
+
+Neither of Y or X is orthogonalized
+
+The S-learner ANN approach estimates the following model non-parametrically without imposing any functional forms on the relationships between the variables.
+
+$$
+Y=f(X, N)+v
+$$
+
+<span style = "color: blue;"> As I said, remove the sentences below. </span>
+
+Once $f(X, N)$ is estimated (denote it as $\hat{f}(X, N)$ ), then the conditional treatment effect of moving from one rate ( $N_0$ ) to another ( $N_1$ ) can be obtained by $\hat{f}\left(X, N_1\right)-\hat{f}\left(X, N_0\right)$.
+The associated change in profit is $P_{\text {corn }} \hat{f}\left(X, N_1\right)-P_{\text {corn }} \hat{f}\left(X, N_0\right)-P_N\left(N_1-N_0\right)$.
+Similar to the CF approach, a grid search is used to identify numerically the optimal nitrogen rate.
+
+### CF-continuous
+
+<span style = "color: blue;"> This is an inaccurate and insufficient explanations. Add what I wrote for CF in the manuscript I shared. </span>
 
@@
 
+<span style = "color: blue;"> This is a description for binary treatment case. After running CF-continuous, you are getting $\hat{y} = \hat{f}(N:X)$</span>
+
 Once $\theta_k\left(X_i\right)$ are estimated, then the conditional treatment effect of moving from one rate $\left(N_0\right)$ to another $\left(N_1\right)$ for observation $i$ can be obtained by $\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_1\right)-\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_0\right)$.
@@
 
-The S-learner ANN approach estimates the following model non-parametrically without imposing any functional forms on the relationships between the variables.
-
-$$
-Y=f(X, N)+v
-$$
-Once $f(X, N)$ is estimated (denote it as $\hat{f}(X, N)$ ), then the conditional treatment effect of moving from one rate ( $N_0$ ) to another ( $N_1$ ) can be obtained by $\hat{f}\left(X, N_1\right)-\hat{f}\left(X, N_0\right)$.
-The associated change in profit is $P_{\text {corn }} \hat{f}\left(X, N_1\right)-P_{\text {corn }} \hat{f}\left(X, N_0\right)-P_N\left(N_1-N_0\right)$.
-Similar to the CF approach, a grid search is used to identify numerically the optimal nitrogen rate.
+### Double-orthogonalized ANN (DO-ANN) and RF (DO-RFF)
 
@@
 As with the other approaches, the profit-maximizing nitrogen rate is identified by conducting a grid search over a sequence of nitrogen values.
+
+### Single-orthogonalized NN (SO-NN) and (SO-RF)
+
+<span style = "color: blue;"> Once you explain how SO works. The difference between SO-NN and SO-RF is just the final stage estimation. Write that way to reduce unnecessary writing</span>
 
@@
 
+## Estimation of site-specific EONR
+
 ## Performance Measurement
@@
 
-$$
-R M S E_{E O N R}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{E O N R}_i-E O N R_i\right)^2}
+```{=tex}
+\begin{align*}
+RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
+\end{align*}
+```
 
-$$
 To assess the predictive accuracy and economic relevance of these modeling approaches, we design a set of simulation experiments that reflect realistic agronomic conditions.

### 2025-07-18 codes/R/CF_Continuous_Loop.Rmd
Update on CF (doing loop on fixed splits)

--- 2025-07-15 codes/R/CF_Continuous_Loop.Rmd
@@
 library(mgcv)
-#library(here)
+# library(here)
 library(reticulate)
+library(readr)
+library(dplyr)
+library(tidyr)
 
 #--- R functions ---#
-here("Shared_codes/R") %>%
-  fs::dir_ls(., full.names = TRUE) %>%
-  lapply(source)
+# here("codes/R") %>%
+#   fs::dir_ls(., full.names = TRUE) %>%
+#   lapply(source)
 
@@
 py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
-source_python("Shared_codes/Python/run_CF_c.py")
+source_python("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
+```
+
+```{r}
+# project‐root path:
+proj_root <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous"
+
+# “.here” file
+here_file <- file.path(proj_root, ".here")
+if (!file.exists(here_file)) {
+  message("Creating .here in ", proj_root)
+  file.create(here_file)
+}
 
@@
 # load data only once, then drop
-raw_sim_data <- readRDS("Data/raw_sim_data.rds")
-rm(list = setdiff(ls(), c("raw_sim_data","np","run_CF_c_py","gm_wrapper_functions_you_need")))
+library(here)
+raw_sim_data <- readRDS(here("data", "raw_sim_data.rds"))
+rm(list = setdiff(ls(), c("raw_sim_data", "np", "run_CF_c_py", "gm_wrapper_functions_you_need")))
 gc()
+```
+
+```{r}
+source("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/R/Utility.R")
 ```
@@
 
+# how many fields to train on?
+n_fields <- 5
+
+# ─── PATHS via here() ──────────────────────────────────────────────────────────
+
+# the right split file under <project_root>/data/train_test_split
+splits_file <- here(
+  "data", "train_test_split",
+  sprintf("train_test_splits_%dfields.csv", n_fields)
+)
+output_dir <- here(
+  "results",
+  sprintf("CF_outcome_%d_fields", n_fields)
+)
+if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
+
+# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────
+
+train_test_splits <- read_csv(splits_file)
+
+# ─── MODEL LOOP ────────────────────────────────────────────────────────────────
+
 x_vars    <- c("Nk", "plateau", "b0")
@@
 
-output_dir <- "Results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
-if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
-
-res_list_one <- lapply(1:500, function(test_sim_id) {
+# cf_results <- lapply(train_test_splits$test_id, function(test_sim_id)
+cf_results <- lapply(train_test_splits$test_id, function(test_sim_id) {
   result_file <- file.path(output_dir, paste0("sim_", test_sim_id, ".rds"))
-
   if (file.exists(result_file)) {
@@
   tryCatch({
-    # Train on five fields
-    train_sim_ids <- sample(setdiff(1:500, test_sim_id), 5)
+
+    split_row     <- train_test_splits[train_test_splits$test_id == test_sim_id, ]
+    train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
+
     sim_id_ls     <- c(train_sim_ids, test_sim_id)
 
+    # assemble train & test data
     data <- raw_sim_data$reg_data[[1]] %>%
@@
     train_data <- data[sim != test_sim_id, ]
-
     test_data  <- raw_sim_data$reg_data[[1]] %>%
-      .[sim %in% test_sim_id, ] %>%
+      .[sim == test_sim_id, ] %>%
       .[, .(sim, data)] %>%
@@
 
+    # fit causal forest
     T_info     <- prepare_T_mat(T_formula, train_data)
@@
     te_hat_cf  <- run_CF_c_py(Y, T_info$T_sp, X, W, n_estimators = 2000)
+    te_info    <- get_te(te_hat_cf, test_data, x_vars)
 
-    te_info     <- get_te(te_hat_cf, test_data, x_vars)
+    # response & EONR
     T_seq      <- train_data[, seq(min(N), max(N), length.out = 200)]
@@
 
+    # true EONR
    true_eonr <- raw_sim_data$reg_data[[1]] %>%
+
   .[sim == test_sim_id, ] %>%
+
   .[, .(sim, data)] %>%
@@
 
-    saveRDS(combined, file = result_file)
+    saveRDS(combined, result_file)
+    combined
 
-    combined
   }, error = function(e) {
     message(sprintf("Error with sim %d: %s", test_sim_id, e$message))
-    return(data.table(sim = test_sim_id, aunit_id = integer(), opt_N_hat = numeric(), opt_N = numeric()))
+    data.table(
+      sim        = test_sim_id,
+      aunit_id   = integer(),
+      opt_N_hat  = numeric(),
+      opt_N      = numeric()
+    )
   })
 })
+
 ```

### 2025-07-27 codes/R/Add_y_tilde_into_raw_sim_data.Rmd
Add Y_tilde into raw_sim_data

# add y_tilde to raw_sim_data

# load raw_sim_data and data_2nd_stage
```{r}
raw_sim_data <- readRDS("/Users/mmousavi2/Dropbox/Local-Cross-Validation/Shared/Data/raw_sim_data.rds")

data_2nd_stage <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/data_2nd_stage.rds")

```

```{r}
# insert y_tilde into raw_sim_data
raw_sim_data$reg_data[[1]] <- raw_sim_data$reg_data[[1]] %>%
  mutate(
    data = pmap(
      list(data, sim),
      function(inner_df, sim_val) {
        inner_df %>%
          # temporarily add sim so we can join
          mutate(sim = sim_val) %>%
          left_join(
            data_2nd_stage %>% select(sim, aunit_id, y_tilde),
            by = c("sim", "aunit_id")
          ) %>%
          # remove the sim column—leave only the original columns + y_tilde
          select(-sim)
      }
    )
  )

```

# save it
```{r}
saveRDS(
  raw_sim_data,
  "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/raw_sim_data.rds"
)

```

### 2025-07-29 Writing/manuscript_CF_continuous.qmd
minor updates
Method significantly changed

--- 2025-07-15 Writing/manuscript_CF_continuous.qmd
@@
+---
+title: "Estimation of the effect of a continuous treatment variable"
+format:
+  html: default
+execute:
+  echo: false
+---
+
+```{r}
+#| include: false
+library(data.table)
+library(dplyr)
+library(flextable)
+library(officer)
+```
+
 # Introduction
 
-# Materials and Method
++ Background
++ Why we need the methods we are proposing? Motivation.
++ Summary of findings
+
+# Method
+
+We run Monte Carlo simulation to test the performance of ...
 
@@
 
-We employ three modeling approaches to estimate the yield response function to nitrogen application rates and to determine the site-specific optimal nitrogen rate.
-First, we implement artificial neural networks (ANNs) using two variations of the R-learner framework.
-In the first variation (RANN), both the outcome variable (yield) and the treatment variable (nitrogen rate) are orthogonalized.
-In the second variation, only the outcome variable is orthogonalized (half_RANN).
-Next, we apply an S-learner ANN, in which yield is modeled as a function of the nitrogen rate and a set of observed covariates (X), without orthogonalization.
-In addition to these ANN-based methods, we also use a causal forest (CF), implemented as a double machine learning approach following the R-learner framework.
-Finally, we include a random forest (RF) model using the R-learner approach with orthogonalization of the outcome variable only (half-RRF).
+We implement three estimation strategies—double‑orthogonalized, single‑orthogonalized, and non‑orthogonalized—that differ only in how (and whether) we orthogonalize the outcome and treatment before the final fit.
+Orthogonalization here means residualizing a variable by fitting it on all covariates and then taking the difference between the observed value and its prediction (we perform all residualizations using a random‑forest model).
+This ensures that our final model focuses on the component of variation that is not explained by the covariates.
 
-The econometric model to estimate the R-learner CF is as following:
+In the double‑orthogonalized approach, we first remove covariate effects from the outcome by predicting the outcome using all covariates and subtracting those predictions from the observed values.
+Next, we repeat this process for the treatment variable, yielding residuals that capture only the variation in treatment unexplained by covariates.
+The final learner is then trained on these two sets of residuals.
+By construction, the model estimates treatment effects using only the variation in outcome and treatment that is orthogonal to the covariates.
 
-$$
+The single‑orthogonalized method follows the same first step for the outcome but omits treatment residualization.
+After removing covariate effects from the outcome, we feed those outcome residuals—together with the original treatment values—into the final learner.
+This approach isolates covariate-driven patterns in the outcome while retaining the full treatment variation for effect estimation.
+
+In the non‑orthogonalized specification, we skip any residualization and directly fit the chosen learner to the original outcome and treatment (and covariates, where applicable).
+This allows us to assess the added value of orthogonalization.
+For each of these three strategies, we implement three different final‑stage learners—neural networks, random forests, and causal forests—to examine how both the orthogonalization scheme and the choice of algorithm influence our estimated treatment effects.
+
+The table below shows all the cases we consider.
+
+```{r}
+#| include: false
+f_est <-
+  data.table(
+    "Approach Label" = c("DO-NN", "SO-RFF", "SO-NN", "DO-CF","NO-NN", "NO-RF"),
+    "Approach Type" = c("Double-orthogonalized", "Single-orthogonalized", "Single-orthogonalized", "Double-orthogonalized", "Non-orthogonalized", "Non-orthogonalized"),
+    "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "Randome Forest", "Randome Forest", "NA", "NA"),
+    "T-orthogonalization" = c("Randome Forest", "NA", "NA", "Randome Forest", "NA", "NA"),
+    "Final stage" = c("Neural Network", "Random Forest", "Neural Network", "Causal Forest", "Neural Network", "Random Forest")
+  ) %>%
+  flextable() %>%
+  fontsize(size = 16, part = "all") %>%
+  autofit()
+
+flextable::save_as_image(f_est, here::here("writing/f_est.png"), res = 300)
+```
+
+```{r}
+#| echo: false
+#| fig-width: 4
+#| out-width: 4in
+knitr::include_graphics(here::here("writing/f_est.png"))
+```
+
+### Double-orthogonalized ANN (DO-ANN) and CF (DO-CF)
+
+#### DO‑ANN
+
+In the DO‑ANN approach, we estimate the yield response function
+
+```{=tex}
+\begin{align*}
+\hat{y}=\hat{f}(N: X)
+\end{align*}
+```
+
+using a structured neural network within a double‑machine‑learning (R‑learner) framework.
+Here $X$ represents observable soil and field characteristics that vary within a field in a spatially correlated manner and govern how yield responds to the nitrogen rate $N$.
+First, we orthogonalize both the outcome and the treatment basis variables by fitting each on all covariates $X$ (using cross‑fitted random forests) and taking residuals.
+For the treatment, we decompose the nitrogen rate into three basis functions and residualize each to obtain $\tilde{T}_1, \tilde{T}_2, \tilde{T}_3$.
+For the outcome, we similarly compute $Y-\hat{g}(X)$.
+Here $\hat{g}(X)$ denotes the predicted yield given covariates.
+The neural network itself has three parallel branches.
+Each branch processes the covariate vector $X$ through a feed‑forward network $f_k(X)$ and multiplies the output by the corresponding residualized treatment basis $\tilde{T}_k$.
+Summing these branch outputs yields the predicted yield
+
+```{=tex}
+\begin{align*}
+\hat{y}_i(N)=\sum_{k=1}^3 f_k\left(X_i\right) \tilde{T}_{k, i}
+\end{align*}
+```
+
+so that $\hat{f}$ maps any nitrogen rate $N$ (via its basis representation) and covariates $X$ to a yield estimate.
+Once trained, we obtain economically optimal nitrogen rates (EONR) by evaluating $\hat{f}$ on a grid of $N$ values for each observation $i$, computing the marginal profit
+
+```{=tex}
+\begin{align*}
+\operatorname{MP}\left(N ; X_i\right)=P_{\text {corn }} \times \hat{y}_i(N)-P_N \times N
+\end{align*}
+```
+
+and selecting the $N$ that maximizes profit function (where $P_{\text {corn }}$ is the market price of corn and $P_{\text {N }}$ is the per‑unit cost of nitrogen.
+This grid‑search procedure is applied separately to each field in the test set to generate site‑specific EONR estimates.
+
+#### DO‑CF (continuous)
+
+The DO‑CF specification follows the same double‑orthogonalization steps as DO‑ANN, but replace the neural network with a causal forest for the final estimation. 
+We specify the CF model
+
+```{=tex}
+\begin{align*}
 Y=\sum_{k=1}^K \theta_k(X) \cdot S_k(N)+g(X)+v
-$$
-where $Y$ is corn yield, $N$ is nitrogen rate (treatment variable), and $X$ is a collection of observed variables, $\alpha, \beta$, and $\gamma$. In this equation, $g(X)$ is the nuisance function.
-Here the treatment variable (nitrogen rate) is split into $K$ basis variables.
-Then, each of the basis variables are orthogonalized by cross-fitting using $X$ as covariates via boosted regression forest.
-Let the treatment basis variables be denoted as $S_k \tilde{(T)}$ for $k=1, \ldots, K$.
-Then, the second stage solves the following problem using the CF algorithm:
+\end{align*}
+```
 
-$$
-\operatorname{Min} \sum_{i=1}^N\left[{Y}_i-\sum_{k=1}^K \theta_k\left(X_i\right) \tilde{S}_k\left(N_i\right)\right]
-$$
+where 
+$Y$ is corn yield, $S_k(N)$ the $k$ th treatment basis function, and $g(X)$ the nuisance component.
+Following the double‑orthogonalization procedure, we first residualize $Y$ on $X$ by cross‑fitting a random forest regression to obtain $\tilde{Y}=Y-\widehat{g}(X)$.
+Next, we residualize each treatment basis $S_k(N)$ on $X$ in the same manner to produce $\tilde{S}_k=S_k(N)-\widehat{m}_k(X)$ for $k=1, \ldots, K$, where $\widehat{m}_k(X)$ is the cross-fitted prediction of the $k$ th basis function from covariates $X$.
+Let the orthogonalized outcome and treatment basis variables be denoted as $\tilde{Y}$ and $\tilde{S}_k(N)$ respectively.
+The final stage fits a causal forest to the residualized data by solving the following minimization problem:
 
-Once $\theta_k\left(X_i\right)$ are estimated, then the conditional treatment effect of moving from one rate $\left(N_0\right)$ to another $\left(N_1\right)$ for observation $i$ can be obtained by $\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_1\right)-\sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_0\right)$.
+```{=tex}
+\begin{align*}
+\operatorname{Min} \sum_{i=1}^N\left[\tilde{Y}_i-\sum_{k=1}^K \theta_k\left(X_i\right) \cdot \tilde{S}_k\left(N_i\right)\right]^2
+\end{align*}
+```
+Once the functions $\theta_k(X)$ are estimated, we recover the predicted yield response for any nitrogen rate $N$ and covariates $X$ as
 
-Profit impact of moving from $N_0$ to $N_1$ is
+```{=tex}
+\begin{align*}
+\hat{Y}(N: X)=\sum_{k=1}^K \hat{\theta}_k(X) \cdot S_k(N)
+\end{align*}
+```
 
-$$
-P_{\text {corn }} \sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_1\right)-P_{\text {corn }} \sum_{k=1}^K \theta_k(X) \tilde{S}_k\left(N_0\right)-P_N\left(N_1-N_0\right)
-$$
+As before, we perform a grid search over a sequence of $N$ values to identify the profit-maximizing N rate.
 
-A grid search over a sequence of $N$ values is used to identify the profit-maximizing N rate.
+### Single-orthogonalized NN (SO-NN) and (SO-RF)
 
-The S-learner ANN approach estimates the following model non-parametrically without imposing any functional forms on the relationships between the variables.
+In the single‑orthogonalized framework, only the outcome is residualized prior to the final-stage estimation, while the treatment variable remains in its original form.
+Specifically, we estimate the yield response function $\hat{Y}=\hat{f}(X, N)$.
+The first step involves orthogonalizing the outcome variable by cross-fitting a RF regression of $Y$ oN $X$ yielding residualized outcomes $\tilde{Y}$.
+In the second stage, the residualized outcome $\tilde{Y}$ is regressed on both $X$ and $N$ simultaneously, using a nonparametric learner.
+The treatment variable is not decomposed into basis functions, and no orthogonalization is applied to it.
+The difference between SO‑NN and SO‑RF lies in the choice of final learner: SO‑NN uses a neural network, while SO‑RF uses a random forest.
+In both cases, the model learns the mapping $\hat{f}(N: X)$ and is then used to predict yield responses over a grid of nitrogen rates.
+As in previous approaches, we compute the marginal profit for each candidate rate and select the nitrogen rate that maximizes this function.
 
-$$
-Y=f(X, N)+v
-$$
-Once $f(X, N)$ is estimated (denote it as $\hat{f}(X, N)$ ), then the conditional treatment effect of moving from one rate ( $N_0$ ) to another ( $N_1$ ) can be obtained by $\hat{f}\left(X, N_1\right)-\hat{f}\left(X, N_0\right)$.
-The associated change in profit is $P_{\text {corn }} \hat{f}\left(X, N_1\right)-P_{\text {corn }} \hat{f}\left(X, N_0\right)-P_N\left(N_1-N_0\right)$.
-Similar to the CF approach, a grid search is used to identify numerically the optimal nitrogen rate.
+### Non-orthogonalized ANN (NO-NN) and RF (NO-RF)
 
-The R-learner ANN (RANN) approach estimates the site-specific nitrogen response function by first orthogonalizing the outcome variable and then modeling the residuals using a neural network.
-Specifically, the outcome residual $y_i-\hat{g}\left(X_i\right)$ is regressed on orthogonalized treatment basis variables $\tilde{T}_1, \tilde{T}_2, \tilde{T}_3$, interacted with covariates $X$, using a structured artificial neural network.
-The model architecture includes three separate branches, each corresponding to one of the orthogonalized basis variables for the nitrogen rate.
-Each branch processes the covariates $X$ and multiplies the output by one of the treatment basis variables.
-The final yield prediction is the sum of the outputs from all three branches:
-
-$$
-\hat{Y}_i=f_1\left(X_i\right) \cdot \tilde{T}_{1 i}+f_2\left(X_i\right) \cdot \tilde{T}_{2 i}+f_3\left(X_i\right) \cdot \tilde{T}_{3 i}
-$$
-where $f_k(\cdot)$ are the outputs from each neural network branch, and $\tilde{T}_{k i}$ are the orthogonalized treatment basis variables for observation $i$.
-These basis variables are obtained by cross-fitting regression forests of each treatment basis on $X$, removing the component of the treatment that is predictable from the covariates.
-
-Once the model is trained, we use it to predict yield for a grid of nitrogen rates. These predictions are then used to compute the marginal profit function:
-
-$$
-\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
-$$
-The site specific nitrogen rate is the value of $N$ that maximizes this marginal profit function.
-This procedure is repeated for each observation in the test set to obtain field-specific EONR estimates.
-As with the other approaches, the profit-maximizing nitrogen rate is identified by conducting a grid search over a sequence of nitrogen values.
-
-The half R-learner artificial neural network (Half-RANN) estimates the site-specific nitrogen response function by orthogonalizing only the outcome variable while keeping the original (non-orthogonalized) nitrogen rate as the treatment input.
-Specifically, the orthogonalized outcome residual $\tilde{Y}_i=Y_i-\hat{g}\left(X_i\right)$, where $\hat{g}\left(X_i\right)$ is a nuisance function estimated from covariates $X$, is modeled using a neural network that takes the original treatment $N_i$ and the covariates $X_i$ as inputs.
-The model learns a function $f(X, N)$ using a fully connected neural network:
-
-$$
-\tilde{Y}_i=f\left(X_i, N_i\right)+\varepsilon_i
-$$
-Here, $X$ includes three observed variables that characterize the yield response function: the nitrogen level at which yield plateaus, the plateau yield level, and the intercept yield level at zero nitrogen.
-The nitrogen rate $N$ is included directly as one of the inputs to the network without basis transformation or orthogonalization.
-Once the network is trained, predicted yield responses $\hat{Y}(X, N)$ are obtained across a grid of nitrogen rates. These predictions are used to compute marginal profit:
-
-$$
-\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
-$$
-As with other models, a grid search over candidate $N$ values is used to identify the profit-maximizing rate.
-
-The half R-learner random forest (Half-RRF) approach estimates the site-specific yield response to nitrogen by orthogonalizing the outcome variable and modeling the resulting residuals using a random forest.
-In this approach, only the outcome $Y$ is orthogonalized with respect to observed covariates $X$, while the nitrogen treatment variable $N$ enters the model in its original form.
-The goal is to capture treatment effect heterogeneity while preserving flexibility in the treatment-response relationship.
-The model estimates the following equation:
-
-$$
-\tilde{Y}_i=f\left(X_i, N_i\right)+\varepsilon_i
-$$
-where $\tilde{Y}_i=Y_i-\hat{g}\left(X_i\right)$ is the residualized outcome, and $f(\cdot)$ is a random forest fitted to the nitrogen rate $N$ and covariates $X$.
-The model is trained using repeated K-fold cross-validation and hyperparameter tuning through grid search.
-Once trained, yield predictions $\hat{Y}(X, N)$ are generated for a grid of nitrogen rates.
-These predictions are used to compute the marginal profit function:
-
-$$
-\operatorname{MP}(N)=P_{\text {corn }} \cdot \hat{Y}(X, N)-P_N \cdot N
-$$
-This yields site-specific recommendations for nitrogen application.
+In the non‑orthogonalized setting, neither the outcome variable nor the treatment is residualized prior to estimation.
+Instead, the model directly fits the observed yield on both the nitrogen rate and covariates.
+We estimate the response function $\hat{Y}=\hat{f}(X, N)$ using the original values of $Y$, $N$ and $X$, without any orthogonalization.
+As in the single‑orthogonalized approaches, the function $\hat{f}(X, N)$ is estimated non‑parametrically.
+The difference between NO‑NN and NO‑RF lies in the choice of final-stage learner: NO‑NN uses a neural network, while NO‑RF relies on a random forest.
+Once the yield response function is estimated, the EONR is determined using the same profit-maximization procedure described in earlier models.
 

### 2025-07-30 codes/R/CF_Continuous_Loop.Rmd
commit changes before pull

@@
 # how many fields to train on?
-n_fields <- 5
+n_fields <- 1
 
@@
 x_vars    <- c("Nk", "plateau", "b0")
-T_formula <- yield ~ s(N, k = 4)
-
+#T_formula <- yield ~ s(N, k = 4)
+T_formula <- y_tilde ~ s(N, k = 4)
 # cf_results <- lapply(train_test_splits$test_id, function(test_sim_id)
@@
   tryCatch({
-
+    # grab exactly n_fields train IDs for this test
+    # base‐R subsetting – unambiguous
     split_row     <- train_test_splits[train_test_splits$test_id == test_sim_id, ]
@@
     T_info     <- prepare_T_mat(T_formula, train_data)
-    Y          <- train_data[, yield]
+    Y          <- train_data[, y_tilde]
     X          <- train_data[, ..x_vars] %>% as.matrix()
@@
    true_eonr <- raw_sim_data$reg_data[[1]] %>%
-
+  # 1) keep only this sim
   .[sim == test_sim_id, ] %>%
-
+  # 2) grab the nested data column
   .[, .(sim, data)] %>%
+  # 3) unnest into rows
   unnest(cols = c(data)) %>%
+  # 4) coerc e to data.table so you can use [ , … ]
   data.table() %>%
+  # 5) pull out just the columns you need
   .[, .(aunit_id, b0, b1, b2, Nk)] %>%
+  # 6) compute true opt_N
   .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
   .[, opt_N := pmin(Nk, pmax(0, opt_N))] %>%
+  # 7) keep only the ID + opt_N
   .[, .(aunit_id, opt_N)]
@@
 })
-
 ```

### 2025-07-30 Writing/manuscript_CF_continuous.qmd
comments added to the manuscript

@@
   echo: false
+bibliography: references.bib
 ---
@@
 ## Models and estimation procedures
+
+<span style = "color: blue;"> Use math here to explain orthogonalization.
+This is where you explain what SO and DO in general.
+The sections below add specifics to the individual approaches we take.
+You also need to provide contexts (past studies).
+</span>
+
+<span style = "color: blue;"> The model of interest is always the below no matter what approaches: </span>
+
+```{=tex}
+\begin{align*}
+y = f(X, N) + g(W) + v
+\end{align*}
+```
+
++ $y$: dependent variable
++ $X$: set of covariates that induce heterogeneity in the impact of $N$ on $y$
++ $W$: set of covariates that affect $y$, but not interactively with $N$
++ $v$: the error term
+
++ <span style = "color: blue;"> Start with non-orthogonalization approach.
+This is what is typically done in the literature (cite such papers).
+Talk about why this can be problematic like @kakimoto2022causal.
+They are not focused on causal impact identification, rather prediction of the level of the dependent variable.
+</span>
+
++ <span style = "color: blue;"> Shift the attention to DML in the binary treatment context (introduce the idea of orthogonalization and why they are done, cite DML papers) as a way to identify causal impact (causal machine learning).
+</span>
+
++ <span style = "color: blue;"> Our context is different because we have a continuous treatment variable.
+Introduce the idea of orthogonalization here and why SO can work but not DO because of interactions of the treatment variable with other covariates (heterogeneity-inducing variables).
+</span>
+
+<span style = "color: blue;"> This is the end of the general discussion of different approaches we use.  </span>
+
+<span style = "color: blue;"> Next talk about individual approaches just like you are doing right now.
+Switch DO and SO sections (Start with SO).
+Expand on the DO-CF approaches and why it works even though it is DO (due to linear-in-parameter nature of the spline-based approach).
+</span>.
 

### 2025-08-05 Writing/manuscript_CF_continuous.qmd
Significant changes on methods

--- 2025-07-29 Writing/manuscript_CF_continuous.qmd
@@
 
-We implement three estimation strategies—double‑orthogonalized, single‑orthogonalized, and non‑orthogonalized—that differ only in how (and whether) we orthogonalize the outcome and treatment before the final fit.
-Orthogonalization here means residualizing a variable by fitting it on all covariates and then taking the difference between the observed value and its prediction (we perform all residualizations using a random‑forest model).
-This ensures that our final model focuses on the component of variation that is not explained by the covariates.
+<span style = "color: blue;"> Use math here to explain orthogonalization.
+This is where you explain what SO and DO in general.
+The sections below add specifics to the individual approaches we take.
+You also need to provide contexts (past studies).
+</span>
 
-In the double‑orthogonalized approach, we first remove covariate effects from the outcome by predicting the outcome using all covariates and subtracting those predictions from the observed values.
-Next, we repeat this process for the treatment variable, yielding residuals that capture only the variation in treatment unexplained by covariates.
-The final learner is then trained on these two sets of residuals.
-By construction, the model estimates treatment effects using only the variation in outcome and treatment that is orthogonal to the covariates.
+The standard approach in the precision-agriculture literature for estimating site-specific input response is to model the outcome level directly as a function of the treatment/input and covariates—e.g., estimate $Y=f(I,C)$ where $I$ is the input (like nitrogen rate) and $C$ are field characteristics—and then derive optimal input decisions from that estimated function.
+This is what is typically done in the recent on-farm precision experimentation literature (e.g., @barbosa2020risk; @barbosa2020modeling; @krause2020random; @gardner2021economic).
+The problem is that these methods optimize prediction of yield levels—not the causal effect of changing the input.
+Good level prediction at the observed input does not guarantee accurate recovery of the response function needed for optimal input choice, because the fitted $f(I,C)$ can conflate input variation with confounding and misattribute the drivers of yield variation.
 
-The single‑orthogonalized method follows the same first step for the outcome but omits treatment residualization.
-After removing covariate effects from the outcome, we feed those outcome residuals—together with the original treatment values—into the final learner.
-This approach isolates covariate-driven patterns in the outcome while retaining the full treatment variation for effect estimation.
+To identify causal effects rather than just predict levels in the single binary treatment case, researchers use double/debiased machine learning (DML), which introduces orthogonalization so the treatment-effect estimator is insensitive to small errors in the auxiliary fits (I need citation here).
+Consider the structural model (Model 1):
 
-In the non‑orthogonalized specification, we skip any residualization and directly fit the chosen learner to the original outcome and treatment (and covariates, where applicable).
-This allows us to assess the added value of orthogonalization.
-For each of these three strategies, we implement three different final‑stage learners—neural networks, random forests, and causal forests—to examine how both the orthogonalization scheme and the choice of algorithm influence our estimated treatment effects.
+```{=tex}
+\begin{align*}
+Y=\theta(X) T+g(X, W)+v
+\end{align*}
+```
+
+where $Y$ is the outcome, $T \in\{0,1\}$  is the treatment, $X$ covariates that also modify the treatment effect $\theta(X)$, $W$ additional predictors of $Y$, and $v$ is noise. 
+Orthogonalization estimates $f(X, W)=\mathbb{E}[Y \mid X, W]$ and $h(X, W)=\mathbb{E}[T \mid X, W]$, forms residuals $Y_{\mathrm{res}}=Y-f(X,W)$ and $T_{\mathrm{res}}=T-h(X, W)$, and then relates $Y_{\mathrm{res}}$ to $T_{\mathrm{res}}$.
+Because this score is Neyman-orthogonal and is typically implemented with cross-fitting, small errors in $f$ and $h$ do not induce first-order bias in $\theta(X)$.
+
+The identifying algebra in Model 1 follows from linearity in $T$:
+
+```{=tex}
+\begin{align*}
+\mathbb{E}[Y \mid X, W]=\theta(X) \mathbb{E}[T \mid X, W]+g(X, W)=\theta(X) h(X, W)+g(X, W)
+\end{align*}
+```
+
+so
+
+```{=tex}
+\begin{align*}
+Y_{\mathrm{res}}=Y-\mathbb{E}[Y \mid X, W]=\theta(X)(T-h(X, W))+v=\theta(X) T_{\mathrm{res}}+v
+\end{align*}
+```
+
+Thus regressing $Y_{\mathrm{res}}$ on $T_{\mathrm{res}}$ recovers $\theta(X)$ up to noise.
+
+In the on farm precision experiment (OFPE) setting, the ideas behind DML have shown promise for causal effect estimation.
+@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield, but their implementation was limited to a single-year experiment from one field and treated heterogeneity in a discrete way.
+They defined multiple “treatments” as increases from the lowest nitrogen rate to each higher rate, which is natural when there are only five or six fixed target rates per field.
+When data are pooled across fields with differing targets, however, applied nitrogen rates become effectively continuous, and the standard residualization of the treatment in DML does not extend cleanly to that setting.
+
+Consider the continuous-treatment setting, where the structural model is
+
+```{=tex}
+\begin{align*}
+Y=\theta(X, T)+g(X, W)+v
+\end{align*}
+```
+
+with $T$ (e.g., nitrogen rate) continuous, and $X$ heterogeneity-inducing covariates that interact with $T$ through the flexible response function $\theta(X, T)$.
+In this case double-orthogonalization fails because $\theta$ is a general (nonlinear and interaction-rich) function of $T$.
+Residualizing the treatment by replacing $T$ with $T-\mathbb{E}[T \mid X, W]$ does not produce a valid adjustment for the heterogeneous effect.
+Formally,
+
+```{=tex}
+\begin{align*}
+\theta(X, T)-\mathbb{E}[\theta(X, T) \mid X, W] \neq \theta(X, T-\mathbb{E}[T \mid X, W])
+\end{align*}
+```
+
+and the clean decomposition that underlies double-orthogonalization in the linear-in-$T$ case breaks down. 
+Instead, we apply single-orthogonalization (SO).
+We first estimate $f(X, W)=\mathbb{E}[Y \mid X, W]$ and form the residual $Y_{\mathrm{res}}=Y-f(X, W)$, and then recover $\theta(X, T)$ by flexibly modeling  $Y_{\mathrm{res}}$ as a function of $(X,T)$.
+This approach removes baseline variation due to $g(X,W)$ while preserving the full interaction structure between $X$ and the continuous treatment.
+
+To study and compare these ideas systematically, we adopt a single underlying structural model and then implement alternative orthogonalization schemes.
+
+The model of interest throughout is (Model 3)
+
+```{=tex}
+\begin{align*}
+y = f(X, N) + g(W) + v
+\end{align*}
+```
+
+where $y$ is crop yield, $N$ is the continuous nitrogen rate, $X$ is a set of covariates that induce heterogeneity in the impact of $N$ on $y$, and $W$ is a set of covariates that affect $y$ but do not interact with $N$.
+We implement three estimation strategies—double-orthogonalized, single-orthogonalized, and non-orthogonalized—that differ only in how (and whether) the outcome and treatment are residualized before the final fit.
+Orthogonalization means fitting a variable on its conditioning covariates and using the residual (observed minus predicted) so that the subsequent learner focuses on variation not explained by those covariates; all residualizations are performed with random forests.
+In the double-orthogonalized approach, both $y$ and $N$ are residualized with respect to $(X,W)$ and the final learner is trained on those residuals.
+In the single-orthogonalized approach, only $y$ is residualized and the raw $N$ is retained.
+The non-orthogonalized specification skips residualization entirely and fits the learner directly to $(y,N,X,W)$.
+For each strategy we consider three final-stage learners—neural networks (NN), random forests (RF), and causal forests (CF)—to evaluate how the orthogonalization scheme and learning algorithm interact in recovering heterogeneous nitrogen response.
 
@@
 
-### Double-orthogonalized ANN (DO-ANN) and CF (DO-CF)
+### Single-orthogonalized NN (SO-NN) and (SO-RF)
+
+In the single-orthogonalized framework only the baseline component $g(W)$ is partialed out before the final-stage estimation; the continuous treatment $N$ and heterogeneity-inducing covariates $X$ remain in their original form.
+Formally, starting from Model 3, we first estimate $g(W)=\mathbb{E}[y \mid W]$ using a cross-fitted random forest and form the residualized outcome
+
+```{=tex}
+\begin{align*}
+\tilde{y}=y-\widehat{g}(W)
+\end{align*}
+```
+
+The second stage fits a flexible model of $\tilde{y}$ on $(X,N)$ to recover the heterogeneous response function $f(X,N)$.
+The distinction between SO-NN and SO-RF is only in the choice of the final learner: SO-NN uses a neural network, while SO-RF uses a random forest.
+In both cases the learner estimates $\hat{f}(X, N)$, which is then used to predict yield under candidate nitrogen rates.
+Marginal profit is computed over a grid of $N$ values and the economically optimal nitrogen rate is selected as the one that maximizes
+
+```{=tex}
+\begin{align*}
+\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
+\end{align*}
+```
+where $P_{\text {Corn }}$ is the corn price, $P_{\text {N }}$ is the nitrogen price, and $\hat{f}$ is the estimated yield response.
+
+### Double-orthogonalized CF (DO-CF) and ANN (DO-ANN)
+
+#### DO‑CF (continuous)
+
+To adapt double-orthogonalization to the model of interest (Model 3), we estimate the heterogeneous response $f(X,N)$ by expanding the continuous treatment $N$ in a set of spline basis functions and allowing their coefficients to vary with the heterogeneity-inducing covariates $X$:
+
+```{=tex}
+\begin{align*}
+y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
+\end{align*}
+```
+
+Although double-orthogonalization generally fails when $f(X,N)$ is an arbitrary nonlinear function of $N$, this spline-based formulation restores the key linear-in-parameter structure that makes DO valid, because the model is linear in the transformed treatment basis $\left\{S_k(N)\right\}$.
+Residualizing each basis component and the outcome with respect to $(X,W)$ then yields a clean decomposition analogous to the linear-in-$T$ case.
+Let $\widetilde{S}_k(N)=S_k(N)-\mathbb{E}\left[S_k(N) \mid X, W\right]$.
+Then
+
+```{=tex}
+\begin{align*}
+\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
+\end{align*}
+```
+
+so causal forest can nonparametrically estimate each heterogeneous weight $\theta_k(X)$  by relating the residualized outcome to the residualized spline bases.
+This Neyman-orthogonal (R-learner) style decomposition preserves the ability to capture complex, heterogeneous, and nonlinear treatment effects while inheriting the robustness of double-orthogonalization: small errors in the nuisance fits for $g(X,W)$ or the conditional means of the bases do not induce first-order bias in $\theta_k(X)$.
+The estimated $\theta_k(X)$ then combine to recover the full heterogeneous response function and, by differencing across candidate $N$ values, site-specific treatment effects and economically optimal nitrogen rates.
 
@@
 
-In the DO‑ANN approach, we estimate the yield response function
+The DO-ANN follows the same double-orthogonalized, linear-in-basis logic as DO-CF, adapted to the Model 3.
+Starting from the residualized outcome $\tilde{y}$, the continuous treatment $N$ is expanded into spline basis functions and each basis is orthogonalized with respect to $(X,W)$; these orthogonalized spline bases are $T_1, T_2, T_3$.
+The model assumes
 
@@
 \begin{align*}
-\hat{y}=\hat{f}(N: X)
+\tilde{y} \approx \theta_1(X) T_1+\theta_2(X) T_2+\theta_3(X) T_3,
 \end{align*}
@@
 
-using a structured neural network within a double‑machine‑learning (R‑learner) framework.
-Here $X$ represents observable soil and field characteristics that vary within a field in a spatially correlated manner and govern how yield responds to the nitrogen rate $N$.
-First, we orthogonalize both the outcome and the treatment basis variables by fitting each on all covariates $X$ (using cross‑fitted random forests) and taking residuals.
-For the treatment, we decompose the nitrogen rate into three basis functions and residualize each to obtain $\tilde{T}_1, \tilde{T}_2, \tilde{T}_3$.
-For the outcome, we similarly compute $Y-\hat{g}(X)$.
-Here $\hat{g}(X)$ denotes the predicted yield given covariates.
-The neural network itself has three parallel branches.
-Each branch processes the covariate vector $X$ through a feed‑forward network $f_k(X)$ and multiplies the output by the corresponding residualized treatment basis $\tilde{T}_k$.
-Summing these branch outputs yields the predicted yield
-
-```{=tex}
-\begin{align*}
-\hat{y}_i(N)=\sum_{k=1}^3 f_k\left(X_i\right) \tilde{T}_{k, i}
-\end{align*}
-```
-
-so that $\hat{f}$ maps any nitrogen rate $N$ (via its basis representation) and covariates $X$ to a yield estimate.
-Once trained, we obtain economically optimal nitrogen rates (EONR) by evaluating $\hat{f}$ on a grid of $N$ values for each observation $i$, computing the marginal profit
-
-```{=tex}
-\begin{align*}
-\operatorname{MP}\left(N ; X_i\right)=P_{\text {corn }} \times \hat{y}_i(N)-P_N \times N
-\end{align*}
-```
-
-and selecting the $N$ that maximizes profit function (where $P_{\text {corn }}$ is the market price of corn and $P_{\text {N }}$ is the per‑unit cost of nitrogen.
-This grid‑search procedure is applied separately to each field in the test set to generate site‑specific EONR estimates.
-
-#### DO‑CF (continuous)
-
-The DO‑CF specification follows the same double‑orthogonalization steps as DO‑ANN, but replace the neural network with a causal forest for the final estimation. 
-We specify the CF model
-
-```{=tex}
-\begin{align*}
-Y=\sum_{k=1}^K \theta_k(X) \cdot S_k(N)+g(X)+v
-\end{align*}
-```
-
-where 
-$Y$ is corn yield, $S_k(N)$ the $k$ th treatment basis function, and $g(X)$ the nuisance component.
-Following the double‑orthogonalization procedure, we first residualize $Y$ on $X$ by cross‑fitting a random forest regression to obtain $\tilde{Y}=Y-\widehat{g}(X)$.
-Next, we residualize each treatment basis $S_k(N)$ on $X$ in the same manner to produce $\tilde{S}_k=S_k(N)-\widehat{m}_k(X)$ for $k=1, \ldots, K$, where $\widehat{m}_k(X)$ is the cross-fitted prediction of the $k$ th basis function from covariates $X$.
-Let the orthogonalized outcome and treatment basis variables be denoted as $\tilde{Y}$ and $\tilde{S}_k(N)$ respectively.
-The final stage fits a causal forest to the residualized data by solving the following minimization problem:
-
-```{=tex}
-\begin{align*}
-\operatorname{Min} \sum_{i=1}^N\left[\tilde{Y}_i-\sum_{k=1}^K \theta_k\left(X_i\right) \cdot \tilde{S}_k\left(N_i\right)\right]^2
-\end{align*}
-```
-Once the functions $\theta_k(X)$ are estimated, we recover the predicted yield response for any nitrogen rate $N$ and covariates $X$ as
-
-```{=tex}
-\begin{align*}
-\hat{Y}(N: X)=\sum_{k=1}^K \hat{\theta}_k(X) \cdot S_k(N)
-\end{align*}
-```
-
-As before, we perform a grid search over a sequence of $N$ values to identify the profit-maximizing N rate.
-
-### Single-orthogonalized NN (SO-NN) and (SO-RF)
-
-In the single‑orthogonalized framework, only the outcome is residualized prior to the final-stage estimation, while the treatment variable remains in its original form.
-Specifically, we estimate the yield response function $\hat{Y}=\hat{f}(X, N)$.
-The first step involves orthogonalizing the outcome variable by cross-fitting a RF regression of $Y$ oN $X$ yielding residualized outcomes $\tilde{Y}$.
-In the second stage, the residualized outcome $\tilde{Y}$ is regressed on both $X$ and $N$ simultaneously, using a nonparametric learner.
-The treatment variable is not decomposed into basis functions, and no orthogonalization is applied to it.
-The difference between SO‑NN and SO‑RF lies in the choice of final learner: SO‑NN uses a neural network, while SO‑RF uses a random forest.
-In both cases, the model learns the mapping $\hat{f}(N: X)$ and is then used to predict yield responses over a grid of nitrogen rates.
-As in previous approaches, we compute the marginal profit for each candidate rate and select the nitrogen rate that maximizes this function.
+so that each $\theta_k(X)$ is the heterogeneous weight on the corresponding orthogonalized spline component of $N$.
+The key difference from DO-CF is the estimator for the coefficient functions $\theta_k(X)$: in DO-CF these are learned nonparametrically via causal forests, whereas in DO-ANN they are approximated by three separate neural-network branches, each producing a scalar weight that multiplies one of the $T_k$; the branch outputs are summed to produce the predicted residual.
+Because the decomposition is linear in the orthogonalized spline bases, the same clean double-orthogonalization justification applies, and the learned branch outputs recover the heterogeneous response without introducing first-order bias from the nuisance fits.
+The combined output reconstructs $\hat{f}(X, N)$, which is then used in the economic objective to compute the EONR.
 

### 2025-08-08 codes/R/Add_Ortho_N_to_data_2nd_stage.qmd
Add N_tilde to the data_2nd_stage

---
title: "Add N_tilde to the data_2nd_stage (<span style='color:blue'>Score: /10 </span>)"
author: "Your Name Here"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

# Load data
```{r setup_not_remove, echo=FALSE, include = F}

data_2nd_stage <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/data_2nd_stage.rds")
evall_N_seq <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/evall_N_seq.rds")

```

# Fit a RF
```{r setup_not_remove, echo=FALSE, include = F}

library(randomForest)

# fitting a RF to predict N from Nk, plateau, and b0

# rf_mod <- randomForest(
#   N ~ Nk + plateau + b0,
#   data = data_2nd_stage,
#   ntree = 100
# )

if (!requireNamespace("ranger", quietly = TRUE)) install.packages("ranger")
library(ranger)

rf_mod <- ranger(
  formula = N ~ Nk + plateau + b0,
  data    = data_2nd_stage,
  num.trees     = 100,       # or fewer if you like
  #mtry          = 2,         # smaller default than sqrt(p)
  min.node.size = 5,         # larger to speed up
  num.threads   = parallel::detectCores()
)

```

# add N_tilde to the data
```{r setup_not_remove, echo=FALSE, include = F}

# # Compute the predicted (fitted) values
# fitted_N <- predict(rf_mod, newdata = data_2nd_stage)
# 
# # Add the residualized treatment to the data 
# data_2nd_stage$N_tilde <- data_2nd_stage$N - fitted_N

data_2nd_stage$fitted_N <- predict(rf_mod, data = data_2nd_stage)$predictions
data_2nd_stage$N_tilde  <- data_2nd_stage$N - data_2nd_stage$fitted_N

```

# save data_2nd_stage with N_tilde
```{r setup_not_remove, echo=FALSE, include = F}
output_dir <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

saveRDS(data_2nd_stage,
        file = file.path(output_dir, "data_2nd_stage.rds"))

```

# add N_tilde seq to the eval_N_seq data
```{r setup_not_remove, echo=FALSE, include = F}
library(dplyr)

# Step 1: Create a sequence of 100 N_tilde values per sim
ntilde_seq <- data_2nd_stage %>%
  group_by(sim) %>%
  summarise(N_tilde = list(seq(min(N_tilde), max(N_tilde), length.out = 100))) %>%
  unnest(N_tilde)

# Step 2: Bind the new N_tilde column to evall_N_seq by sim
evall_N_seq <- evall_N_seq %>%
  arrange(sim) %>%
  group_by(sim) %>%
  mutate(row = row_number()) %>%
  ungroup()

ntilde_seq <- ntilde_seq %>%
  group_by(sim) %>%
  mutate(row = row_number()) %>%
  ungroup()

evall_N_seq <- evall_N_seq %>%
  left_join(ntilde_seq, by = c("sim", "row")) %>%
  select(-row)

```

# save eval_N_seq data with N_Tilde seq

```{r setup_not_remove, echo=FALSE, include = F}
output_dir <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

saveRDS(evall_N_seq,
        file = file.path(output_dir, "evall_N_seq.rds"))

```

### 2025-08-10 codes/Python/Functions/RANN.py
Changing RANN to N_tilde

@@ -13,17 +13,33 @@ from sklearn.preprocessing import StandardScaler
 # %%
 # === MODEL 1: RANN ===
 def train_rann_model_from_splits(data_2nd_stage, evall_N_seq, split_csv_path, folder_name, device):
+    #class MyModel(nn.Module):
+        #def __init__(self):
+            #super(MyModel, self).__init__()
+            #self.branch1 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
+            #self.branch2 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
+            #self.branch3 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
+
+        #def forward(self, inp):
+            #x = inp[:, :3]
+            #T = inp[:, 3:]
+            #return self.branch1(x)*T[:,0:1] + self.branch2(x)*T[:,1:2] + self.branch3(x)*T[:,2:3]
     class MyModel(nn.Module):
         def __init__(self):
             super(MyModel, self).__init__()
-            self.branch1 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
-            self.branch2 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
-            self.branch3 = nn.Sequential(nn.Linear(3, 64), nn.ReLU(), nn.Linear(64, 64), nn.ReLU(), nn.Linear(64, 1))
+            # single branch for the 3 covariates
+            self.branch = nn.Sequential(
+                nn.Linear(3, 64),
+                nn.ReLU(),
+                nn.Linear(64, 64),
+                nn.ReLU(),
+                nn.Linear(64, 1)
+            )
 
         def forward(self, inp):
-            x = inp[:, :3]
-            T = inp[:, 3:]
-            return self.branch1(x)*T[:,0:1] + self.branch2(x)*T[:,1:2] + self.branch3(x)*T[:,2:3]
+           x = inp[:, :3]     # Nk, plateau, b0
+           n = inp[:, 3:4]    # N_tilde
+           return self.branch(x) * n        
 
[…]

### 2025-08-10 codes/R/debugging.Rmd
Check CF simulation to make sure there is no mistakes

# Objectives

+ Test whether python codes for estimating R-learner models are working properly using a dataset following a simple data generating process

+ Test whether spline approach (based on the `mgcv` package) is working properly

# Preparation

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
#library(here)
#library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

library(reticulate)
use_condaenv("cf_conda311", required = TRUE)
py_config()

py_run_string("
import platform, numpy as np, sklearn, scipy, econml
print('arch:', platform.machine())
print('numpy:', np.__version__)
print('sklearn:', sklearn.__version__)
print('scipy:', scipy.__version__)
print('econml:', econml.__version__)
")

# make sure Python can see the folder that contains run_CF_c.py
py_run_string("import sys, os; sys.path.append('/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python')")

# clear any stale bytecode that might have been written under a different toolchain
system("find /Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python -name '__pycache__' -type d -prune -exec rm -rf {} +")

# now import without executing R helpers:
py_run_string("import run_CF_c; print('run_CF_c imported OK; has run_CF_c_py:', hasattr(run_CF_c, 'run_CF_c_py'))")

#--- python functions ---#
# use_condaenv("cf_conda311", required = TRUE)
# py_config()
# py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
source_python("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
```

## Data

```{r}
all_data <- raw_sim_data

test_sim_id=1

split_row     <- train_test_splits_10fields[train_test_splits_10fields$test_id == test_sim_id, ]
train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:10)]))

sim_id_ls     <- c(train_sim_ids, test_sim_id)

    # assemble train & test data
    data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% sim_id_ls, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    train_data <- data[sim != test_sim_id, ]

    # aunit level
    test_data  <- raw_sim_data$reg_data[[1]] %>%
      .[sim == test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    # cell level
  # test_data <-
  # all_data$field_pars[[1]] %>%
  # .[sim == test_sim_id, ]

```

# Train CF

```{r, eval = FALSE}
x_vars <- c("Nk", "plateau", "b0")
# x_vars <-
#   c(
#     "b0_1", "b0_2",
#     "b1_2_1", "b1_2_2", "b1_1_1", "b1_1_2",
#     # "b2_2_1", "b2_2_2", "b2_1_1", "b2_1_2",
#     "Nk_2_1", "Nk_2_2", "Nk_1_1", "Nk_1_2",
#     "plateau_2_1", "plateau_2_2", "plateau_1_1", "plateau_1_2"
#     # "theta_plateau_2", "theta_Nk_2", "theta_b0_2"
#   )

T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
Y <- train_data[, yield]
X <- train_data[, ..x_vars] %>% as.matrix()
W <- X
X_test <- test_data[, ..x_vars] %>% as.matrix()

te_hat_cf <- run_CF_c_py(Y, T_info$T_sp, X, W, n_estimators = 2000)

te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")
#te_info <- get_te(te_hat_cf, test_data, x_vars, "cell_id")

T_seq <- train_data[, seq(min(N), max(N), length = 200)]

response_data <- find_response_semi(T_seq, T_info, te_info)

saveRDS(list(te_hat_cf = te_hat_cf, te_info = te_info, T_seq = T_seq, response_data = response_data, test_data = test_data, train_data = train_data), "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/results/Test_CF_any_mis/cf-continuous-unit_10_fields_double_check.rds")
```

# Post-estimation analysis
```{r}
cf_results <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/results/Test_CF_any_mis/cf-continuous-unit_10_fields_double_check.rds")
```

## Site-specific EONR

```{r}
pCorn <- 6.25 / 25.4 # $/kg
pN <- 1 / 0.453592 # $/kg

# aunit id ###############################
ss_eonr <-
  cf_results$response_data %>%
  .[, profit := est * pCorn - pN * T] %>%
  .[, .SD[which.max(profit), ], by = aunit_id] %>%
  .[, .(aunit_id, T)] %>%
  setnames("T", "opt_N_hat")

true_ss_eonr <-
  cf_results$test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
  .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
  .[, opt_N := pmin(Nk, opt_N)] %>%
  .[, opt_N := pmax(0, opt_N)] %>%
  .[, .(aunit_id, opt_N)]

combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]

combined_eonr[, cor(opt_N, opt_N_hat)]

############################################################################

# cell ###########################
# ss_eonr <-
#   cf_results$response_data %>%
#   .[, profit := est * pCorn - pN * T] %>%
#   .[, .SD[which.max(profit), ], by = cell_id] %>%
#   .[, .(cell_id, T)] %>%
#   setnames("T", "opt_N_hat")
#  
#  true_ss_eonr <-
#   cf_results$test_data[, .(cell_id, b0, b1, b2, Nk)] %>%
#   .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
#   .[, opt_N := pmin(Nk, opt_N)] %>%
#   .[, opt_N := pmax(0, opt_N)] %>%
#   .[, .(cell_id, opt_N)]
#  
# combined_eonr <- ss_eonr[true_ss_eonr, on = "cell_id"]

#combined_eonr[, cor(opt_N, opt_N_hat)]

#######################################################################################

rmse_10_fields_unit <- sqrt(mean((combined_eonr$opt_N_hat - combined_eonr$opt_N)^2, na.rm = TRUE))  
rmse_10_fields_unit
```

# dependency functions
```{r}
prepare_T_mat <- function(gam_formula, data) {
  gam_setup <- gam(gam_formula, data = data)
  T_var_name <- all.vars(gam_formula)[2]
  T_sp <- predict(gam_setup, data = data, type = "lpmatrix") %>%
    .[, -1] %>%
    data.table() %>%
    setnames(names(.), paste0("T_", 1:ncol(.)))
  return(list(gam_setup = gam_setup, T_sp = T_sp, T_var_name = T_var_name))
}

################################################################################
# aunit

get_te <- function(trained_model, test_data, x_vars, id_var = "aunit_id") {
  X_test <- test_data[, ..x_vars] %>% as.matrix()
  te_hat <- trained_model$const_marginal_effect(X_test)
  return(list(te_hat = data.table(te_hat), id_data = test_data[, ..id_var]))
}

# cell

# get_te <- function(trained_model, test_data, x_vars, id_var = "cell_id") {
#   X_test <- test_data[, ..x_vars] %>% as.matrix()
#   te_hat <- trained_model$const_marginal_effect(X_test)
#   return(list(te_hat = data.table(te_hat), id_data = test_data[, ..id_var]))
# }
################################################################################
find_response_semi <- function(T_seq, T_info, te_info) {
  eval_T <- data.table(T = T_seq) %>%
    setnames("T", T_info$T_var_name) %>%
    predict(T_info$gam_setup, newdata = ., type = "lpmatrix") %>%
    .[, -1] %>%
    data.table()

  curv_data <- as.matrix(te_info$te_hat) %*% t(as.matrix(eval_T)) %>%
    data.table() %>%
    setnames(names(.), as.character(T_seq)) %>%
    .[, id := 1:.N] %>%
    melt(id.var = "id") %>%
    setnames(c("variable", "value"), c("T", "est")) %>%
    .[, T := as.numeric(as.character(T))]

  id_data <- te_info$id_data[, id := 1:.N]
  final_data <- id_data[curv_data, on = "id"][, id := NULL]
  return(final_data)
}

```

### 2025-08-12 CF_loop.Rmd
CF loop, still needs attention to make it organazied

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
# library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

#--- R functions ---#
# here("codes/R") %>%
#   fs::dir_ls(., full.names = TRUE) %>%
#   lapply(source)

use_virtualenv("cf_env311", required = TRUE)
Sys.setenv(DYLD_LIBRARY_PATH = "/opt/homebrew/Cellar/llvm@15/15.0.7/lib/c++")
py_config()

source_python("/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")

```

# CF

```{r}

## ───────────────────────────────────────
train_test_splits <- train_test_splits_1fields
n_fields <- 1  # how many train sims

## ── Output paths ──────────────────────────
output_dir   <- "/Users/monamousavi/Dropbox/CF_res_1"
dir.create(path.expand(output_dir), recursive = TRUE, showWarnings = FALSE)
progress_csv <- file.path(output_dir, "progress.csv")

## ─  keep an in-memory list ────────────
#cf_all_results <- vector("list", length = nrow(train_test_splits))
#names(cf_all_results) <- train_test_splits$test_id
train_test_splits <- subset(train_test_splits, !test_id %in% as.integer(sub("^sim_(\\d+)\\.rds$","\\1", basename(list.files(path.expand(output_dir), "^sim_\\d+\\.rds$")))))

## ── Loop over test sims ───────────────────────────────────────────────────────
for (i in seq_len(nrow(train_test_splits))) {

  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  # 
  tryCatch({

    ## 1) Train/test IDs
    split_row     <- train_test_splits[i, ]
    train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
    sim_id_ls     <- c(train_sim_ids, test_sim_id)

    ## 2) train & test data
    data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% sim_id_ls, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    train_data <- data[sim != test_sim_id, ]
    test_data  <- raw_sim_data$reg_data[[1]] %>%
      .[sim == test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    ## 3)  CF
    x_vars <- c("Nk", "plateau", "b0")
    T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
    Y      <- train_data[, yield]
    X      <- as.matrix(train_data[, ..x_vars])
    W      <- X
    X_test <- as.matrix(test_data[, ..x_vars])

    ## 4) Run CF in Python
    Sys.setenv(
      OMP_NUM_THREADS        = "1",
      MKL_NUM_THREADS        = "1",
      KMP_AFFINITY           = "disabled",
      KMP_DUPLICATE_LIB_OK   = "TRUE"
    )

    te_hat_cf <- run_CF_c_py(
      Y,
      T_info$T_sp,
      X,
      W,
      n_estimators = 2000
    )
    te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

    ## 5) Response curves
    T_seq <- train_data[, seq(min(N), max(N), length = 200)]
    response_data <- find_response_semi(T_seq, T_info, te_info)

    ## 6) Site-specific EONR
    pCorn <- 6.25 / 25.4
    pN    <- 1 / 0.453592

    ss_eonr <- response_data %>%
      .[, profit := est * pCorn - pN * T] %>%
      .[, .SD[which.max(profit), ], by = aunit_id] %>%
      .[, .(aunit_id, opt_N_hat = T)]

    true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
      .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
      .[, opt_N := pmin(Nk, opt_N)] %>%
      .[, opt_N := pmax(0, opt_N)] %>%
      .[, .(aunit_id, opt_N)]

    combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
    corr_val      <- combined_eonr[, cor(opt_N, opt_N_hat)]

    ## 7) Package result for THIS sim
    res <- list(
      te_hat_cf     = te_hat_cf,
      te_info       = te_info,
      T_seq         = T_seq,
      response_data = response_data,
      test_data     = test_data,
      train_data    = train_data,
      combined_eonr = combined_eonr,
      corr          = corr_val
    )

    #  keep in memory too
    #cf_all_results[[as.character(test_sim_id)]] <- res

    ## 8) SAVE THIS SIM IMMEDIATELY 
    tmp_fn <- paste0(out_fn, ".tmp")
    saveRDS(res, tmp_fn, compress = "xz")
    file.rename(tmp_fn, out_fn)

    ## 9) Log progress
    prog <- data.frame(
      test_sim_id = test_sim_id,
      corr        = corr_val,
      n_train     = nrow(train_data),
      saved_file  = out_fn,
      ts          = as.character(Sys.time())
    )
    write.table(
      prog,
      file      = progress_csv,
      sep       = ",",
      row.names = FALSE,
      col.names = !file.exists(progress_csv),
      append    = TRUE
    )

    message("Finished sim ", test_sim_id, " → ", basename(out_fn),
            " (corr = ", round(corr_val, 3), ")")

    # Free RAM periodically
    rm(res, te_hat_cf, te_info, response_data, train_data, test_data)
    if (i %% 10 == 0) gc()

  }, error = function(e) {
    message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
    #  which sim failed
    err_row <- data.frame(
      test_sim_id = test_sim_id,
      error       = conditionMessage(e),
      ts          = as.character(Sys.time())
    )
    write.table(
      err_row,
      file      = file.path(output_dir, "errors.csv"),
      sep       = ",",
      row.names = FALSE,
      col.names = !file.exists(file.path(output_dir, "errors.csv")),
      append    = TRUE
    )
  })

}

## ── save the whole list at the end ─────────────────────────────────
# saveRDS(cf_all_results, file.path(output_dir, "_all_results.rds"))

```
# read in

```{r}
library(data.table)

in_dir   <- "/Users/monamousavi/Dropbox/CF_res_1"         # where current sim_###.rds live
out_dir  <- "~/Desktop/CF_res_1_eonr"    # new lean outputs
dir.create(path.expand(out_dir), recursive = TRUE, showWarnings = FALSE)

files  <- list.files(path.expand(in_dir), pattern = "^sim_\\d+\\.rds$", full.names = TRUE)
sim_id <- as.integer(sub("^.*sim_(\\d+)\\.rds$", "\\1", files))

for (i in seq_along(files)) {
  r <- readRDS(files[i])  # don't View(); just extract the piece we need
  e <- as.data.table(r$combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
  e[, sim := sim_id[i]]
  setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))

  fn_rds <- file.path(out_dir, sprintf("sim_%03d_eonr.rds", sim_id[i]))
  tmp    <- paste0(fn_rds, ".tmp")
  saveRDS(e, tmp, compress = "xz")
  file.rename(tmp, fn_rds)

  # optional CSV
  #fwrite(e, sub("\\.rds$", ".csv", fn_rds))

  rm(r, e)
  if (i %% 50 == 0) gc()
}

```

### 2025-08-14 codes/R/CF_Continuous_Loop_new.Rmd
Initial commit to CF_Continuous repo

--- 2025-07-30 codes/R/CF_Continuous_Loop.Rmd
@@
 library(mgcv)
-# library(here)
-library(reticulate)
+#library(here)
+#library(reticulate)
 library(readr)
@@
 library(tidyr)
+
+Sys.setenv(
+  OMP_NUM_THREADS = "1",
+  OPENBLAS_NUM_THREADS = "1",
+  MKL_NUM_THREADS = "1",
+  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
+)
+
+library(reticulate)
+use_condaenv("cf_conda311", required = TRUE)
+py_config()
+
+py_run_string("import sys, os; sys.path.append('/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python')")
+
+# clear any stale bytecode that might have been written under a different toolchain
+system("find /Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python -name '__pycache__' -type d -prune -exec rm -rf {} +")
+
+# now import without executing R helpers:
+py_run_string("import run_CF_c; print('run_CF_c imported OK; has run_CF_c_py:', hasattr(run_CF_c, 'run_CF_c_py'))")
+
+#--- python functions ---#
+# use_condaenv("cf_conda311", required = TRUE)
+# py_config()
+# py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
+source_python("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
 
@@
 ```
-
-```{r}
+```{r setup, include=FALSE}
 # project‐root path:
@@
 
-# “.here” file
+# ensure a `.here` exists at the project root (optional)
 here_file <- file.path(proj_root, ".here")
-if (!file.exists(here_file)) {
-  message("Creating .here in ", proj_root)
-  file.create(here_file)
-}
+if (!file.exists(here_file)) file.create(here_file)
 
+# turn off multi-threading
+Sys.setenv(OMP_NUM_THREADS = 1, MKL_NUM_THREADS = 1)
+
+# load here (no need to set root—see next chunk)
+library(here)
 ```
@@
 ```{r}
-# turn off multi-threading
-Sys.setenv(OMP_NUM_THREADS = 1, MKL_NUM_THREADS = 1)
+# Data (use here() with absolute root to avoid any cached project root)
+rds_path <- here(proj_root, "data", "raw_sim_data.rds")
+if (!file.exists(rds_path)) stop("File not found: ", rds_path)
 
-# load data only once, then drop
-library(here)
-raw_sim_data <- readRDS(here("data", "raw_sim_data.rds"))
-rm(list = setdiff(ls(), c("raw_sim_data", "np", "run_CF_c_py", "gm_wrapper_functions_you_need")))
+raw_sim_data <- readRDS(rds_path)
+
+# keep memory tidy
+rm(list = setdiff(ls(), c("raw_sim_data", "np", "run_CF_c_py", "gm_wrapper_functions_you_need", "proj_root")))
 gc()
@@
 # how many fields to train on?
-n_fields <- 1
+n_fields <- 10
 
@@
 # the right split file under <project_root>/data/train_test_split
+proj_root <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous"
+
 splits_file <- here(
+  proj_root,
   "data", "train_test_split",
@@
 )
-output_dir <- here(
-  "results",
+
+output_dir <- file.path(
+  "/Users/mmousavi2/Desktop/tes_one_loop",
   sprintf("CF_outcome_%d_fields", n_fields)
@@
 
+#train_test_splits <- train_test_splits_10fields
+
 # ─── MODEL LOOP ────────────────────────────────────────────────────────────────
@@
 # cf_results <- lapply(train_test_splits$test_id, function(test_sim_id)
-cf_results <- lapply(train_test_splits$test_id, function(test_sim_id) {
+cf_results <- lapply(train_test_splits$test_id[1], function(test_sim_id) {
   result_file <- file.path(output_dir, paste0("sim_", test_sim_id, ".rds"))
@@
 ```
+# rmse
+
+```{r}
+rmse_1_fields_one_loop <- sqrt(mean((sim_001_eonr$opt_N_hat - sim_001_eonr$opt_N)^2, na.rm = TRUE))  
+rmse_1_fields_one_loop
+```

### 2025-08-14 codes/R/CF_loop.Rmd
CF loop still needs attention to make it organaized
Initial commit to CF_Continuous repo

--- 2025-08-12 CF_loop.Rmd
@@
 source_python("/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
-
 ```
 
-# CF
+# CF loop
 
 ```{r}
-
 ## ───────────────────────────────────────
@@
 ## ─  keep an in-memory list ────────────
-#cf_all_results <- vector("list", length = nrow(train_test_splits))
-#names(cf_all_results) <- train_test_splits$test_id
-train_test_splits <- subset(train_test_splits, !test_id %in% as.integer(sub("^sim_(\\d+)\\.rds$","\\1", basename(list.files(path.expand(output_dir), "^sim_\\d+\\.rds$")))))
+# cf_all_results <- vector("list", length = nrow(train_test_splits))
+# names(cf_all_results) <- train_test_splits$test_id
+train_test_splits <- subset(train_test_splits, !test_id %in% as.integer(sub("^sim_(\\d+)\\.rds$", "\\1", basename(list.files(path.expand(output_dir), "^sim_\\d+\\.rds$")))))
 
@@
 for (i in seq_len(nrow(train_test_splits))) {
-
   test_sim_id <- train_test_splits$test_id[i]
@@
   #
-  tryCatch({
-
+  tryCatch(
+    {
       ## 1) Train/test IDs
@@
       #  keep in memory too
-    #cf_all_results[[as.character(test_sim_id)]] <- res
+      # cf_all_results[[as.character(test_sim_id)]] <- res
 
@@
 
-    message("Finished sim ", test_sim_id, " → ", basename(out_fn),
-            " (corr = ", round(corr_val, 3), ")")
+      message(
+        "Finished sim ", test_sim_id, " → ", basename(out_fn),
+        " (corr = ", round(corr_val, 3), ")"
+      )
 
@@
       if (i %% 10 == 0) gc()
-
-  }, error = function(e) {
+    },
+    error = function(e) {
       message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
@@
       )
-  })
-
+    }
+  )
 }
@@
 # saveRDS(cf_all_results, file.path(output_dir, "_all_results.rds"))
+```
 
-```
 # read in
@@
   # optional CSV
-  #fwrite(e, sub("\\.rds$", ".csv", fn_rds))
+  # fwrite(e, sub("\\.rds$", ".csv", fn_rds))
 
@@
 }
-
 ```

### 2025-08-14 codes/R/CF_Loop.qmd
Reproducable CF loop

---
title: "CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

#--- python functions ---#
use_condaenv("cf_conda311", required = TRUE)
py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")

#--- R functions ---#
source(here("codes", "R", "Utility.R"))

#--- python functions ---#
source_python(here::here("codes/Python/run_CF_c.py"))
```

## Data

```{r}
raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 10

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

progress_csv <- file.path(output_dir, "progress.csv")

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────

train_test_splits <- read_csv(splits_file)

```

# CF loop

```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(nrow(train_test_splits)))
#for (i in seq_len(1))
for (i in seq_len(nrow(train_test_splits))) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch(
    {
      ## 1) train/test IDs
      split_row     <- train_test_splits[i, ]
      train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
      sim_id_ls     <- c(train_sim_ids, test_sim_id)

      ## 2) train & test data
      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      train_data <- data[sim != test_sim_id, ]
      test_data  <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      ## 3) CF 
      x_vars <- c("Nk", "plateau", "b0")
      T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
      Y      <- train_data[, yield]
      X      <- as.matrix(train_data[, ..x_vars])
      W      <- X
      X_test <- as.matrix(test_data[, ..x_vars])

      ## 4) run cf in python
      Sys.setenv(
        OMP_NUM_THREADS      = "1",
        MKL_NUM_THREADS      = "1",
        KMP_AFFINITY         = "disabled",
        KMP_DUPLICATE_LIB_OK = "TRUE"
      )

      te_hat_cf <- run_CF_c_py(
        Y,
        T_info$T_sp,
        X,
        W,
        n_estimators = 2000
      )
      te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

      ## 5) Response curves
      T_seq <- train_data[, seq(min(N), max(N), length = 200)]
      response_data <- find_response_semi(T_seq, T_info, te_info)

      ## 6) Site-specific EONR
      pCorn <- 6.25 / 25.4
      pN    <- 1 / 0.453592

      ss_eonr <- response_data %>%
        .[, profit := est * pCorn - pN * T] %>%
        .[, .SD[which.max(profit), ], by = aunit_id] %>%
        .[, .(aunit_id, opt_N_hat = T)]

      true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
        .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
        .[, opt_N := pmin(Nk, opt_N)] %>%
        .[, opt_N := pmax(0,  opt_N)] %>%
        .[, .(aunit_id, opt_N)]

      combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
      corr_val <- combined_eonr[, cor(opt_N, opt_N_hat)]

      ## 7) SAVE EONR 
      e <- as.data.table(combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
      e[, sim := as.integer(test_sim_id)]
      data.table::setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))

      fn_eonr   <- file.path(eonr_dir, sprintf("sim_%03d_eonr.rds", as.integer(test_sim_id)))
      tmp_eonr  <- paste0(fn_eonr, ".tmp")
      saveRDS(e, tmp_eonr, compress = "xz")
      file.rename(tmp_eonr, fn_eonr)

      # optional
      csv_path <- file.path(eonr_dir, "eonr_progress.csv")
      data.table::fwrite(e, csv_path, append = file.exists(csv_path))

      ## 8) save the full result object 
      # res <- list(
      #   te_hat_cf     = te_hat_cf,
      #   te_info       = te_info,
      #   T_seq         = T_seq,
      #   response_data = response_data,
      #   test_data     = test_data,
      #   train_data    = train_data,
      #   combined_eonr = combined_eonr,
      #   corr          = corr_val
      # )
      # 
      # tmp_fn <- paste0(out_fn, ".tmp")
      # saveRDS(res, tmp_fn, compress = "xz")
      # file.rename(tmp_fn, out_fn)

      ## 9) Log progress
      prog <- data.frame(
        test_sim_id = test_sim_id,
        corr        = corr_val,
        n_train     = nrow(train_data),
        saved_file  = out_fn,
        eonr_file   = fn_eonr,
        ts          = as.character(Sys.time())
      )
      write.table(
        prog,
        file      = progress_csv,
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(progress_csv),
        append    = TRUE
      )

      message(
        "Finished sim ", test_sim_id, " → ", basename(out_fn),
        " | eonr: ", basename(fn_eonr),
        " (corr = ", round(corr_val, 3), ")"
      )

      # Free ram 
      rm(res, te_hat_cf, te_info, response_data, train_data, test_data, e)
      if (i %% 10 == 0) gc()
    },
    error = function(e) {
      message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
      err_row <- data.frame(
        test_sim_id = test_sim_id,
        error       = conditionMessage(e),
        ts          = as.character(Sys.time())
      )
      write.table(
        err_row,
        file      = file.path(output_dir, "errors.csv"),
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(file.path(output_dir, "errors.csv")),
        append    = TRUE
      )
    }
  )
}

```

### 2025-08-17 codes/R/visualize_yield_response_function.R
Dependency function to visualize yield response function

@@ -1,0 +1,91 @@
+# Requirements
+suppressPackageStartupMessages({
+  library(tidyverse)
+  library(here)
+})
+
+# Folder label helper
+fields_label <- function(n_fields) {
+  if (n_fields == 1) "one_field"
+  else if (n_fields == 5) "five_fields"
+  else if (n_fields == 10) "ten_fields"
+  else paste0(n_fields, "_fields")
+}
+
+# Build path to the CSV using here()
+results_csv_path <- function(model, n_fields, test_id = 1,
+                             base_dir = here("results", "yield_response_function_for_one_iteration")) {
+  subdir <- paste0("YRF_", model, "_", fields_label(n_fields))
+  file.path(base_dir, subdir, sprintf("yield_response_%s.csv", test_id))
+}
+
+# Main plotting function
+# Example: plot_yield_response(model = "NO_ANN", n_fields = 5)
+plot_yield_response <- function(model = "NO_ANN",
+                                n_fields = 5,
+                                test_id = 1,
+                                n_ids = 5,
+                                seed = 123,
+                                fixed_scales = TRUE,
+                                group_col = NULL) {
+  csv_path <- results_csv_path(model, n_fields, test_id)
+  
+  if (!file.exists(csv_path)) {
+    stop("File not found: ", csv_path, call. = FALSE)
+  }
+  
+  df <- readr::read_csv(csv_path, show_col_types = FALSE)
+  
+  # x-axis: prefer N, else N_tilde
+  x_col <- if ("N" %in% names(df)) "N" else if ("N_tilde" %in% names(df)) "N_tilde" else
+    stop("Neither 'N' nor 'N_tilde' column found in: ", csv_path, call. = FALSE)
+  
+  # grouping column: user-specified or auto (aunit_id -> row_id -> cell_id -> field_id)
+  if (is.null(group_col)) {
+    candidates <- c("aunit_id", "row_id", "cell_id", "field_id")
+    group_col <- candidates[candidates %in% names(df)][1]
+    if (is.na(group_col)) stop("No suitable grouping column found in file.")
+  } else if (!group_col %in% names(df)) {
+    stop("Requested group_col '", group_col, "' not found in file.")
+  }
+  
+  # pick IDs
+  set.seed(seed)
[…]

### 2025-08-17 codes/R/YRF_visualized.qmd
present results (yield response)

---
title: "yield response function"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- R functions ---#
library(here)
source(here("codes", "R", "visualize_yield_response_function.R"))
```

# Use it:

```{r}
plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
plot_yield_response(model = "SO_ANN", n_fields = 10) 
plot_yield_response(model = "NO_RF", n_fields = 5) 
plot_yield_response(model = "SO_RF", n_fields = 5) 
```

### 2025-08-19 codes/R/visualized_results.qmd
minor edits

---
title: "yield response function"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- R functions ---#
library(here)
source(here("codes", "R", "visualize_yield_response_function.R"))
```

# Use it:

```{r}
plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
plot_yield_response(model = "SO_ANN", n_fields = 10) 
plot_yield_response(model = "NO_RF", n_fields = 5) 
plot_yield_response(model = "SO_RF", n_fields = 5) 
```
# RMSE

```{r}
# ============================
# create per-sim RMSE table 
# ============================

# ---- Packages ----
library(tidyverse)
library(fs)
library(here)
library(readr)
library(stringr)

# ---- Constants: all 18 result folders ----
MODEL_DIRS <- c(
  "CF_outcome_1_fields","CF_outcome_5_fields","CF_outcome_10_fields",
  "RANN_outcome_one_field","RANN_outcome_five_fields","RANN_outcome_ten_fields",
  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields",
  "NO_RF_one_field","NO_RF_five_fields","NO_RF_ten_fields",
  "HalfRRF_one_field","HalfRRF_five_fields","HalfRRF_ten_fields",
  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields"
)

EXPECTED_ROWS <- 1440L

# ---- function ----
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

parse_n_fields <- function(folder) {
  case_when(
    str_detect(folder, "(?:_1_fields|one_field)\\b")   ~ 1L,
    str_detect(folder, "(?:_5_fields|five_fields)\\b") ~ 5L,
    str_detect(folder, "(?:_10_fields|ten_fields)\\b") ~ 10L,
    TRUE ~ NA_integer_
  )
}

parse_model <- function(folder) {
  f <- str_to_lower(folder)
  case_when(
    str_starts(f, "cf_")             ~ "CF",
    str_starts(f, "rann_outcome")    ~ "RANN",
    str_starts(f, "half_rann")       ~ "HalfRANN",
    str_starts(f, "simple_ann")      ~ "Simple_ANN",
    str_starts(f, "no_rf")           ~ "NO_RF",
    str_starts(f, "halfrrf")         ~ "HalfRRF",
    TRUE                             ~ folder
  )
}

# Read any object: RDS, CSV
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

# Standardize to two numeric vectors (per-sim rows)
read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.list(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF list
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# 
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    # Names like sim_055_eonr or EONR_1 
    keep(~ str_detect(path_file(.x),
           regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                 ignore_case = TRUE))) %>%
    # drop anything starting with "validation"
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    # drop macOS Icon files
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# get sim id from filename base
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

# 
row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

    e <- read_eonr_file(f)
    n_rows <- length(e$pred)

    tibble(
      sim      = sim_id,
      Model    = model,
      `#field` = n_f,
      n_rows   = n_rows,
      RMSE     = rmse(e$pred, e$true),
      ok_rows  = isTRUE(n_rows == expected_rows),
      file     = f
    )
  }, error = function(e) NULL)
}

# ---- create table  ----
build_rmse_table <- function(base_dir = here("results"),
                             include_dirs = MODEL_DIRS,
                             expected_rows = EXPECTED_ROWS) {
  dirs <- file.path(base_dir, include_dirs)
  dirs <- dirs[dir_exists(dirs)]

  map_dfr(dirs, function(dir_path) {
    folder <- path_file(dir_path)
    model  <- parse_model(folder)
    n_f    <- parse_n_fields(folder)

    files <- files_in_dir(dir_path)

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(sim), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}

```

# RMSE run

```{r}
# ---- run it (all models) ----
rmse_tbl_raw <- build_rmse_table(
  base_dir = here("results"),
  include_dirs = MODEL_DIRS,
  expected_rows = EXPECTED_ROWS
)

# save
# write_csv(rmse_tbl_raw, here("results", "rmse_table_all_models.csv"))
#saveRDS(rmse_tbl_raw, here("results", "rmse_tbl_raw.rds"))

```

# his

```{r}

library(dplyr)
library(ggplot2)
library(stringr)

#  model names: row labels in the graph
model_to_row <- c(
  "Simple_ANN" = "No_ANN",
  "RANN"       = "DO_ANN",
  "NO_RF"      = "NO_RF",
  "HalfRANN" = "SO_ANN",
  "HalfRRF"  = "SO_RF",
   "CF"       = "CF"
)

rmse_plot_df <- rmse_tbl_raw %>%
  filter(ok_rows, Model %in% names(model_to_row)) %>%
  mutate(
    row_lab = recode(Model, !!!model_to_row),
    col_lab = factor(`#field`,
                     levels = c(1, 5, 10),
                     labels = c("Number of fields: 1",
                                "Number of fields: 5",
                                "Number of fields: 10"))
  )

#  check 
rmse_plot_df %>% count(row_lab, col_lab)

# --- setup ---
xmin <- 0
nice_ceiling <- function(x, base = 5) base * ceiling(x / base)
xmax <- nice_ceiling(max(rmse_plot_df$RMSE, na.rm = TRUE), base = 5)

bins <- 25
binwidth <- (xmax - xmin) / bins

# --- Plot ---
p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
  facet_grid(row_lab ~ col_lab, scales = "fixed") +  
  scale_x_continuous(limits = c(xmin, xmax),
                     breaks = seq(xmin, xmax, length.out = 6),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "RMSE by iteration", y = "Count") +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 14)
  )

p

# ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)

```

# create rmse_summary

```{r}
# =========
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)

# 
rmse_df <- rmse_tbl_raw %>%
  dplyr::group_by(Model, `#field`, sim) %>%  
  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
  # Rename models
  dplyr::mutate(
    Model = dplyr::recode(
      Model,
      "Simple_ANN" = "NO_ANN",
      "HalfRANN"   = "SO_ANN",
      "RANN"       = "DO_ANN",
      "HalfRRF"    = "SO_RF",
      "CF"         = "CF",
      "NO_RF"      = "NO_RF",
      .default     = Model
    ),
    field_f = factor(`#field`, levels = c(1, 5, 10),
                     labels = c("1 field", "5 fields", "10 fields"))
  )

# 3) summary stats 
rmse_summary <- rmse_df %>%
  dplyr::group_by(Model, field_f) %>%
  dplyr::summarise(
    n      = dplyr::n(),
    mean   = mean(RMSE, na.rm = TRUE),
    median = median(RMSE, na.rm = TRUE),
    sd     = sd(RMSE, na.rm = TRUE),
    se     = sd / sqrt(n),
    lower  = mean + qt(0.025, df = n - 1) * se,
    upper  = mean + qt(0.975, df = n - 1) * se,
    .groups = "drop"
  )

```

# plot 1

```{r}
# 
p_mean <- ggplot(rmse_summary, aes(x = field_f, y = mean, group = Model, color = Model)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 3) +
  labs(x = "# fields", y = "Mean RMSE", color = "Model") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 13),
    strip.text = element_text(face = "bold")
  )

p_mean +
  geom_text(aes(label = sprintf("%.1f", mean)), vjust = -0.6, size = 3, show.legend = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.10)))

p_mean

# ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)
```

# plot 2

```{r}

fill_min <- floor(min(rmse_summary$mean, na.rm = TRUE))
fill_max <- ceiling(max(rmse_summary$mean, na.rm = TRUE))

p_clear <- ggplot(rmse_summary, aes(x = field_f, y = Model, fill = mean)) +
  geom_tile(color = "white", linewidth = 0.7) +
  geom_text(aes(label = sprintf("%.1f", mean)),
            size = 3.5, fontface = "bold", color = "black") +
  scale_fill_gradient(
    name = "Mean RMSE",
    low = "#f7fbff",  
    high = "#6baed6", 
    limits = c(fill_min, fill_max)
  ) +
  coord_equal() +
  labs(x = "# fields", y = "model") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
    #plot.margin = margin(10, 10, 10, 10)
  )

p_clear

# ggsave(here::here("results","rmse_mean_graph.png"), p_clear, width = 7.5, height = 4.5, dpi = 300)

```

# table

```{r}

# ------
rmse_summary %>%
  arrange(Model, field_f) %>%
  mutate(`mean (sd)` = sprintf("%.2f (%.2f)", mean, sd)) %>%
  select(Model, field_f, n, `mean (sd)`, median, lower, upper) %>%
  print(n = Inf)

```

### 2025-08-22 Writing/manuscript_CF_continuous.qmd
Changing DO-ANN (not using spline)

@@
 
-The DO-ANN follows the same double-orthogonalized, linear-in-basis logic as DO-CF, adapted to the Model 3.
-Starting from the residualized outcome $\tilde{y}$, the continuous treatment $N$ is expanded into spline basis functions and each basis is orthogonalized with respect to $(X,W)$; these orthogonalized spline bases are $T_1, T_2, T_3$.
-The model assumes
+In the DO-ANN approach, we apply double-orthogonalization to Model 3 by residualizing both the outcome $y$ and the continuous treatment $N$ with respect to $(X, W)$ using cross-fitted random forests:
 
 ```{=tex}
-\begin{align*}
-\tilde{y} \approx \theta_1(X) T_1+\theta_2(X) T_2+\theta_3(X) T_3,
+\tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
 \end{align*}
@@
 
-so that each $\theta_k(X)$ is the heterogeneous weight on the corresponding orthogonalized spline component of $N$.
-The key difference from DO-CF is the estimator for the coefficient functions $\theta_k(X)$: in DO-CF these are learned nonparametrically via causal forests, whereas in DO-ANN they are approximated by three separate neural-network branches, each producing a scalar weight that multiplies one of the $T_k$; the branch outputs are summed to produce the predicted residual.
-Because the decomposition is linear in the orthogonalized spline bases, the same clean double-orthogonalization justification applies, and the learned branch outputs recover the heterogeneous response without introducing first-order bias from the nuisance fits.
-The combined output reconstructs $\hat{f}(X, N)$, which is then used in the economic objective to compute the EONR.
+The final-stage learner is a neural network that models $\tilde{y}$ as a linear function of $\tilde{N}$, with heterogeneity in the slope captured by covariates $X$:
+
+```{=tex}
+\hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
+\end{align*}
+```
+
+This structure preserves the linear-in-treatment-residual form required for valid double-orthogonalization while allowing $\hat{\theta}(X)$ to vary flexibly across sites.
+The neural network receives $X$ as input and outputs the site-specific treatment effect.
+
+After training, we reconstruct $\hat{f}(X, N)$ over a grid of candidate nitrogen rates by evaluating the model across values of $N$, and select the economically optimal nitrogen rate.
 

### 2025-08-24 codes/R/visualized_results.qmd
comparison plot added

@@
 ```
+
+# comparison plot
+
+```{r}
+# compare to which model?
+baseline <- "NO_RF"
+
+# average RMSE by Model × field
+avg_by <- rmse_df %>%
+  dplyr::group_by(Model, field_f) %>%
+  dplyr::summarise(mean_rmse = mean(RMSE), .groups = "drop")
+
+# baseline per field 
+baseline_by_field <- avg_by %>%
+  dplyr::filter(Model == baseline) %>%
+  dplyr::select(field_f, baseline_rmse = mean_rmse)
+
+# join and compute % improvement
+rel <- avg_by %>%
+  dplyr::left_join(baseline_by_field, by = "field_f") %>%
+  dplyr::filter(!is.na(baseline_rmse)) %>%
+  dplyr::mutate(
+    pct_impr = 100 * (baseline_rmse - mean_rmse) / baseline_rmse
+    # or: pct_impr = (1 - mean_rmse / baseline_rmse) * 100
+  ) %>%
+  dplyr::filter(Model != baseline)
+
+# Plot
+library(ggplot2)
+p_rel <- ggplot(rel, aes(x = field_f, y = pct_impr, fill = Model)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.65) +
+  geom_hline(yintercept = 0, linewidth = 0.4) +
+  geom_text(aes(label = sprintf("%.1f%%", pct_impr)),
+            position = position_dodge(width = 0.7), vjust = -0.4, size = 3, show.legend = FALSE) +
+  scale_y_continuous(expand = expansion(mult = c(0.02, 0.10))) +
+  labs(x = "# fields", y = sprintf("Improvement vs %s (%%)", baseline), fill = "Model") +
+  theme_bw()
+
+p_rel
+
+```

### 2025-08-24 codes/R/visualize_yield_response_function.R
Some modifications to the yield response function

@@ -86,6 +86,7 @@ plot_yield_response <- function(model = "NO_ANN",
     p <- p + scale_y_continuous(limits = c(y_min, y_max))
   }
   
-  print(p)
-  invisible(list(plot = p, file = csv_path, ids = pick, x_col = x_col, group_col = group_col))
+  #print(p)
+  #invisible(list(plot = p, file = csv_path, ids = pick, x_col = x_col, group_col = group_col))
+  p
 }

### 2025-08-26 Writing/manuscript_CF_continuous.qmd
Results

@@
 
-# Results and Discussions
+# Simulation Results
+
+We compare six estimators—CF, SO_ANN, DO_ANN, NO_ANN, SO_RF, and NO_RF—over 500 Monte-Carlo replications and three training sizes (1, 5, and 10 fields).
+Each replication contains 1,440 evaluation units.
+Accuracy is assessed by the RMSE of the estimated economically optimal nitrogen rate (EONR) with respect to the true (simulated) EONR.
+Figures 1–4 summarize errors; Figure 5 panels illustrate representative yield–N response functions that help interpret the error patterns.
+
+```{r setup, include=FALSE}
+
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+
+suppressPackageStartupMessages(library(here))
+suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(stringr))
+
+# load data
+rmse_tbl_raw <- readRDS(here::here("results", "rmse_tbl_raw.rds"))
+```
+
+@fig-rmse-hist shows a clear ordering across methods.
+CF concentrates tightly at low RMSE for all training sizes, indicating both accuracy and stability.
+SO_ANN shifts steadily left as the number of fields increases, with visibly narrower spreads at 5 and 10 fields.
+In contrast, NO_ANN exhibits a heavy right tail with a single field but collapses toward the leading methods at 5–10 fields, showing its sensitivity to sample size.
+SO_RF improves with more fields yet remains less accurate than SO_ANN or CF.
+NO_RF has broad, right-shifted distributions and is a weak baseline.
+DO_ANN performs poorly across all sizes, with distributions centered at very high RMSE and little improvement from additional fields.
+
+```{r}
+#| label: fig-rmse-hist
+#| fig-cap: "RMSE histograms by model and number of fields."
+#| fig-width: 12
+#| fig-height: 7
+
+# model names: row labels in the graph
+model_to_row <- c(
+  "Simple_ANN" = "No_ANN",
+  "RANN"       = "DO_ANN",
+  "NO_RF"      = "NO_RF",
+  "HalfRANN"   = "SO_ANN",
+  "HalfRRF"    = "SO_RF",
+  "CF"         = "CF"
+)
+
+rmse_plot_df <- rmse_tbl_raw %>%
+  dplyr::filter(ok_rows, Model %in% names(model_to_row)) %>%
+  dplyr::mutate(
+    row_lab = dplyr::recode(Model, !!!model_to_row),
+    col_lab = factor(`#field`,
+                     levels = c(1, 5, 10),
+                     labels = c("Number of fields: 1",
+                                "Number of fields: 5",
+                                "Number of fields: 10"))
+  )
+
+# consistent axes
+xmin <- 0
+nice_ceiling <- function(x, base = 5) base * ceiling(x / base)
+xmax <- nice_ceiling(max(rmse_plot_df$RMSE, na.rm = TRUE), base = 5)
+bins <- 25
+binwidth <- (xmax - xmin) / bins
+
+p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
+  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
+  facet_grid(row_lab ~ col_lab, scales = "fixed") +
+  scale_x_continuous(limits = c(xmin, xmax),
+                     breaks = seq(xmin, xmax, length.out = 6),
+                     expand = c(0, 0)) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  labs(x = "RMSE by iteration", y = "Count") +
+  theme_bw() +
+  theme(
+    strip.text = element_text(face = "bold", size = 16),
+    axis.title = element_text(size = 14)
+  )
+
+p
+```
+
+Mean RMSE by model and training size (@fig-avg-performance and @tbl-mean-rmse-heatmap) confirms these patterns.
+CF is consistently low (≈15 across sizes).
+SO_ANN improves monotonically as data increase and attains the best mean at 10 fields (≈11).
+NO_ANN is unusable with a single field (mean ≈82) but becomes competitive by 5–10 fields (≈14 and ≈13).
+SO_RF yields moderate means (≈22 at 10 fields), while NO_RF remains highest among the non-ANN baselines (≈36–40).
+DO_ANN is uniformly poor (≈90–93).
+The heatmap (Fig.
+3) prints the means in each cell and emphasizes the strong gains from orthogonalization (SO_ANN and CF) relative to their non-orthogonal counterparts.
+
+```{r}
+#| label: fig-avg-performance
+#| fig-cap: "Average performance by model and number of fields."
+#| fig-width: 12
+#| fig-height: 7
+
+rmse_df <- rmse_tbl_raw %>%
+  dplyr::group_by(Model, `#field`, sim) %>%  
+  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
+  # Rename models
+  dplyr::mutate(
+    Model = dplyr::recode(
+      Model,
+      "Simple_ANN" = "NO_ANN",
+      "HalfRANN"   = "SO_ANN",
+      "RANN"       = "DO_ANN",
+      "HalfRRF"    = "SO_RF",
+      "CF"         = "CF",
+      "NO_RF"      = "NO_RF",
+      .default     = Model
+    ),
+    field_f = factor(`#field`, levels = c(1, 5, 10),
+                     labels = c("1 field", "5 fields", "10 fields"))
+  )
+
+# 3) summary stats 
+rmse_summary <- rmse_df %>%
+  dplyr::group_by(Model, field_f) %>%
+  dplyr::summarise(
+    n      = dplyr::n(),
+    mean   = mean(RMSE, na.rm = TRUE),
+    median = median(RMSE, na.rm = TRUE),
+    sd     = sd(RMSE, na.rm = TRUE),
+    se     = sd / sqrt(n),
+    lower  = mean + qt(0.025, df = n - 1) * se,
+    upper  = mean + qt(0.975, df = n - 1) * se,
+    .groups = "drop"
+  )
+
+# 
+p_mean <- ggplot(rmse_summary, aes(x = field_f, y = mean, group = Model, color = Model)) +
+  geom_line(linewidth = 0.7, alpha = 0.8) +
+  geom_point(size = 3) +
+  labs(x = "# fields", y = "Mean RMSE", color = "Model") +
+  theme_bw() +
+  theme(
+    axis.title = element_text(size = 13),
+    strip.text = element_text(face = "bold")
+  )
+
+p_mean
+
+# ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)
+```
+
+```{r}
+#| label: tbl-mean-rmse-heatmap
+#| tbl-cap: "Mean RMSE by model and number of fields."
+#| tbl-cap-location: bottom
+#| echo: false
+#| message: false
+#| warning: false
+
+library(dplyr)
+library(tidyr)
+library(gt)
+library(scales)
+
+# Range for coloring
+fill_min <- floor(min(rmse_summary$mean, na.rm = TRUE))
+fill_max <- ceiling(max(rmse_summary$mean, na.rm = TRUE))
+
+# Wide table of means
+rmse_table <- rmse_summary %>%
+  select(Model, field_f, mean) %>%
+  mutate(mean = round(mean, 1)) %>%
+  pivot_wider(names_from = field_f, values_from = mean) %>%
+  arrange(Model)
+
+# Light → mid-blue palette
+pal <- col_numeric(
+  palette = c("#f7fbff", "#6baed6"),
+  domain  = c(fill_min, fill_max)
+)
+
+# HTML colorbar legend to place under the table
+legend_html <- sprintf(
+  '<div style="display:flex;align-items:center;gap:8px;margin-top:6px;">
+     <span style="font-size:12px;color:#444;">Low (%.1f)</span>
+     <div style="flex:1;height:10px;background:linear-gradient(to right,#f7fbff,#6baed6);
+                 border:1px solid #ddd;border-radius:3px;"></div>
+     <span style="font-size:12px;color:#444;">High (%.1f)</span>
+   </div>', fill_min, fill_max)
+
+rmse_table %>%
+  gt(rowname_col = "Model") %>%
+  data_color(
+    columns = c(`1 field`, `5 fields`, `10 fields`),
+    colors  = pal
+  ) %>%
+  fmt_number(columns = c(`1 field`, `5 fields`, `10 fields`), decimals = 1) %>%
+  cols_label(
+    `1 field`  = "1 field",
+    `5 fields` = "5 fields",
+    `10 fields`= "10 fields"
+  ) %>%
+  tab_options(
+    table.font.size = px(13),
+    data_row.padding = px(6),
+    column_labels.font.weight = "bold"
+  ) %>%
+  tab_source_note(source_note = gt::html(legend_html))
+```
+
+Using NO_RF as a baseline, CF reduces mean RMSE by roughly 58–61 % across training sizes (@fig-improvement).
+SO_ANN delivers comparable or larger gains (≈54–69 %), with the largest improvement at 10 fields.
+SO_RF provides modest gains (≈38–42 %).
+NO_ANN is worse than the baseline with 1 field but improves substantially at 5–10 fields (≈63–65 % reduction).
+DO_ANN is consistently inferior to the baseline at all sizes.
+
+```{r}
+#| label: fig-improvement
+#| fig-cap: "Improvement relative to NO_RF."
+#| fig-width: 12
+#| fig-height: 7
+
+# compare to which model?
+baseline <- "NO_RF"
+
+# average RMSE by Model × field
+
+rmse_df <- rmse_tbl_raw %>%
+  dplyr::group_by(Model, `#field`, sim) %>%  
+  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
+  # Rename models
+  dplyr::mutate(
+    Model = dplyr::recode(
+      Model,
+      "Simple_ANN" = "NO_ANN",
+      "HalfRANN"   = "SO_ANN",
+      "RANN"       = "DO_ANN",
+      "HalfRRF"    = "SO_RF",
+      "CF"         = "CF",
+      "NO_RF"      = "NO_RF",
+      .default     = Model
+    ),
+    field_f = factor(`#field`, levels = c(1, 5, 10),
+                     labels = c("1 field", "5 fields", "10 fields"))
+  )
+
+avg_by <- rmse_df %>%
+  dplyr::group_by(Model, field_f) %>%
+  dplyr::summarise(mean_rmse = mean(RMSE), .groups = "drop")
+
+# baseline per field 
+baseline_by_field <- avg_by %>%
+  dplyr::filter(Model == baseline) %>%
+  dplyr::select(field_f, baseline_rmse = mean_rmse)
+
+# join and compute % improvement
+rel <- avg_by %>%
+  dplyr::left_join(baseline_by_field, by = "field_f") %>%
+  dplyr::filter(!is.na(baseline_rmse)) %>%
+  dplyr::mutate(
+    pct_impr = 100 * (baseline_rmse - mean_rmse) / baseline_rmse
+    # or: pct_impr = (1 - mean_rmse / baseline_rmse) * 100
+  ) %>%
+  dplyr::filter(Model != baseline)
+
+# Plot
+library(ggplot2)
+p_rel <- ggplot(rel, aes(x = field_f, y = pct_impr, fill = Model)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.65) +
+  geom_hline(yintercept = 0, linewidth = 0.4) +
+  geom_text(aes(label = sprintf("%.1f%%", pct_impr)),
+            position = position_dodge(width = 0.7), vjust = -0.4, size = 3, show.legend = FALSE) +
+  scale_y_continuous(expand = expansion(mult = c(0.02, 0.10))) +
+  labs(x = "# fields", y = sprintf("Improvement vs %s (%%)", baseline), fill = "Model") +
+  theme_bw()
+
+p_rel
+
+```
+
+Predicted yield–N curves (@fig-yield-res-1 through @fig-yield-res-5) for five randomly selected analysis units clarify why methods differ.
+SO_ANN produces agronomically plausible, monotone-concave responses with clear plateaus (ten-field panel), yielding interior optima and low EONR error.
+NO_ANN resembles SO_ANN once trained on 5–10 fields, but with a single field it is noticeably noisier, aligning with its heavy-tailed errors.
+Tree-based models (NO_RF and SO_RF) generate step-like, piecewise-constant responses; the implied EONR can only move at jump points, which coarsens the optimum and limits accuracy.
+DO_ANN often yields nearly linear increasing curves across the tested N range, pushing the optimum to the upper bound and explaining its uniformly high error.
+Although CF is not naturally expressed as per-unit yield curves, its error distribution indicates high stability across all sizes.
+
+```{r}
+#| label: fig-yield-res-1
+#| fig-cap: "Yield response function (NO_ANN; 5 fields)."
+#| fig-cap-location: bottom
+#| fig-width: 12
+#| fig-height: 7
+#| message: false
+#| warning: false
+
+source(here("codes", "R", "visualize_yield_response_function.R"))
+
+plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
+```
+```{r}
+#| label: fig-yield-res-2
+#| fig-cap: "Yield response function (DO_ANN; 5 fields)."
+#| fig-width: 12
+#| fig-height: 7
+
+source(here("codes", "R", "visualize_yield_response_function.R"))
+plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde
+
+```
+
+```{r}
+#| label: fig-yield-res-3
+#| fig-cap: "Yield response function SO_ANN (SO_ANN; 10 fields)."
+#| fig-width: 12
+#| fig-height: 7
+#| 
+source(here("codes", "R", "visualize_yield_response_function.R"))
+plot_yield_response(model = "SO_ANN", n_fields = 10)
+
+```
+
+```{r}
+#| label: fig-yield-res-4
+#| fig-cap: "Yield response function NO_RF (NO_RF; 5 fields)."
+#| fig-width: 12
+#| fig-height: 7
+#| 
+source(here("codes", "R", "visualize_yield_response_function.R"))
+plot_yield_response(model = "NO_RF", n_fields = 5)
+
+```
+
+```{r}
+#| label: fig-yield-res-5
+#| fig-cap: "Yield response function SO_RF (SO_RF; 5 fields)."
+#| fig-width: 12
+#| fig-height: 7
+#| 
+source(here("codes", "R", "visualize_yield_response_function.R"))
+plot_yield_response(model = "SO_RF", n_fields = 5) 
+```
 
 # Conclusion
+
+Across 500 replications, SO_ANN (with ≥10 fields) and CF deliver the best overall accuracy.
+Orthogonalization is critical: SO_ANN and CF are consistently superior to NO_ANN and NO_RF, especially in small-to-moderate samples.
+NO_ANN becomes competitive only once training data reach at least five fields.
+Tree-based baselines remain substantially less accurate, and the doubly-orthogonal ANN (DO_ANN) is inappropriate for this setting because it fails to recover the saturating response shape.
+These findings are robust across training sizes and are consistent.

### 2025-08-28 Writing/manuscript_CF_continuous.qmd
Adding citations

@@
 
-To identify causal effects rather than just predict levels in the single binary treatment case, researchers use double/debiased machine learning (DML), which introduces orthogonalization so the treatment-effect estimator is insensitive to small errors in the auxiliary fits (I need citation here).
+To identify causal effects rather than just predict levels in the single binary treatment case, researchers use double/debiased machine learning (DML), which introduces orthogonalization so the treatment-effect estimator is insensitive to small errors in the auxiliary fits (@chernozhukov2018double; @mackey2018orthogonal).
 Consider the structural model (Model 1):

### 2025-08-28 Writing/manuscript_CF_continuous.qmd
Citation added

@@
 The standard approach in the precision-agriculture literature for estimating site-specific input response is to model the outcome level directly as a function of the treatment/input and covariates—e.g., estimate $Y=f(I,C)$ where $I$ is the input (like nitrogen rate) and $C$ are field characteristics—and then derive optimal input decisions from that estimated function.
-This is what is typically done in the recent on-farm precision experimentation literature (e.g., @barbosa2020risk; @barbosa2020modeling; @krause2020random; @gardner2021economic).
+This is what is typically done in the recent on-farm precision experimentation literature (e.g., @de2023predicting; @barbosa2020risk; @barbosa2020modeling; @krause2020random; @gardner2021economic).
 The problem is that these methods optimize prediction of yield levels—not the causal effect of changing the input.

### 2025-08-29 codes/R/Add_b1_b2_to_data_2nd_stage.qmd
Add b1 and b2 to the data_2nd_stage

---
title: "Add b1 & b2 to the data_2nd_stage"
author: "Your Name Here"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

# Load data
```{r setup_not_remove, echo=FALSE, include = F}

library(here)

data_2nd_stage <- readRDS(here("data", "data_20230504", "data_2nd_stage.rds"))

```

# Add b1 & b2 into data_2nd_stage
```{r setup_not_remove, echo=FALSE, include = F}
library(data.table)

reg_data  <- raw_sim_data$reg_data[[1]]
# flatten the list-column `data` inside `reg_data`
reg_data_flat <- rbindlist(
  lapply(seq_len(nrow(reg_data)), function(i) {
    df <- reg_data$data[[i]]
    df$sim <- reg_data$sim[i]
    df
  }),
  use.names = TRUE, fill = TRUE
)

# keep only needed columns for the join
reg_data_subset <- reg_data_flat[, .(sim, aunit_id, b1, b2)]

# set as data.table 
setDT(data_2nd_stage)

# merge by sim and aunit_id
data_2nd_stage <- merge(
  data_2nd_stage,
  reg_data_subset,
  by = c("sim", "aunit_id"),
  all.x = TRUE,
  sort = FALSE
)

# save
saveRDS(
  data_2nd_stage,
  here("data", "data_20230504", "data_2nd_stage.rds")
)

```

### 2025-08-29 codes/R/visualize_yield_response_function.R
YRF changed to include true yield

@@ -1,10 +1,112 @@
-# Requirements
+# # Requirements
+# suppressPackageStartupMessages({
+#   library(tidyverse)
+#   library(here)
+# })
+# 
+# # Folder label helper
+# fields_label <- function(n_fields) {
+#   if (n_fields == 1) "one_field"
+#   else if (n_fields == 5) "five_fields"
+#   else if (n_fields == 10) "ten_fields"
+#   else paste0(n_fields, "_fields")
+# }
+# 
+# # Build path to the CSV using here()
+# results_csv_path <- function(model, n_fields, test_id = 1,
+#                              base_dir = here("results", "yield_response_function_for_one_iteration")) {
+#   subdir <- paste0("YRF_", model, "_", fields_label(n_fields))
+#   file.path(base_dir, subdir, sprintf("yield_response_%s.csv", test_id))
+# }
+# 
+# # Main plotting function
+# # Example: plot_yield_response(model = "NO_ANN", n_fields = 5)
+# plot_yield_response <- function(model = "NO_ANN",
+#                                 n_fields = 5,
+#                                 test_id = 1,
+#                                 n_ids = 5,
+#                                 seed = 123,
+#                                 fixed_scales = TRUE,
+#                                 group_col = NULL) {
+#   csv_path <- results_csv_path(model, n_fields, test_id)
+#   
+#   if (!file.exists(csv_path)) {
+#     stop("File not found: ", csv_path, call. = FALSE)
+#   }
+#   
+#   df <- readr::read_csv(csv_path, show_col_types = FALSE)
+#   
+#   # x-axis: prefer N, else N_tilde
+#   x_col <- if ("N" %in% names(df)) "N" else if ("N_tilde" %in% names(df)) "N_tilde" else
+#     stop("Neither 'N' nor 'N_tilde' column found in: ", csv_path, call. = FALSE)
+#   
+#   # grouping column: user-specified or auto (aunit_id -> row_id -> cell_id -> field_id)
+#   if (is.null(group_col)) {
+#     candidates <- c("aunit_id", "row_id", "cell_id", "field_id")
+#     group_col <- candidates[candidates %in% names(df)][1]
+#     if (is.na(group_col)) stop("No suitable grouping column found in file.")
+#   } else if (!group_col %in% names(df)) {
[…]

### 2025-08-29 codes/R/visualized_results.qmd
Results visualization changed to include true yield

@@
 ```{r}
-plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
-plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
-plot_yield_response(model = "SO_ANN", n_fields = 10) 
-plot_yield_response(model = "NO_RF", n_fields = 5) 
-plot_yield_response(model = "SO_RF", n_fields = 5) 
+# plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
+# plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
+# plot_yield_response(model = "SO_ANN", n_fields = 10) 
+# plot_yield_response(model = "NO_RF", n_fields = 5) 
+# plot_yield_response(model = "SO_RF", n_fields = 5)
+
+models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF")
+
+# compute global limits for you
+plot_yield_response("NO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("NO_RF",  1, global_models = models_to_compare)
+plot_yield_response("NO_RF",  5, global_models = models_to_compare)
+plot_yield_response("NO_RF",  10, global_models = models_to_compare)
+plot_yield_response("SO_RF",  1, global_models = models_to_compare)
+plot_yield_response("SO_RF",  5, global_models = models_to_compare)
+plot_yield_response("SO_RF",  10, global_models = models_to_compare)
 ```

### 2025-09-05 codes/R/Add_mean_match_to_YRF.qmd
Mean match for yield response function

---
title: "Add mean match to YRF"
author: "Your Name Here"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

# Add mean match
```{r setup_not_remove, echo=FALSE, include = F}

suppressPackageStartupMessages({
  library(here)
  library(readr)
  library(dplyr)
  library(stringr)
  library(purrr)
})

base_dir <- here::here("results", "yield_response_function_for_one_iteration")
if (!dir.exists(base_dir)) {
  stop("Base results directory not found: ", base_dir, call. = FALSE)
}

# mean-match predicted yield to true yield 
mean_match_file <- function(csv_path) {
  df <- suppressMessages(readr::read_csv(csv_path, show_col_types = FALSE))

  # Need both columns present
  if (!all(c("pred_yield", "true_yield") %in% names(df))) {
    message("Skipping (missing pred_yield/true_yield): ", csv_path)
    return(invisible(FALSE))
  }

  # global mean-match (single shift per file)
  delta <- mean(df$true_yield, na.rm = TRUE) - mean(df$pred_yield, na.rm = TRUE)

  df <- df %>%
    mutate(pred_yield_mm = pred_yield + delta) %>%
    relocate(pred_yield_mm, .after = pred_yield)

  readr::write_csv(df, csv_path)
  message(sprintf("Updated: %s  (delta = %.6f)", csv_path, delta))
  invisible(TRUE)
}

# find all YRF_* subfolders, but skip NO_* models
subdirs <- list.dirs(base_dir, recursive = FALSE, full.names = TRUE)
targets <- subdirs[!grepl("^YRF_NO_", basename(subdirs), ignore.case = TRUE)]

# process each yield_response_*.csv in folders
walk(targets, function(dir_path) {
  csvs <- list.files(dir_path, pattern = "^yield_response_.*\\.csv$", full.names = TRUE)
  if (!length(csvs)) {
    message("No yield_response_*.csv in: ", dir_path)
    return(invisible(NULL))
  }
  walk(csvs, mean_match_file)
})

```

### 2025-09-05 codes/R/visualize_yield_response_function.R
slightly modified

@@ -94,7 +94,7 @@
 
 
 
-#############$%^&&&&&&&&&&&&&&&&&&& add true yrf as well
+#############$%^&&&&&&&&&&&&&&&&&&& add true yrf and mean match as well
 
 
 suppressPackageStartupMessages({
@@ -138,31 +138,33 @@ compute_global_limits <- function(models,
                                   test_id = 1,
                                   base_dir = here("results", "yield_response_function_for_one_iteration")) {
   xmins <- c(); xmaxs <- c(); ymins <- c(); ymaxs <- c()
-
+  
   for (m in models) {
     f <- results_csv_path(m, n_fields, test_id, base_dir = base_dir)
     if (!file.exists(f)) next
     df <- suppressMessages(readr::read_csv(f, show_col_types = FALSE))
-
+    
     # choose x column (N preferred, else N_tilde)
     x_col <- if ("N" %in% names(df)) "N" else if ("N_tilde" %in% names(df)) "N_tilde" else NULL
     if (!is.null(x_col)) {
       xmins <- c(xmins, suppressWarnings(min(df[[x_col]], na.rm = TRUE)))
       xmaxs <- c(xmaxs, suppressWarnings(max(df[[x_col]], na.rm = TRUE)))
     }
-
-    # y from both predicted and true
-    y_vec <- df$pred_yield
+    
+    # y from predicted (prefer pred_yield_mm if present) and true
+    y_pred_col <- if ("pred_yield_mm" %in% names(df)) "pred_yield_mm" else "pred_yield"
+    y_vec <- df[[y_pred_col]]
     if ("true_yield" %in% names(df)) y_vec <- c(y_vec, df$true_yield)
+    
     ymins <- c(ymins, suppressWarnings(min(y_vec, na.rm = TRUE)))
     ymaxs <- c(ymaxs, suppressWarnings(max(y_vec, na.rm = TRUE)))
   }
-
+  
   x_lo <- suppressWarnings(min(xmins, na.rm = TRUE))
   x_hi <- suppressWarnings(max(xmaxs, na.rm = TRUE))
   y_lo <- suppressWarnings(min(ymins, na.rm = TRUE))
   y_hi <- suppressWarnings(max(ymaxs, na.rm = TRUE))
-
+  
   list(
     x = .pad_range(x_lo, x_hi),
     y = .pad_range(y_lo, y_hi)
@@ -183,17 +185,21 @@ plot_yield_response <- function(model = "NO_ANN",
[…]

### 2025-09-21 codes/R/visualized_results.qmd
Add visualization

--- 2025-08-24 codes/R/visualized_results.qmd
@@
 ```{r}
-plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
-plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
-plot_yield_response(model = "SO_ANN", n_fields = 10) 
-plot_yield_response(model = "NO_RF", n_fields = 5) 
-plot_yield_response(model = "SO_RF", n_fields = 5) 
+# plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
+# plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
+# plot_yield_response(model = "SO_ANN", n_fields = 10) 
+# plot_yield_response(model = "NO_RF", n_fields = 5) 
+# plot_yield_response(model = "SO_RF", n_fields = 5)
+
+models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF")
+
+# 
+plot_yield_response("NO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 5, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("NO_RF",  1, global_models = models_to_compare)
+plot_yield_response("NO_RF",  5, global_models = models_to_compare)
+plot_yield_response("NO_RF",  10, global_models = models_to_compare)
+plot_yield_response("SO_RF",  1, global_models = models_to_compare)
+plot_yield_response("SO_RF",  5, global_models = models_to_compare)
+plot_yield_response("SO_RF",  10, global_models = models_to_compare)
 ```
+
 # RMSE
@@
   "RANN_outcome_one_field","RANN_outcome_five_fields","RANN_outcome_ten_fields",
-  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields",
+  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_two_fields","Simple_ANN_outcome_three_fields", "Simple_ANN_outcome_four_fields",
+  "Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields","Simple_ANN_outcome_twenty_fields",
   "NO_RF_one_field","NO_RF_five_fields","NO_RF_ten_fields",
   "HalfRRF_one_field","HalfRRF_five_fields","HalfRRF_ten_fields",
-  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields"
+  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields", "half_RANN_outcome_twenty_fields"
 )
@@
     str_detect(folder, "(?:_1_fields|one_field)\\b")      ~ 1L,
-    str_detect(folder, "(?:_5_fields|five_fields)\\b") ~ 5L,
-    str_detect(folder, "(?:_10_fields|ten_fields)\\b") ~ 10L,
+    str_detect(folder, "(?:_2_fields|two_fields?)\\b")    ~ 2L,
+    str_detect(folder, "(?:_3_fields|three_fields?)\\b")  ~ 3L,
+    str_detect(folder, "(?:_4_fields|four_fields?)\\b")   ~ 4L,
+    str_detect(folder, "(?:_5_fields|five_fields?)\\b")   ~ 5L,
+    str_detect(folder, "(?:_10_fields|ten_fields?)\\b")   ~ 10L,
+    str_detect(folder, "(?:_20_fields|twenty_fields?)\\b")~ 20L,
     TRUE ~ NA_integer_
@@
     col_lab = factor(`#field`,
-                     levels = c(1, 5, 10),
+                     levels = c(1, 2, 3, 4, 5, 10, 20),
                      labels = c("Number of fields: 1",
+                                "Number of fields: 2",
+                                "Number of fields: 3",
+                                "Number of fields: 4",
                                 "Number of fields: 5",
-                                "Number of fields: 10"))
+                                "Number of fields: 10",
+                                "Number of fields: 20"))
   )
@@
     ),
-    field_f = factor(`#field`, levels = c(1, 5, 10),
-                     labels = c("1 field", "5 fields", "10 fields"))
+    field_f = factor(`#field`, levels = c(1, 2, 3, 4, 5, 10, 20),
+                     labels = c("1 field", "2 fields", "3 fields", "4 fields",  "5 fields", "10 fields", "20 fields"))
   )
@@
 ```
+
+# Best model by order
+
+```{r}
+library(dplyr)
+library(stringr)
+
+best_models <- rmse_summary %>%
+  mutate(field_n = as.numeric(str_extract(field_f, "\\d+"))) %>%
+  group_by(field_n) %>%
+  #  the best model 
+  slice_min(order_by = mean, with_ties = FALSE) %>%
+  ungroup() %>%
+  # order by RMSE 
+  arrange(mean, field_n) %>%
+  select(Model, field_f, mean)
+
+## Vis
+
+library(gt)
+
+best_models %>%
+  gt() %>%
+  tab_header(
+    title = "Best Models by Field Count",
+    subtitle = "Ordered by lowest mean RMSE"
+  ) %>%
+  fmt_number(
+    columns = c(mean),
+    decimals = 3
+  ) %>%
+  cols_label(
+    field_f = "# Fields",
+    mean    = "mean RMSE"
+
+  ) %>%
+  tab_options(table.font.size = "small")
+
+```

### 2025-09-21 codes/Python/Functions/simple_ANN.py
Codes modification

@@ -130,8 +130,12 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):
     split_csv_path = f'./data/train_test_split/train_test_splits_{n_fields}fields.csv'
     folder_map = {
         ("SimpleANN", 1): "Simple_ANN_outcome_one_field",
+        ("SimpleANN", 2): "Simple_ANN_outcome_two_fields",
+        ("SimpleANN", 3): "Simple_ANN_outcome_three_fields",
+        ("SimpleANN", 4): "Simple_ANN_outcome_four_fields",
         ("SimpleANN", 5): "Simple_ANN_outcome_five_fields",
-        ("SimpleANN", 10): "Simple_ANN_outcome_ten_fields"
+        ("SimpleANN", 10): "Simple_ANN_outcome_ten_fields",
+        ("SimpleANN", 20): "Simple_ANN_outcome_twenty_fields"
     }
     folder_name = folder_map.get((model_type, n_fields))
     if folder_name is None:

### 2025-09-21 codes/Python/Functions/Generate_train_test_splits.py
Codes modification

@@ -41,10 +41,22 @@ print("GPU is", "available" if torch.cuda.is_available() else "NOT AVAILABLE")
 
 # %%
 # === set directory ===
+#import os
+#notebook_dir = "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous"  
+#os.chdir(notebook_dir)  # Change directory
+#print(os.getcwd())  
+
+# === set directory ===
+from pathlib import Path
 import os
-notebook_dir = "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous"  
-os.chdir(notebook_dir)  # Change directory
-print(os.getcwd())  
+
+# project root = two levels up from this script (../CF_Continuous)
+project_root = Path(__file__).resolve().parents[3]
+notebook_dir = project_root  # CF_Continuous root
+
+os.chdir(notebook_dir)
+print("Working dir:", os.getcwd())
+
 
 # %%
 # === LOAD DATA ===
@@ -67,10 +79,28 @@ def create_split_csv(n_fields, output_path):
     df = pd.DataFrame(splits)
     df.to_csv(output_path, index=False)
 
+from pathlib import Path
+import os
+
+def maybe_create_split(n_fields, filename):
+    if not Path(filename).exists():
+        create_split_csv(n_fields, filename)
+        print(f"✅ Created {filename}")
+    else:
+        print(f"⚡ Already exists: {filename}")
+
+
+
 os.makedirs("./data/train_test_split", exist_ok=True)
-create_split_csv(1, './data/train_test_split/train_test_splits_1fields.csv')
-create_split_csv(5, './data/train_test_split/train_test_splits_5fields.csv')
-create_split_csv(10, './data/train_test_split/train_test_splits_10fields.csv')
+
+# list of splits
+field_counts = [1,2,3,4, 5, 10, 20]
+
+
+# loop over them
+for n in field_counts:
+    filename = f'./data/train_test_split/train_test_splits_{n}fields.csv'
+    maybe_create_split(n, filename)
 
 # %%
 

### 2025-09-21 codes/Python/Functions/half_RANN.py

@@ -132,7 +132,8 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):
     folder_map = {
         ("half_RANN", 1): "half_RANN_outcome_one_field",
         ("half_RANN", 5): "half_RANN_outcome_five_fields",
-        ("half_RANN", 10): "half_RANN_outcome_ten_fields"
+        ("half_RANN", 10): "half_RANN_outcome_ten_fields",
+        ("half_RANN", 20): "half_RANN_outcome_twenty_fields" 
     }
     folder_name = folder_map.get((model_type, n_fields))
     if folder_name is None:

### 2025-09-21 codes/Python/Functions/main_simulation_run.py

@@ -13,14 +13,19 @@ if __name__ == "__main__":
 
     # === Simple ANN ===
     #run_simpleann("SimpleANN", 1, data_2nd_stage, evall_N_seq, device)
-    run_simpleann("SimpleANN", 5, data_2nd_stage, evall_N_seq, device)
+    #run_simpleann("SimpleANN",2, data_2nd_stage, evall_N_seq, device)
+    #run_simpleann("SimpleANN",3, data_2nd_stage, evall_N_seq, device)
+    #run_simpleann("SimpleANN", 4, data_2nd_stage, evall_N_seq, device)
+    #run_simpleann("SimpleANN", 5, data_2nd_stage, evall_N_seq, device)
     #run_simpleann("SimpleANN", 10, data_2nd_stage, evall_N_seq, device)
+    run_simpleann("SimpleANN", 20, data_2nd_stage, evall_N_seq, device)
 
 
  # === half RANN ===
     #run_half_RANN("half_RANN", 1, data_2nd_stage, evall_N_seq, device)
     #run_half_RANN("half_RANN", 5, data_2nd_stage, evall_N_seq, device)
     #run_half_RANN("half_RANN", 10, data_2nd_stage, evall_N_seq, device)
+    #run_half_RANN("half_RANN", 20, data_2nd_stage, evall_N_seq, device)
 
 
 # === half RRF ===

### 2025-09-21 Writing/manuscript_CF_continuous.qmd
working on writing

--- 2025-08-28 Writing/manuscript_CF_continuous.qmd
@@
 + Summary of findings
+
+**Widespread use of machine learning methods:**
+
++ ML is being widely used because it captures complex interactions better than traditional models, helping improve yield prediction and site-specific decisions.
+
+The rapid expansion of data from precision agriculture technologies has encouraged widespread adoption of machine learning (ML) methods in agricultural economics and crop management research.
+Traditional linear and quadratic models often fall short in capturing the nonlinear and spatially complex relationships between crop yields, management inputs, and environmental variables, which has led to a shift toward more flexible ML approaches.
+For example, random forest regression has been applied to optimize variable planting rates for corn and soybean, showing the ability to handle high-dimensional interactions between soil, topography, and weather while often performing better than linear models (@krause2020random).
+Similarly, convolutional neural networks (CNNs) have been developed to model crop yield response to seed and nitrogen management, making use of spatial features in field data and achieving substantial gains in predictive accuracy over conventional models (@barbosa2020modeling).
+More recent work has extended these deep learning approaches to incorporate ensemble methods, further improving prediction reliability and practical applicability in farm-level decision support (@barbosa2020risk).
+In addition, on-farm precision experimentation combined with ML-based yield response estimation has provided tools for empirically evaluating the economic performance of site-specific input recommendations, highlighting the potential of these methods for guiding practical decision-making (@gardner2021economic).
+Taken together, these studies point to the growing role of machine learning in advancing both predictive modeling and economic optimization in agriculture, marking a shift in how site-specific management is studied and implemented.
+
+**Causal machine learning becoming more and more imporatnt**
+
++ prediction alone is not enough, which is why causal ML is becoming increasingly important.
+
+While prediction-oriented machine learning methods have advanced site-specific management by improving yield modeling and decision support, there is a growing recognition that accurate prediction alone is not sufficient.
+For farm-level recommendations, the central question is not just how well a model fits observed yields, but how input changes cause yield outcomes.
+This shift in focus has increased the importance of causal machine learning methods, which are specifically designed to estimate heterogeneous treatment effects rather than simply predict yield levels.
+In agricultural economics and precision agriculture, these methods are becoming increasingly relevant as researchers look for tools that provide more reliable directions for optimizing input use across diverse field conditions.
+
+**example using CF**
+
++ In this paragraph I talked about an example using CF.
+Their work was limited to discrete treatments, but as more  data become available, N become continuous.
+Since CF is strong, we need to adapt it for continuous treatments.
+
+One prominent example of this trend is the causal forest (@athey2019generalized), which has become a valuable tool in the context of On-Farm Precision Experimentation (OFPE).
+Unlike prediction-based methods, causal forest is designed to non-parametrically identify heterogeneous treatment effects, making it well suited for revealing complex interactions among variables when the underlying functional forms are unknown.
+@kakimoto2022causal showed that it can estimate heterogeneous treatment effects of inputs on yield more accurately than prediction-based methods such as random forest, boosted forest, or convolutional neural networks.
+Their study, however, was limited to a single-year experiment on one field.
+Causal forest was originally developed for discrete treatments, and OFPE trials often include five or six distinct experimental input levels.
+In @kakimoto2022causal, the method estimated the effect of increasing nitrogen rates relative to the lowest rate, resulting in multiple treatment contrasts within the trial.
+As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights.
+But pooling brings together trials with different target rates, so applied nitrogen covers a near-continuous range, which creates a need for methods that handle continuous treatments.
+Using data across many fields offers clear advantages, and it becomes especially important to adapt causal forest methods to continuous treatments without losing their statistical strengths over other machine-learning approaches.
+
+**Objective**
+
++ the objective is to extend CF to continuous treatments, so we can recover full yield response functions and test performance.
+
+**Q**
+
+We also have SO. How to talk about it as an objective here?
+
+The objective of this study is to develop and evaluate a causal machine learning approach that can recover the full yield response function from on-farm precision experimentation data, where treatment variables such as nitrogen rates are continuous and randomized.
+Existing methods, including random forests, boosted forests, and neural networks, often fail to capture heterogeneous treatment effects of continuous treatments, while most causal machine learning frameworks have been designed for binary treatments.
+To address this gap, we extend the causal forest framework to continuous treatment settings by incorporating smoothing splines, allowing the model to non-parametrically capture complex interactions among covariates without imposing restrictive functional form assumptions.
+We assess the performance of the proposed method through Monte Carlo simulations, comparing it with commonly used machine learning approaches, and report results across scenarios using data from one, five, and ten fields.
 
@@
 
-In the on farm precision experiment (OFPE) setting, the ideas behind DML have shown promise for causal effect estimation.
+In the OFPE setting, the ideas behind DML have shown promise for causal effect estimation.
 @kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield, but their implementation was limited to a single-year experiment from one field and treated heterogeneity in a discrete way.

### 2025-09-28 codes/R/CF_YRF.qmd
CF YRF

---
title: "CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)
library(pasim)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

#--- python functions ---#
# use_condaenv("cf_conda311", required = TRUE)
# Apple M2
use_condaenv("cf_conda311",
             conda = "/Users/monamousavi/miniforge3/bin/conda",
             required = TRUE)
py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")

#--- R functions ---#
source(here("codes", "R", "Utility.R"))

#--- python functions ---#
source_python(here::here("codes/Python/run_CF_c.py"))
```

## load data

```{r}
#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 20

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

progress_csv <- file.path(output_dir, "progress.csv")

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────

train_test_splits <- read_csv(splits_file)

```

# CF YRF

```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(nrow(train_test_splits)))
for (i in seq_len(1)){
#for (i in seq_len(nrow(train_test_splits))) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch(
    {
      ## 1) train/test IDs
      split_row     <- train_test_splits[i, ]
      train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
      sim_id_ls     <- c(train_sim_ids, test_sim_id)

      ## 2) train & test data
      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      train_data <- data[sim != test_sim_id, ]
      test_data  <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      ## 3) CF 
      x_vars <- c("Nk", "plateau", "b0")
      T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
      Y      <- train_data[, yield]
      X      <- as.matrix(train_data[, ..x_vars])
      W      <- X
      X_test <- as.matrix(test_data[, ..x_vars])

      ## 4) run cf in python
      Sys.setenv(
        OMP_NUM_THREADS      = "1",
        MKL_NUM_THREADS      = "1",
        KMP_AFFINITY         = "disabled",
        KMP_DUPLICATE_LIB_OK = "TRUE"
      )

      te_hat_cf <- run_CF_c_py(
        Y,
        T_info$T_sp,
        X,
        W,
        n_estimators = 2000
      )
      te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

      ## 5) Response curves
      T_seq <- train_data[, seq(min(N), max(N), length = 100)]
      response_data <- find_response_semi(T_seq, T_info, te_info)

out_dir <- here::here(
  "results",
  "yield_response_function_for_one_iteration"
)
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

saveRDS(
  list(
    te_hat_cf     = te_hat_cf,
    te_info       = te_info,
    T_seq         = T_seq,
    response_data = response_data,
    test_data     = test_data,
    train_data    = train_data
  ),
  file = file.path(out_dir, sprintf("YRF_CF_%dfields.rds", n_fields))
)

})

}

```

# Post-estimation analysis
```{r}
cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_10fields.rds"))
```

## Visualization of yield response curve fit

```{r}
test_data <- cf_results$test_data
train_data <- cf_results$train_data
repsponse_data <- cf_results$repsponse_data

true_response <-
  test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
  .[, ] %>%
  expand_grid_df(., data.table(N = train_data[, seq(min(N), max(N), length = 100)])) %>%
  .[, yield := gen_yield_QP(b0, b1, b2, Nk, N)]

check_id <- test_data[order(Nk), aunit_id] %>%
  .[seq(1, nrow(test_data), length = 5)] %>%
  .[-1]

cf_curves <-
  response_data[aunit_id %in% check_id, ] %>% .[, type := "Estiamated by CF"]

true_curves <-
  true_response[aunit_id %in% check_id, ] %>%
  .[, type := "True"] %>%
  .[, .(aunit_id, N, yield, type)] %>%
  setnames("N", "T")

yield_intercept <- true_curves[, .(yield = mean(yield)), by = aunit_id]

yield_adjustment <-
  cf_curves[, .(est = mean(est)), by = aunit_id] %>%
  yield_intercept[., on = "aunit_id"] %>%
  .[, yield_adj := yield - est] %>%
  .[, .(aunit_id, yield_adj)]

cf_curves <-
  yield_adjustment[cf_curves, on = "aunit_id"] %>%
  .[, yield := est + yield_adj] %>%
  .[, .(aunit_id, yield, T, type)]

curves_data_il <-
  rbind(cf_curves, true_curves) %>%
  .[, type := factor(type, levels = c("True", "Estiamated by CF"))] %>%
  .[, group := .GRP, by = aunit_id]

g_response_estimated <-
  ggplot() +
  geom_line(
    data = curves_data_il[T <= 260, ],
    aes(
      y = yield,
      x = T,
      color = factor(group),
      linetype = type
    )
  ) +
  scale_color_discrete(name = "aunit_id Number") +
  scale_linetype_discrete(name = "Curve Type") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Corn Yield (bu/acre)") +
  xlab("Nitrogen Application Rate (lb/acre)")

# saveRDS(
#   object = g_response_estimated,
#   file   = here::here("Results", "g_response_estimated.rds")
# )

```

# All YRF

```{r}
#--- libraries ---#
library(pasim)
library(data.table)
library(dplyr)
library(ggplot2)
library(here)
library(purrr)
library(tidyr)     
library(patchwork)

process_cf_file <- function(file_path, label) {
  cf_results <- readRDS(file_path)

  test_data     <- cf_results$test_data
  train_data    <- cf_results$train_data
  response_data <- cf_results$response_data

  # true response curves
  true_response <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
    tidyr::expand_grid(N = seq(min(train_data$N),
                               max(train_data$N),
                               length = 100)) %>%
    as.data.table() %>%
    .[, yield := gen_yield_QP(b0, b1, b2, Nk, N)]

  # pick a few ids
  check_id <- test_data[order(Nk), aunit_id] %>%
    .[seq(1, nrow(test_data), length = 5)] %>%
    .[-1]

  # estimated by CF
  cf_curves <- response_data[aunit_id %in% check_id, ] %>%
    .[, type := "Estimated by CF"]

  # true curves
  true_curves <- true_response[aunit_id %in% check_id, ] %>%
    .[, type := "True"] %>%
    .[, .(aunit_id, N, yield, type)] %>%
    setnames("N", "T")

  # adjust intercepts
  yield_intercept <- true_curves[, .(yield = mean(yield)), by = aunit_id]
  yield_adjustment <- cf_curves[, .(est = mean(est)), by = aunit_id] %>%
    yield_intercept[., on = "aunit_id"] %>%
    .[, yield_adj := yield - est] %>%
    .[, .(aunit_id, yield_adj)]

  cf_curves <- yield_adjustment[cf_curves, on = "aunit_id"] %>%
    .[, yield := est + yield_adj] %>%
    .[, .(aunit_id, yield, T, type)]

  # combine
  curves_data_il <- rbind(cf_curves, true_curves) %>%
    .[, type := factor(type, levels = c("True", "Estimated by CF"))] %>%
    .[, group := .GRP, by = aunit_id]

  # plot
  ggplot() +
    geom_line(
      data = curves_data_il[T <= 260, ],
      aes(y = yield, x = T, color = factor(group), linetype = type)
    ) +
    scale_color_discrete(name = "aunit_id Number") +
    scale_linetype_discrete(name = "Curve Type") +
    theme_bw() +
    theme(legend.position = "bottom") +
    ylab("Corn Yield (bu/acre)") +
    xlab("Nitrogen Application Rate (lb/acre)") +
    ggtitle(paste("Yield Response Function -", label))
}

file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
file_paths <- here::here(
  "Results/yield_response_function_for_one_iteration",
  paste0("YRF_CF_", file_labels, ".rds")
)

#---all plots ---#
all_plots <- map2(file_paths, file_labels, process_cf_file)

# saveRDS(
#   all_plots,
#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds")
# )

```

# Visualize YRF

```{r}

all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds"))

# Example one plot
#all_plots[[1]]

# arrange 
wrap_plots(all_plots, ncol = 2)

```

### 2025-09-30 codes/R/visualized_results.qmd
Codes modified (#fields added)

@@
 
-# ---- Constants: all 18 result folders ----
-MODEL_DIRS <- c(
-  "CF_outcome_1_fields","CF_outcome_5_fields","CF_outcome_10_fields",
-  "RANN_outcome_one_field","RANN_outcome_five_fields","RANN_outcome_ten_fields",
-  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_two_fields","Simple_ANN_outcome_three_fields", "Simple_ANN_outcome_four_fields",
-  "Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields","Simple_ANN_outcome_twenty_fields",
-  "NO_RF_one_field","NO_RF_five_fields","NO_RF_ten_fields",
-  "HalfRRF_one_field","HalfRRF_five_fields","HalfRRF_ten_fields",
-  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields", "half_RANN_outcome_twenty_fields"
-)
-
 EXPECTED_ROWS <- 1440L
 
-# ---- function ----
+# --------
 rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))
 
+# number of fields 
 parse_n_fields <- function(folder) {
+  f <- str_to_lower(folder)
+
+  num <- str_match(f, "(_|\\b)(\\d+)_?fields?\\b")[,3]
+  if (!is.na(num)) return(as.integer(num))
+
+  words <- c("one"=1,"two"=2,"three"=3,"four"=4,"five"=5,
+             "ten"=10,"twenty"=20,"thirty"=30,"forty"=40)
+  word <- str_match(f, "(one|two|three|four|five|ten|twenty|thirty|forty)_?fields?")[,2]
+  if (!is.na(word)) return(words[[word]])
+
+  NA_integer_
+}
+
+# --- Model name mapping ---
+parse_model <- function(folder) {
+  f <- str_to_lower(folder)
+
   case_when(
-    str_detect(folder, "(?:_1_fields|one_field)\\b")      ~ 1L,
-    str_detect(folder, "(?:_2_fields|two_fields?)\\b")    ~ 2L,
-    str_detect(folder, "(?:_3_fields|three_fields?)\\b")  ~ 3L,
-    str_detect(folder, "(?:_4_fields|four_fields?)\\b")   ~ 4L,
-    str_detect(folder, "(?:_5_fields|five_fields?)\\b")   ~ 5L,
-    str_detect(folder, "(?:_10_fields|ten_fields?)\\b")   ~ 10L,
-    str_detect(folder, "(?:_20_fields|twenty_fields?)\\b")~ 20L,
-    TRUE ~ NA_integer_
+    str_detect(f, "^cf")          ~ "CF",
+    str_detect(f, "^rann")        ~ "DO_ANN",
+    str_detect(f, "^half_rann")   ~ "SO_ANN",
+    str_detect(f, "^simple_ann")  ~ "No_ANN",
+    str_detect(f, "^no_rf")       ~ "NO_RF",
+    str_detect(f, "^halfrrf")     ~ "SO_RF",
+    TRUE                          ~ NA_character_
   )
@@
 
-parse_model <- function(folder) {
-  f <- str_to_lower(folder)
-  case_when(
-    str_starts(f, "cf_")             ~ "CF",
-    str_starts(f, "rann_outcome")    ~ "RANN",
-    str_starts(f, "half_rann")       ~ "HalfRANN",
-    str_starts(f, "simple_ann")      ~ "Simple_ANN",
-    str_starts(f, "no_rf")           ~ "NO_RF",
-    str_starts(f, "halfrrf")         ~ "HalfRRF",
-    TRUE                             ~ folder
-  )
-}
-
-# Read any object: RDS, CSV
+# read RDS, CSV
 read_any <- function(f) {
@@
 
-# Standardize to two numeric vectors (per-sim rows)
 read_eonr_file <- function(f) {
@@
 
-# 
+# list files in dir 
 files_in_dir <- function(dir_path) {
   dir_ls(dir_path, type = "file", recurse = TRUE) %>%
-    # Names like sim_055_eonr or EONR_1 
     keep(~ str_detect(path_file(.x),
@@
                             ignore_case = TRUE))) %>%
-    # drop anything starting with "validation"
     discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
-    # drop macOS Icon files
     discard(~ grepl("Icon", basename(.x), fixed = TRUE))
@@
 
-# get sim id from filename base
+# extract sim id from filename
 extract_sim_id <- function(base) {
@@
 
-# 
 row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
@@
 
-# ---- create table  ----
+# --- RMSE table ----
 build_rmse_table <- function(base_dir = here("results"),
-                             include_dirs = MODEL_DIRS,
                              expected_rows = EXPECTED_ROWS) {
-  dirs <- file.path(base_dir, include_dirs)
-  dirs <- dirs[dir_exists(dirs)]
+
+  dirs <- fs::dir_ls(base_dir, type = "directory", recurse = FALSE) %>%
+    keep(~ str_detect(basename(.x),
+                      regex("^(cf|rann|half_rann|simple_ann|no_rf|halfrrf)",
+                            ignore_case = TRUE)))
 
@@
   }) %>%
-    filter(!is.na(sim), !is.na(`#field`), !is.na(RMSE)) %>%
+    filter(!is.na(Model), !is.na(`#field`), !is.na(RMSE)) %>%
     arrange(Model, `#field`, sim)
@@
   base_dir = here("results"),
-  include_dirs = MODEL_DIRS,
   expected_rows = EXPECTED_ROWS
@@
 
-#  model names: row labels in the graph
-model_to_row <- c(
+# -- labels for models ---
+model_names <- function(models) {
+  map <- c(
     "Simple_ANN" = "No_ANN",
@@
     "NO_RF"      = "NO_RF",
-  "HalfRANN" = "SO_ANN",
-  "HalfRRF"  = "SO_RF",
+    "HALFRANN"   = "SO_ANN",
+    "HALFRRF"    = "SO_RF",
     "CF"         = "CF"
   )
+  # use mapping if available, else leave unchanged
+  out <- unname(map[models])
+  ifelse(is.na(out), models, out)
+}
 
+# --- plotting data ---
 rmse_plot_df <- rmse_tbl_raw %>%
-  filter(ok_rows, Model %in% names(model_to_row)) %>%
+  filter(ok_rows) %>%
   mutate(
-    row_lab = recode(Model, !!!model_to_row),
-    col_lab = factor(`#field`,
-                     levels = c(1, 2, 3, 4, 5, 10, 20),
-                     labels = c("Number of fields: 1",
-                                "Number of fields: 2",
-                                "Number of fields: 3",
-                                "Number of fields: 4",
-                                "Number of fields: 5",
-                                "Number of fields: 10",
-                                "Number of fields: 20"))
+    row_lab = model_names(Model),
+    col_lab = factor(
+      `#field`,
+      levels = sort(unique(`#field`)),   # discover field counts
+      labels = paste0("Number of fields: ", sort(unique(`#field`)))
+    )
   )
 
-#  check 
+# --- Check ---
 rmse_plot_df %>% count(row_lab, col_lab)
 
-# --- setup ---
+# ------
 xmin <- 0
@@
   facet_grid(row_lab ~ col_lab, scales = "fixed") +  
-  scale_x_continuous(limits = c(xmin, xmax),
+  scale_x_continuous(
+    limits = c(xmin, xmax),
     breaks = seq(xmin, xmax, length.out = 6),
-                     expand = c(0, 0)) +
+    expand = c(0, 0)
+  ) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
@@
 p
-
 # ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)
@@
 
-# 
+# --- Prepare rmse_df ---
 rmse_df <- rmse_tbl_raw %>%
@@
   dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
-  # Rename models
   dplyr::mutate(
-    Model = dplyr::recode(
-      Model,
-      "Simple_ANN" = "NO_ANN",
-      "HalfRANN"   = "SO_ANN",
-      "RANN"       = "DO_ANN",
-      "HalfRRF"    = "SO_RF",
-      "CF"         = "CF",
-      "NO_RF"      = "NO_RF",
-      .default     = Model
-    ),
-    field_f = factor(`#field`, levels = c(1, 2, 3, 4, 5, 10, 20),
-                     labels = c("1 field", "2 fields", "3 fields", "4 fields",  "5 fields", "10 fields", "20 fields"))
+    # Field labels (auto from unique values)
+    field_f = factor(
+      `#field`,
+      levels = sort(unique(`#field`)),
+      labels = paste0(sort(unique(`#field`)),
+                      ifelse(sort(unique(`#field`)) == 1, " field", " fields"))
+    )
   )
 
-# 3) summary stats 
+# --- Summary stats ---
 rmse_summary <- rmse_df %>%
@@
     .groups = "drop"
-  )
+  ) %>%
+  # remove 2-field and 4-field combinations
+  dplyr::filter(!field_f %in% c("2 fields", "4 fields"))
 
@@
 
-# Best model by order
+# Best model by the numbers of fields
 
@@
   tab_header(
-    title = "Best Models by Field Count",
+    title = "best models by #field",
     subtitle = "Ordered by lowest mean RMSE"
@@
   cols_label(
-    field_f = "# Fields",
+    field_f = "# fields",
     mean    = "mean RMSE"

### 2025-09-30 codes/R/CF_Loop.qmd
CF loop (reproducable)
Adjust for python settings

@@
 use_condaenv("cf_conda311", required = TRUE)
+
+# use_condaenv("cf_conda311",
+#              conda = "/Users/monamousavi/miniforge3/bin/conda",
+#              required = TRUE)
+
 py_config()
@@
 
-## Data
+## load data
 
 ```{r}
-raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))
+#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))
 
 # how many fields to train on?
-n_fields <- 10
+n_fields <- 20
 
@@
       # Free ram 
-      rm(res, te_hat_cf, te_info, response_data, train_data, test_data, e)
+      rm(te_hat_cf, te_info, response_data, train_data, test_data, e)
       if (i %% 10 == 0) gc()

### 2025-10-08 codes/R/Corner_sol_CF.qmd
CF corner solution

---
title: "CF corner solution"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(data.table)
library(here)
library(tidyr)
```

## load data and find corner solutions
```{r}

# raw sim data
# raw_sim_data <- readRDS(here("data", "raw_sim_data.rds"))

# List of field settings you ran
n_fields_list <- c(1, 3, 5, 10, 20)

# Function to process one n_fields case
process_field_setting <- function(n_fields) {

  # train/test splits for this field count
  splits_file <- here("data", "train_test_split",
                      sprintf("train_test_splits_%dfields.csv", n_fields))
  splits <- fread(splits_file)

  # estimated EONR results for this field count
  eonr_dir <- here("results", sprintf("CF_outcome_%d_fields", n_fields))
  files <- list.files(eonr_dir, pattern = "_eonr\\.rds$", full.names = TRUE)
  eonr_all <- rbindlist(lapply(files, readRDS))

  # Loop on all test sims
  corner_results <- lapply(1:nrow(splits), function(i) {

    test_id <- splits$test_id[i]
    train_ids <- as.integer(unlist(splits[i, grep("^train_", names(splits)), with = FALSE]))

    # get all train sims used for this test (1, 3, 5, 10, 20)
    train_data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% train_ids, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    N_min <- min(train_data$N)
    N_max <- max(train_data$N)

    # get estimated opt_N_hat for this test sim
    eonr_sim <- eonr_all[sim == test_id]

    #  corner solutions
    eonr_sim[, corner := (opt_N_hat == N_min | opt_N_hat == N_max)]
    eonr_sim[, n_fields := n_fields]

    return(eonr_sim)
  })

  return(rbindlist(corner_results))
}

# run for all field counts
corner_all <- rbindlist(lapply(n_fields_list, process_field_setting))

# summarize: % of corner solutions by n_fields
corner_summary <- corner_all[, .(
  n_total = .N,
  n_corners = sum(corner, na.rm = TRUE)
), by = n_fields]

corner_summary

# save 
# saveRDS(corner_summary,
#         here("results", "corner_summary.rds"))

```

### 2025-10-13 codes/Python/Functions/CF_python.py
CF wrote entierly in python

@@ -1,0 +1,160 @@
+import os
+import numpy as np
+import pandas as pd
+from tqdm import tqdm
+from econml.dml import CausalForestDML
+from sklearn.ensemble import GradientBoostingRegressor
+from sklearn.multioutput import MultiOutputRegressor
+from sklearn.preprocessing import StandardScaler
+import subprocess
+import sys
+
+# === Ensure ipywidgets is installed (for tqdm / Jupyter progress bar) ===
+import subprocess
+import sys
+
+# === Ensure ipywidgets is installed (for tqdm / Jupyter progress bar) ===
+try:
+    import ipywidgets
+except ImportError:
+    print("Installing ipywidgets...")
+    subprocess.check_call([sys.executable, "-m", "pip", "install", "ipywidgets", "-q"])
+    # Try enabling the Jupyter widget extension, but skip if it fails (VS Code doesn’t need it)
+    try:
+        subprocess.check_call([sys.executable, "-m", "jupyter", "nbextension", "enable", "--py", "widgetsnbextension"])
+    except Exception as e:
+        print(f"(Skipping nbextension enable: {e})")
+
+
+
+
+# ================================================================
+# === MODEL: Causal Forest (Python / raw inputs)
+# ================================================================
+def train_cf_model_from_splits(data_2nd_stage, eval_N_seq, split_csv_path, folder_name):
+    """
+    Train EconML CausalForestDML on raw (non-orthogonalized) data
+    to match the R 'CF Loop' logic.
+    """
+
+    # --- Setup results folder ---
+    base_results_dir = os.path.join(os.getcwd(), "results")
+    output_folder = os.path.join(base_results_dir, folder_name)
+    os.makedirs(output_folder, exist_ok=True)
+
+    splits_df = pd.read_csv(split_csv_path)
+    p_corn = 6.25 / 25.4
+    p_N = 1 / 0.453592
+
+    os.environ["OMP_NUM_THREADS"] = "1"
+    os.environ["MKL_NUM_THREADS"] = "1"
+
+    for _, row in tqdm(splits_df.iterrows(), total=len(splits_df)):
+    #for _, row in tqdm(splits_df.iloc[:1].iterrows(), total=1):
+        test_id = int(row["test_id"])
+        train_ids = row[[c for c in row.index if c.startswith("train_")]].values
+
+        # --- Split train/test ---
+        train_df = data_2nd_stage[data_2nd_stage["sim"].isin(train_ids)].reset_index(drop=True)
[…]

### 2025-10-13 codes/Python/Functions/main_simulation_run.py
minor modification

@@ -3,11 +3,14 @@ from simple_ANN import run_model as run_simpleann
 from half_RANN import run_model as run_half_RANN
 from half_RRF import run_model as run_halfrrf
 from NO_RF import run_model as run_NO_RF
+from CF_python import run_model as run_CF_python
+from CF_savedT import run_model as run_CF_savedT
 from Generate_train_test_splits import data_2nd_stage, evall_N_seq, device
 
 if __name__ == "__main__":
     # === RANN ===
     #run_rann("RANN", 1, data_2nd_stage, evall_N_seq, device)
+    #run_rann("RANN", 3, data_2nd_stage, evall_N_seq, device)
     #run_rann("RANN", 5, data_2nd_stage, evall_N_seq, device)
     #run_rann("RANN", 10, data_2nd_stage, evall_N_seq, device)
 
@@ -18,11 +21,12 @@ if __name__ == "__main__":
     #run_simpleann("SimpleANN", 4, data_2nd_stage, evall_N_seq, device)
     #run_simpleann("SimpleANN", 5, data_2nd_stage, evall_N_seq, device)
     #run_simpleann("SimpleANN", 10, data_2nd_stage, evall_N_seq, device)
-    run_simpleann("SimpleANN", 20, data_2nd_stage, evall_N_seq, device)
+    #run_simpleann("SimpleANN", 20, data_2nd_stage, evall_N_seq, device)
 
 
  # === half RANN ===
     #run_half_RANN("half_RANN", 1, data_2nd_stage, evall_N_seq, device)
+    #run_half_RANN("half_RANN", 3, data_2nd_stage, evall_N_seq, device)
     #run_half_RANN("half_RANN", 5, data_2nd_stage, evall_N_seq, device)
     #run_half_RANN("half_RANN", 10, data_2nd_stage, evall_N_seq, device)
     #run_half_RANN("half_RANN", 20, data_2nd_stage, evall_N_seq, device)
@@ -30,14 +34,30 @@ if __name__ == "__main__":
 
 # === half RRF ===
     #run_halfrrf("HalfRRF", 1, data_2nd_stage, evall_N_seq, device)
+    #run_halfrrf("HalfRRF", 3, data_2nd_stage, evall_N_seq, device)
     #run_halfrrf("HalfRRF", 5, data_2nd_stage, evall_N_seq, device)
[…]

### 2025-10-13 codes/R/visualized_results.qmd
CF YRF added
modified to include CF python

@@
 
-# --- Model name mapping ---
+# --- model name mapping ---
 parse_model <- function(folder) {
@@
   case_when(
+    str_detect(f, "^cf_python")     ~ "CF_Python",
     str_detect(f, "^cf")            ~ "CF",
@@
   obj <- read_any(f)
-  if (is.list(obj) && all(c("pred","true") %in% names(obj))) {
-    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
+  if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
+    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF or CF_python csv
+  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
+    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))
   } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
     list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
-  } else if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
-    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF
-  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
-    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF list
   } else {
@@
     keep(~ str_detect(basename(.x),
-                      regex("^(cf|rann|half_rann|simple_ann|no_rf|halfrrf)",
+                      regex("^(cf|cf_python|rann|half_rann|simple_ann|no_rf|halfrrf)",
                             ignore_case = TRUE)))
@@
 )
-
 # save
@@
 
-# -- labels for models ---
+# -- label mapping for models ---
 model_names <- function(models) {
@@
     "RANN"       = "DO_ANN",
-    "NO_RF"      = "NO_RF",
     "HALFRANN"   = "SO_ANN",
     "HALFRRF"    = "SO_RF",
-    "CF"         = "CF"
+    "NO_RF"      = "NO_RF",
+    "CF"         = "CF",
+    "CF_Python"  = "CF_Python"
   )
-  # use mapping if available, else leave unchanged
-  out <- unname(map[models])
+
+  # normalize capitalization 
+  models_upper <- toupper(models)
+  out <- map[models_upper]
   ifelse(is.na(out), models, out)
@@
 
-# --- plotting data ---
+# ---  data for plotting ---
 rmse_plot_df <- rmse_tbl_raw %>%
@@
       `#field`,
-      levels = sort(unique(`#field`)),   # discover field counts
+      levels = sort(unique(`#field`)),
       labels = paste0("Number of fields: ", sort(unique(`#field`)))
@@
 
-# --- Check ---
+# ---number of sims per facet ---
 rmse_plot_df %>% count(row_lab, col_lab)
 
-# ------
+# --- axis ranges ---
 xmin <- 0
@@
 p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
-  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
+  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left", fill = "skyblue3") +
   facet_grid(row_lab ~ col_lab, scales = "fixed") +
@@
   scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  labs(x = "RMSE by iteration", y = "Count") +
+  labs(
+    title = "distribution of RMSE across models and field counts",
+    x = "RMSE by iteration",
+    y = "Count"
+  ) +
   theme_bw() +
   theme(
-    strip.text = element_text(face = "bold", size = 16),
-    axis.title = element_text(size = 14)
+    strip.text = element_text(face = "bold", size = 14),
+    axis.title = element_text(size = 13),
+    plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
   )
@@
 p
+
 # ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)
@@
 ```
+
+# YRF for CF
+
+```{r}
+all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds"))
+
+# Example one plot
+#all_plots[[1]]
+
+# arrange 
+wrap_plots(all_plots, ncol = 2)
+
+```

### 2025-10-14 codes/Python/Functions/CF_python.py
CF python minor modfications

@@ -9,11 +9,11 @@ from sklearn.preprocessing import StandardScaler
 import subprocess
 import sys
 
-# === Ensure ipywidgets is installed (for tqdm / Jupyter progress bar) ===
+# ===  ipywidgets is installed (for tqdm / Jupyter progress bar) ===
 import subprocess
 import sys
 
-# === Ensure ipywidgets is installed (for tqdm / Jupyter progress bar) ===
+# ===  ipywidgets is installed (for tqdm / Jupyter progress bar) ===
 try:
     import ipywidgets
 except ImportError:
@@ -34,7 +34,7 @@ except ImportError:
 def train_cf_model_from_splits(data_2nd_stage, eval_N_seq, split_csv_path, folder_name):
     """
     Train EconML CausalForestDML on raw (non-orthogonalized) data
-    to match the R 'CF Loop' logic.
+    
     """
 
     # --- Setup results folder ---

### 2025-10-14 codes/Python/Functions/NO_RF.py
Field counts added

@@ -79,8 +79,10 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):  # <- 
     split_csv_path = f'./data/train_test_split/train_test_splits_{n_fields}fields.csv'
     folder_map = {
         ("NO_RF", 1): "NO_RF_one_field",
+        ("NO_RF", 3): "NO_RF_three_fields",
         ("NO_RF", 5): "NO_RF_five_fields",
-        ("NO_RF", 10): "NO_RF_ten_fields"
+        ("NO_RF", 10): "NO_RF_ten_fields",
+        ("NO_RF", 20): "NO_RF_twenty_fields"
     }
     folder_name = folder_map.get((model_type, n_fields))
     if folder_name is None:

### 2025-10-14 codes/Python/Functions/RANN.py
field counts added

@@ -133,8 +133,10 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):
     split_csv_path = f'./data/train_test_split/train_test_splits_{n_fields}fields.csv'
     folder_map = {
         ("RANN", 1): "RANN_outcome_one_field",
+        ("RANN", 3): "RANN_outcome_three_fields",
         ("RANN", 5): "RANN_outcome_five_fields",
-        ("RANN", 10): "RANN_outcome_ten_fields"
+        ("RANN", 10): "RANN_outcome_ten_fields",
+        ("RANN", 20): "RANN_outcome_twenty_fields",
     }
     folder_name = folder_map.get((model_type, n_fields))
     if folder_name is None:

### 2025-10-14 codes/Python/Functions/half_RANN.py
three fields added

@@ -131,6 +131,7 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):
     split_csv_path = f'./data/train_test_split/train_test_splits_{n_fields}fields.csv'
     folder_map = {
         ("half_RANN", 1): "half_RANN_outcome_one_field",
+        ("half_RANN", 3): "half_RANN_outcome_three_fields",
         ("half_RANN", 5): "half_RANN_outcome_five_fields",
         ("half_RANN", 10): "half_RANN_outcome_ten_fields",
         ("half_RANN", 20): "half_RANN_outcome_twenty_fields" 

### 2025-10-14 codes/Python/Functions/half_RRF.py
counts for fields added

@@ -79,8 +79,10 @@ def run_model(model_type, n_fields, data_2nd_stage, evall_N_seq, device):  # <- 
     split_csv_path = f'./data/train_test_split/train_test_splits_{n_fields}fields.csv'
     folder_map = {
         ("HalfRRF", 1): "HalfRRF_one_field",
+        ("HalfRRF", 3): "HalfRRF_three_fields",
         ("HalfRRF", 5): "HalfRRF_five_fields",
-        ("HalfRRF", 10): "HalfRRF_ten_fields"
+        ("HalfRRF", 10): "HalfRRF_ten_fields",
+        ("HalfRRF", 20): "HalfRRF_twenty_fields"
     }
     folder_name = folder_map.get((model_type, n_fields))
     if folder_name is None:

### 2025-10-14 codes/Python/Functions/main_simulation_run.py
Additional CF checking (CF_savedT) but not necessary

@@ -59,5 +59,9 @@ if __name__ == "__main__":
  
 # === CF using saved spline columns (T_1, T_2, T_3, …) ===
     run_CF_savedT("CF_savedT", 1, data_2nd_stage, evall_N_seq)
+    #run_CF_savedT("CF_savedT", 3, data_2nd_stage, evall_N_seq)
+    #run_CF_savedT("CF_savedT", 5, data_2nd_stage, evall_N_seq)
+   #run_CF_savedT("CF_savedT", 10, data_2nd_stage, evall_N_seq)
+    #run_CF_savedT("CF_savedT", 20, data_2nd_stage, evall_N_seq)
 
   

### 2025-10-14 codes/Python/Functions/CF_savedT.py
new CF code in python but results are the same so not necessary

@@ -1,0 +1,142 @@
+# %% [markdown]
+# # === MODEL: CF_savedT ===
+
+# %%
+import os
+import numpy as np
+import pandas as pd
+from tqdm import tqdm
+from econml.dml import CausalForestDML
+from sklearn.ensemble import GradientBoostingRegressor
+from sklearn.multioutput import MultiOutputRegressor
+
+# %%
+def train_cf_savedT_model_from_splits(data_2nd_stage, evall_N_seq, split_csv_path, folder_name):
+    """
+    Train EconML CausalForestDML using saved spline columns (T_1_tilde, T_2_tilde, T_3_tilde)
+    """
+
+    # === Output setup ===
+    base_results_dir = os.path.join(os.getcwd(), "results")
+    output_folder = os.path.join(base_results_dir, folder_name)
+    os.makedirs(output_folder, exist_ok=True)
+
+    # Subfolders for TE and response matrices
+    te_dir = os.path.join(output_folder, "te_hat")
+    resp_dir = os.path.join(output_folder, "response")
+    eonr_dir = os.path.join(output_folder, "eonr")
+
+    os.makedirs(te_dir, exist_ok=True)
+    os.makedirs(resp_dir, exist_ok=True)
+    os.makedirs(eonr_dir, exist_ok=True)
+
+    # === Load split file ===
+    splits_df = pd.read_csv(split_csv_path)
+
+    # === Constants ===
+    p_corn = 6.25 / 25.4   # $/kg
+    p_N = 1 / 0.453592     # $/kg
+
+    os.environ["OMP_NUM_THREADS"] = "1"
+    os.environ["MKL_NUM_THREADS"] = "1"
+
+    # === Loop over each split row ===
+    #for _, row in tqdm(splits_df.iloc[:1].iterrows(), total=1):
+    for _, row in tqdm(splits_df.iterrows(), total=len(splits_df)):
+        test_id = int(row["test_id"])
+        train_ids = row[[c for c in row.index if c.startswith("train_")]].dropna().astype(int).values
+
+        # --- Split train/test data ---
+        train_df = data_2nd_stage[data_2nd_stage["sim"].isin(train_ids)].reset_index(drop=True)
+        test_df  = data_2nd_stage[data_2nd_stage["sim"] == test_id].reset_index(drop=True)
+
+        # --- Prepare variables ---
+        Y = train_df["yield"].to_numpy()
+        X = train_df[["Nk", "plateau", "b0"]].to_numpy()
+        W = X.copy()
[…]

### 2025-10-26 Writing/manuscript_CF_continuous.qmd
major update on writing plus introduction

--- 2025-08-28 Writing/manuscript_CF_continuous.qmd
@@
 format:
-  html: default
+  docx:
+    toc: true
+
 execute:
@@
 library(officer)
+library(data.table)
+library(tidyverse)
+library(officedown)
+library(officer)
+library(flextable)
+library(stringr)
+library(sf)
+library(lfe)
+library(modelsummary)
+library(patchwork)
+library(gridExtra)
+library(kableExtra)
+library(measurements)
 ```
@@
 
-+ Background
-+ Why we need the methods we are proposing? Motivation.
-+ Summary of findings
+Efficient nitrogen (N) management remains one of the most critical challenges in modern agricultural production.
+Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm.
+The principle of applying the right amount of nitrogen at the right place and time requires understanding the spatial and temporal variability in crop nitrogen needs.
+This variability arises from complex interactions among soil nitrogen availability, soil and field characteristics, and weather conditions.
+Accurately identifying the optimal nitrogen rate for each site and season is therefore fundamental to improving both profitability and sustainability in crop production systems.
+
+On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management.
+Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions.
+Analyzing these data enables estimation of yield response functions and identification of site-specific economically optimal nitrogen rates.
+Recent advances in data analytics—particularly the widespread adoption of machine learning (ML) methods—have expanded the potential of on-farm experimentation.
+Two major methodological directions have emerged: (1) yield prediction–based methods, which recommend site-specific input rates by predicting yield as a function of nitrogen and environmental covariates (@krause2020random; @barbosa2020modeling; @barbosa2020risk; @gardner2021economic), and (2) causal machine learning (causal ML) methods (@kakimoto2022causal), which estimate the underlying causal relationship between nitrogen application and yield response.
+
+In the OFPE setting, causal machine learning methods have shown to be promising to recover the full yield response curve.
+@kakimoto2022causal showed that causal forest can estimate heterogeneous treatment effects of an input on yield more accurately than prediction-oriented machined approaches like random forest, boosted forest, and convolutional neural network.
+However, their application of CF is only limited to a single-year OFPE experiment from a single field.
+Causal forest was developped primarily to estimate
+the heterogeneous treatment effects of a dicrete treatment. In OFPE, there are typically five
+or six distinctive target experimental rates (see the first panel of @fig-distribution-N).
+In @kakimoto2022causal, CF estimates the effect of multiple treatments defined as the increase in nitrogen rate from the lowest rate to the other rates (so n − 1 tretments in n-rate trials).
+However,
+once data from multiple fields with different target rates are combined, applied nitrogen rates
+become rather continuous as seen in the second panel of @fig-distribution-N.
+Existing approaches—such as direct applications of random forests or neural networks and CF—often tend to fail to capture heterogeneous treatment effects of a continuous variable (@kakimoto2022causal).
+
+```{r n-histogram, include = FALSE}
+# sim_data <-
+#   readRDS(here::here("data/raw_sim_data.rds")) %>%
+#    data.table::data.table()
+# 
+# reg_data_all <-
+#   sim_data[, .(design_name, reg_data)] %>%
+#   unnest(reg_data) %>%
+#   data.table::data.table() %>%
+#   .[, design := case_when(
+#     design_name == "Latin Square Fixed 5" ~ "ls5",
+#     design_name == "Latin Square Fixed 6" ~ "ls6"
+#   )]
+# 
+# data_guide <-
+#   lapply(
+#     1:nrow(reg_data_all),
+#     function(x) {
+#       temp_data <- reg_data_all[x, ]
+#       temp_sim <- temp_data[, sim]
+#       temp_design <- temp_data[, design]
+# 
+#       data_file_name <- paste0("data/sim_data_", temp_sim, "_design_", temp_design, ".rds")
+# 
+#       saveRDS(
+#         temp_data,
+#         here::here(data_file_name)
+#       )
+# 
+#       return(
+#         data.table::data.table(
+#           sim = temp_sim,
+#           design = temp_design,
+#           file_name = data_file_name
+#         )
+#       )
+#     }
+#   ) %>%
+#   data.table::rbindlist()
+# 
+# saveRDS(data_guide, here::here("data/data_guide.rds"))
+
+# sim_des_ls <- data_guide[, sim_des]
+# train_site_num <- 20
+# 
+# set.seed(87924354)
+# 
+# test_train <-
+#   data_guide %>%
+#   rowwise() %>%
+#   mutate(train_sim = list(
+#     sample(
+#       sim_des_ls[sim_des_ls != sim_des],
+#       train_site_num
+#     )
+#   )) %>%
+#   data.table::data.table()
+# 
+# saveRDS(test_train, here::here("data/test_train_ids.rds"))
+
+#####
+
+# data_guide <-
+#   readRDS(here::here("data/data_guide.rds")) %>%
+#   .[, sim_des := paste0(sim, "_", design)]
+# 
+# test_train <- readRDS(here::here("data/test_train_ids.rds"))
+# 
+# temp_data <- test_train[8, ]
+# 
+# test_sim_des <- temp_data$sim_des
+# 
+# #--- get the sim_des ids for training data ---#
+# train_sim_des <- temp_data$train_sim[[1]]
+# 
+# #--- train data ---#
+# train_data <-
+#   data_guide[sim_des %in% train_sim_des, file_name] %>%
+#   lapply(., function(x) readRDS(here::here(x))) %>%
+#   data.table::rbindlist() %>%
+#   .[, data] %>%
+#   data.table::rbindlist() %>%
+#   .[, type := "Multiple Fields"]
+# 
+# #--- test data ---#
+# test_data <-
+#   readRDS(here::here(temp_data$file_name))$data[[1]] %>%
+#   .[, type := "Single Field"]
+# 
+# data_illust <-
+#   rbind(test_data, train_data) %>%
+#   .[, type := factor(type, levels = c("Single Field", "Multiple Fields"))]
+# 
+# (
+#   g_N_hist_illust <-
+#     ggplot(data_illust) +
+#     geom_histogram(
+#       aes(x = N),
+#       bins = 50,
+#       fill = "gray",
+#       color = "blue"
+#     ) +
+#     facet_grid(type ~ ., scales = "free_y") +
+#     xlab("Nitrogen Application Rate (lb/acre)") +
+#     ylab("Count")
+# )
+
+# ggsave(
+#   g_N_hist_illust,
+#   file = here::here("writing/Figures/g_N_hist_single_multiple.pdf"),
+#   width = 6.5,
+#   height = 4
+# )
+
+```
+
+```{r fig-distribution-N, fig.cap="Distribution of nitrogen rates from a single OFPE and multiple OFPEs"}
+knitr::include_graphics(here::here("writing/Figures/g_N_hist_single_multiple.pdf"))
+```
+As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights.
+It creates a need for methods that handle continuous treatments and it becomes especially important to adapt causal methods to continuous treatments without losing their statistical strengths over other machine-learning approaches.
+
+The key objective of this article is to identify a causal machine learning approach capable of estimating heterogeneous treatment effects for yield response function estimation using data from on-farm trials, where the treatment variable—nitrogen rate—is continuous and randomized.
+We investigate two complementary approaches: a Single-Orthogonalized Neural Network (SO-ANN), designed to flexibly capture nonlinear treatment–response relationships, and an extended Causal Forest (CF) that incorporates smoothing splines to model continuous treatments.
+Both methods aim to preserve the strengths of causal ML—nonparametric modeling of complex interactions without restrictive assumptions.
+Their performance is evaluated through Monte Carlo simulations that mimic typical OFPE conditions and crop yield responses to nitrogen.
+
+We find that the SO-ANN approach outperforms the non-orthogonalized model, which represents the common practice in yield response function estimation.
+In contrast, the Double-Orthogonalized approach performs poorly, and its accuracy does not improve even with larger sample sizes, consistent with theoretical expectations.
+Our extended CF method successfully captures heterogeneous and nonlinear effects of a continuous treatment and accurately estimates site-specific profit-maximizing nitrogen rates with minimal bias underscoring the promise of this approach.
+Although the methodological extension proposed here is primarily motivated by the needs of OFPE, it holds potential for a wide range of applications in applied economics where continuous treatments and heterogeneous effects are of interest.
 
@@
 
-We run Monte Carlo simulation to test the performance of ...
-
-## Data generation
-
 ## Models and estimation procedures
-
-<span style = "color: blue;"> Use math here to explain orthogonalization.
-This is where you explain what SO and DO in general.
-The sections below add specifics to the individual approaches we take.
-You also need to provide contexts (past studies).
-</span>
 
@@
 
-In the on farm precision experiment (OFPE) setting, the ideas behind DML have shown promise for causal effect estimation.
-@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield, but their implementation was limited to a single-year experiment from one field and treated heterogeneity in a discrete way.
-They defined multiple “treatments” as increases from the lowest nitrogen rate to each higher rate, which is natural when there are only five or six fixed target rates per field.
-When data are pooled across fields with differing targets, however, applied nitrogen rates become effectively continuous, and the standard residualization of the treatment in DML does not extend cleanly to that setting.
+@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield and treated heterogeneity in a discrete way.
+However, when data are pooled across fields and applied nitrogen rates become continuous, the standard residualization of the treatment in DML does not extend cleanly to that setting.
 
@@
 For each simulation round, one field is selected as the test field.
-The remaining fields are used to construct the training set in three scenarios: (1) one-field training, (2) five-field training, and (3) ten-field training.
+The remaining fields are used to construct the training set in three scenarios: (1) one-field training, (2) three-field training, (3) five-field training, (4) ten-field training and (5) twenty-field training.
 These scenarios allow us to assess how model performance changes with the size and diversity of training data.
@@
 
+## Data generation
+
+First, a 92.17-acre rectangular field is created, which consists of $384$ of `r measurements::conv_unit(18.3, "m", "ft") %>% round(digits = 0)` feet $\times$ `r measurements::conv_unit(73.2, "m", "ft") %>% round(digits = 0)` feet plots.
+each of which will be assigned an N fertilizer application rate.
+Each plot is made up of four (4-rows $\times$ 1-column) `r measurements::conv_unit(18.3, "m", "ft") %>% round(digits = 0)` feet $\times$ `r measurements::conv_unit(18.3, "m", "ft") %>% round(digits = 0)` feet "subplots," which are the unit of analysis used in subsequent statistical analysis.
+Each subplot consists of a 6-rows $\times$ 6-columns grid of thirty-six `r measurements::conv_unit(3.05, "m", "ft") %>% round(digits = 0)` feet $\times$ `r measurements::conv_unit(3.05, "m", "ft") %>% round(digits = 0)` feet "cells."
+Visual representation of the field layout is presented in @fig-field-layout. The data are generated at the cell-level first and then aggregated up to analysis-unit level.
+
+```{r}
+
+# /*+++++++++++++++++++++++++++++++++++
+#' ## Plots
+# /*+++++++++++++++++++++++++++++++++++
+
+#library(patchwork)
+
+# all_data <- readRDS(here::here("data/raw_sim_data.rds"))
+# field_sf <- all_data$field_sf[[1]]
+# 
+# 
+# plot_sf <-
+#   field_sf %>%
+#   nest_by(plot_id) %>%
+#   mutate(data = list(
+#     st_union(data)
+#   )) %>%
+#   unnest() %>%
+#   st_as_sf()
+# 
+# plot_focus <- filter(plot_sf, plot_id == 283)
+# 
+# plot_focus_xy <-
+#   st_centroid(plot_focus) %>%
+#   st_coordinates()
+# 
+# bbox_field <- st_bbox(plot_sf)
+# min_x_field <- bbox_field["xmin"]
+# max_x_field <- bbox_field["xmax"]
+# center_x_field <- mean(c(min_x_field, max_x_field))
+# min_y_field <- bbox_field["ymin"]
+# max_y_field <- bbox_field["ymax"]
+# center_y_field <- mean(c(min_y_field, max_y_field))
+# 
+# g_field <-
+#   ggplot() +
+#   geom_sf(data = plot_sf, linewidth = 1, fill = NA) +
+#   geom_sf(data = plot_focus, fill = "red", alpha = 0.5) +
+#   #--- horizontal width ---#
+#   geom_segment(
+#     aes(x = center_x_field - 72, xend = min_x_field, y = max_y_field + 18, yend = max_y_field + 18),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   geom_segment(
+#     aes(x = center_x_field + 72, xend = max_x_field, y = max_y_field + 18, yend = max_y_field + 18),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   annotate(
+#     "text",
+#     x = center_x_field,
+#     y = max_y_field + 18,
+#     label = "2,835 feet"
+#   ) +
+#   #--- vertical length ---#
+#   geom_segment(
+#     aes(x = max_x_field + 18, xend = max_x_field + 18, y = center_y_field + 60, yend = max_y_field),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   geom_segment(
+#     aes(x = max_x_field + 18, xend = max_x_field + 18, y = center_y_field - 60, yend = min_y_field),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   annotate(
+#     "text",
+#     x = max_x_field + 18,
+#     y = center_y_field,
+#     label = "1,417 feet",
+#     angle = "270"
+#   ) +
+#   #--- expansion to the focus plot ---#
+#   geom_segment(
+#     aes(x = plot_focus_xy[1, "X"] - 36, xend = min_x_field, y = plot_focus_xy[1, "Y"] - 9, yend = plot_focus_xy[1, "Y"] - 120),
+#     color = "red"
+#   ) +
+#   geom_segment(
+#     aes(x = plot_focus_xy[1, "X"] + 36, xend = max_x_field, y = plot_focus_xy[1, "Y"] - 9, yend = plot_focus_xy[1, "Y"] - 120),
+#     color = "red"
+#   ) +
+#   annotate(
+#     "text",
+#     y = plot_focus_xy[1, "Y"],
+#     x = plot_focus_xy[1, "X"],
+#     label = "Plot",
+#     size = 3
+#   ) +
+#   theme_void()
+# 
+# # /*+++++++++++++++++++++++++++++++++++
+# #' ## inside a plot
+# # /*+++++++++++++++++++++++++++++++++++
+# plot_sf_focus <- filter(field_sf, plot_id == 1)
+# 
+# bbox_plot <- st_bbox(plot_sf_focus)
+# min_x_plot <- bbox_plot["xmin"]
+# max_x_plot <- bbox_plot["xmax"]
+# center_x_plot <- mean(c(min_x_plot, max_x_plot))
+# min_y_plot <- bbox_plot["ymin"]
+# max_y_plot <- bbox_plot["ymax"]
+# center_y_plot <- mean(c(min_y_plot, max_y_plot))
+# 
+# subplot_sf <-
+#   plot_sf_focus %>%
+#   nest_by(aunit_id, buffer) %>%
+#   mutate(data = list(
+#     st_union(data)
+#   )) %>%
+#   unnest() %>%
+#   st_as_sf() %>%
+#   mutate(label = ifelse(
+#     buffer == 1,
+#     "buffer",
+#     paste0("subplot-", aunit_id - 1)
+#   )) %>%
+#   mutate(buf_or_not = ifelse(label == "buffer", "buffer", "subplot"))
+# 
+# subplot_text_sf <- st_centroid(subplot_sf)
+# 
+# cell_sf <- plot_sf_focus[7, ]
+# 
+# bbox_cell <- st_bbox(cell_sf)
+# min_x_cell <- bbox_cell["xmin"]
+# max_x_cell <- bbox_cell["xmax"]
+# center_x_cell <- mean(c(min_x_cell, max_x_cell))
+# min_y_cell <- bbox_cell["ymin"]
+# max_y_cell <- bbox_cell["ymax"]
+# center_y_cell <- mean(c(min_y_cell, max_y_cell))
+# 
+# g_inside_plot <-
+#   ggplot() +
+#   geom_sf(data = plot_sf_focus, size = 0.2, fill = NA) +
+#   #--- horizontal width ---#
+#   geom_segment(
+#     aes(x = center_x_plot - 6, xend = min_x_plot, y = min_y_plot - 3, yend = min_y_plot - 3),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   geom_segment(
+#     aes(x = center_x_plot + 6, xend = max_x_plot, y = min_y_plot - 3, yend = min_y_plot - 3),
+#     arrow = arrow(length = unit(0.3, "cm"))
+#   ) +
+#   annotate(
+#     "text",
+#     x = center_x_plot,
+#     y = min_y_plot - 3,
+#     label = "236 feet"
+#   ) +
+#   geom_sf(data = filter(subplot_sf, buffer == 1), fill = "yellow", alpha = 0.4) +
+#   geom_sf(data = subplot_sf, size = 1.2, fill = NA, color = "blue") +
+#   geom_sf_text(data = subplot_text_sf, aes(label = label), size = 4, color = "blue") +
+#   geom_sf(data = cell_sf, fill = "green", alpha = 0.3) +
+#   #--- horizontal width of the cell ---#
+#   geom_segment(
+#     aes(x = center_x_cell, xend = min_x_cell, y = max_y_cell + 1, yend = max_y_cell + 1),
+#     arrow = arrow(length = unit(0.2, "cm"))
+#   ) +
+#   geom_segment(
+#     aes(x = center_x_cell, xend = max_x_cell, y = max_y_cell + 1, yend = max_y_cell + 1),
+#     arrow = arrow(length = unit(0.2, "cm"))
+#   ) +
+#   annotate(
+#     "text",
+#     x = center_x_cell,
+#     y = max_y_plot + 3,
+#     label = "19.7 feet"
+#   ) +
+#   annotate("text", x = 39, y = 429, label = "cell") +
+#   scale_fill_discrete(name = "") +
+#   theme_void() +
+#   theme(
+#     legend.position = "bottom",
+#     legend.text = element_text(size = 12)
+#   )
+# 
+# g_layout <- g_field / g_inside_plot
+# 
+# saveRDS(g_layout, here::here("writing/Figures/g_layout.rds"))
+# 
+# ggsave(
+#   g_layout,
+#   file = here::here("writing/Figures/g_layout.pdf"),
+#   width = 6.5,
+#   height = 6
+# )
+```
+
+```{r fig-field-layout, fig.cap = "Visualization of the field layout and components"}
+knitr::include_graphics(here::here("Writing/Figures/g_layout.pdf"))
+```
+Cell-level yields are generated following the quadratic-plateau functional form as follows:
+
+```{=latex}
+\begin{align} 
+y_{i,j} = 
+  \begin{cases}
+  \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
+  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
+  \end{cases}
+\label{eq:yield-gen-sim}
+\end{align}
+```
+
+In this formulation, yield increases as $N_{i,j}$ increase until $N_{i,j}$ reaches $\tau_{i,j}$, at which yield hits the plateau, $\eta_{i,j}$.
+The yield level at $N_{i,j}= 0$ is $\alpha_{i,j}$.
+The rate at which yield increases with respect to $N_{i,j}$ is governed by $\beta_{i,j}$ and $\gamma_{i,j}$.
+For a given cell, the quadratic-plateu function has five parameters, but the degree of freedom is only three.
+That is, once any three of the five parameters are determined, then the rest of the parameters are automatically determined.
+Specifically, we first generated $\alpha_{i,j}$, $\eta_{i,j}$, and $\tau_{i,j}$ in a spatially correlated manner following the Gaussian bluh bluh process (cite), and then $\beta_{i,j}$ and $\gamma_{i,j}$ were derived.
+The erorr term ($\varepsilon_{i,t}$) is generated using the same spatial process.
+This is repeated independenty 500 times to create 500 fields of the same size but different spatial distribution and values of the three parameters.
+
+Trial designs vary field by field based on the value of $\tau_{i,j}$.
+Specifically, the lowest nitrogen rate is set at the minimum of $\tau_{i,j}$ of the field minus 50, and the highest rate is set at the maximum of $\tau_{i,j}$ of the field plus 30.
+We then create a sequence of equi-distance 5-level nitrogen rates.
+Once the unique nitrogen levels are determined, a Latin square design is used as a trial design (see  @fig-trial-design) for an example trial design).
+This ensures that nitrogen rate is orthogonal to any variable including the error term, avoiding endogeneity problem as done in OFPE in real world.
+Now, yield values at the cell level can be calculated by plugging in the value of $N_{i,j}$, $\alpha_{i,j}$, $\eta_{i,j}$, $\tau_{i,j}$, and $\varepsilon_{i,j}$ into Equation~\ref{eq:yield-gen-sim}.
+
+```{r}
+# trial_design <-
+#   all_data$trial_design[[1]] %>%
+#   .[sim == 1, ]
+# 
+# field <- data.table(field_sf)
+# 
+# td_sf <-
+#   trial_design[field, on = c("block_id", "plot_in_block_id")] %>%
+#   st_as_sf()
+# 
+# g_td <-
+#   ggplot(td_sf) +
+#   geom_sf(aes(fill = factor(N_tgt)), color = NA) +
+#   scale_fill_viridis_d(name = "Nitrogen Rate (lb/acre)") +
+#   theme_void() +
+#   theme(legend.position = "bottom")
+# 
+# saveRDS(g_td, here::here("writing/Figures/g_td.rds"))
+```
+
+```{r fig-trial-design, fig.cap = "Example trial design"}
+readRDS(here::here("writing/Figures/g_td.rds"))
+```
+
 # Simulation Results
 
-We compare six estimators—CF, SO_ANN, DO_ANN, NO_ANN, SO_RF, and NO_RF—over 500 Monte-Carlo replications and three training sizes (1, 5, and 10 fields).
+We compare six estimators—CF, SO_ANN, DO_ANN, NO_ANN, SO_RF, and NO_RF—over 500 Monte-Carlo replications and five training sizes (1, 3, 5, 10 and 20 fields).
 Each replication contains 1,440 evaluation units.
@@
 
-```{r setup, include=FALSE}
+```{r sim-setup, include=FALSE}
 
@@
 rmse_tbl_raw <- readRDS(here::here("results", "rmse_tbl_raw.rds"))
+
 ```
@@
 #| fig-height: 7
+#| fig-show: hold
 
-# model names: row labels in the graph
-model_to_row <- c(
+library(dplyr)
+library(ggplot2)
+library(stringr)
+
+model_names <- function(models) {
+  map <- c(
     "Simple_ANN" = "No_ANN",
     "RANN"       = "DO_ANN",
+    "HALFRANN"   = "SO_ANN",
+    "HALFRRF"    = "SO_RF",
     "NO_RF"      = "NO_RF",
-  "HalfRANN"   = "SO_ANN",
-  "HalfRRF"    = "SO_RF",
     "CF"         = "CF"
   )
+  models_upper <- toupper(models)
+  out <- map[models_upper]
+  ifelse(is.na(out), models, out)
+}
 
 rmse_plot_df <- rmse_tbl_raw %>%
-  dplyr::filter(ok_rows, Model %in% names(model_to_row)) %>%
-  dplyr::mutate(
-    row_lab = dplyr::recode(Model, !!!model_to_row),
+  filter(ok_rows) %>%
+  mutate(
+    row_lab = model_names(Model),
     col_lab = factor(`#field`,
-                     levels = c(1, 5, 10),
-                     labels = c("Number of fields: 1",
-                                "Number of fields: 5",
-                                "Number of fields: 10"))
+      levels = sort(unique(`#field`)),
+      labels = paste0("Number of fields: ", sort(unique(`#field`)))
+    )
   )
 
-# consistent axes
+# prevent table output from creating phantom figure
+invisible(rmse_plot_df %>% count(row_lab, col_lab))
+
 xmin <- 0
@@
 p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
-  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
+  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left", fill = "skyblue3") +
   facet_grid(row_lab ~ col_lab, scales = "fixed") +
-  scale_x_continuous(limits = c(xmin, xmax),
-                     breaks = seq(xmin, xmax, length.out = 6),
-                     expand = c(0, 0)) +
+  scale_x_continuous(limits = c(xmin, xmax), breaks = seq(xmin, xmax, length.out = 6), expand = c(0, 0)) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  labs(x = "RMSE by iteration", y = "Count") +
+  labs(
+    #title = "Distribution of RMSE across models and field counts",
+    x = "RMSE by iteration",
+    y = "Count"
+  ) +
   theme_bw() +
   theme(
-    strip.text = element_text(face = "bold", size = 16),
-    axis.title = element_text(size = 14)
+    strip.text = element_text(face = "bold", size = 14),
+    axis.title = element_text(size = 13),
+    plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
   )
 
-p
+print(p)
+
 ```
 
-Mean RMSE by model and training size (@fig-avg-performance and @tbl-mean-rmse-heatmap) confirms these patterns.
+Mean RMSE by model and training size (@fig-avg-performance) confirms these patterns.
 CF is consistently low (≈15 across sizes).
@@
   dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
-  # Rename models
   dplyr::mutate(
-    Model = dplyr::recode(
-      Model,
-      "Simple_ANN" = "NO_ANN",
-      "HalfRANN"   = "SO_ANN",
-      "RANN"       = "DO_ANN",
-      "HalfRRF"    = "SO_RF",
-      "CF"         = "CF",
-      "NO_RF"      = "NO_RF",
-      .default     = Model
-    ),
-    field_f = factor(`#field`, levels = c(1, 5, 10),
-                     labels = c("1 field", "5 fields", "10 fields"))
+    # Field labels (auto from unique values)
+    field_f = factor(
+      `#field`,
+      levels = sort(unique(`#field`)),
+      labels = paste0(sort(unique(`#field`)),
+                      ifelse(sort(unique(`#field`)) == 1, " field", " fields"))
+    )
   )
@@
   )
-
 # 
@@
 ```{r}
-#| label: tbl-mean-rmse-heatmap
-#| tbl-cap: "Mean RMSE by model and number of fields."
-#| tbl-cap-location: bottom
-#| echo: false
-#| message: false
-#| warning: false
-
-library(dplyr)
-library(tidyr)
-library(gt)
-library(scales)
-
-# Range for coloring
-fill_min <- floor(min(rmse_summary$mean, na.rm = TRUE))
-fill_max <- ceiling(max(rmse_summary$mean, na.rm = TRUE))
-
-# Wide table of means
-rmse_table <- rmse_summary %>%
-  select(Model, field_f, mean) %>%
-  mutate(mean = round(mean, 1)) %>%
-  pivot_wider(names_from = field_f, values_from = mean) %>%
-  arrange(Model)
-
-# Light → mid-blue palette
-pal <- col_numeric(
-  palette = c("#f7fbff", "#6baed6"),
-  domain  = c(fill_min, fill_max)
-)
-
-# HTML colorbar legend to place under the table
-legend_html <- sprintf(
-  '<div style="display:flex;align-items:center;gap:8px;margin-top:6px;">
-     <span style="font-size:12px;color:#444;">Low (%.1f)</span>
-     <div style="flex:1;height:10px;background:linear-gradient(to right,#f7fbff,#6baed6);
-                 border:1px solid #ddd;border-radius:3px;"></div>
-     <span style="font-size:12px;color:#444;">High (%.1f)</span>
-   </div>', fill_min, fill_max)
-
-rmse_table %>%
-  gt(rowname_col = "Model") %>%
-  data_color(
-    columns = c(`1 field`, `5 fields`, `10 fields`),
-    colors  = pal
-  ) %>%
-  fmt_number(columns = c(`1 field`, `5 fields`, `10 fields`), decimals = 1) %>%
-  cols_label(
-    `1 field`  = "1 field",
-    `5 fields` = "5 fields",
-    `10 fields`= "10 fields"
-  ) %>%
-  tab_options(
-    table.font.size = px(13),
-    data_row.padding = px(6),
-    column_labels.font.weight = "bold"
-  ) %>%
-  tab_source_note(source_note = gt::html(legend_html))
+# #| label: tbl-mean-rmse-heatmap
+# #| tbl-cap: "Mean RMSE by model and number of fields."
+# #| tbl-cap-location: bottom
+# #| echo: false
+# #| message: false
+# #| warning: false
+# 
+# library(dplyr)
+# library(tidyr)
+# library(gt)
+# library(scales)
+# 
+# # Range for coloring
+# fill_min <- floor(min(rmse_summary$mean, na.rm = TRUE))
+# fill_max <- ceiling(max(rmse_summary$mean, na.rm = TRUE))
+# 
+# # Wide table of means
+# rmse_table <- rmse_summary %>%
+#   select(Model, field_f, mean) %>%
+#   mutate(mean = round(mean, 1)) %>%
+#   pivot_wider(names_from = field_f, values_from = mean) %>%
+#   arrange(Model)
+# 
+# # Light → mid-blue palette
+# pal <- col_numeric(
+#   palette = c("#f7fbff", "#6baed6"),
+#   domain  = c(fill_min, fill_max)
+# )
+# 
+# # HTML colorbar legend to place under the table
+# legend_html <- sprintf(
+#   '<div style="display:flex;align-items:center;gap:8px;margin-top:6px;">
+#      <span style="font-size:12px;color:#444;">Low (%.1f)</span>
+#      <div style="flex:1;height:10px;background:linear-gradient(to right,#f7fbff,#6baed6);
+#                  border:1px solid #ddd;border-radius:3px;"></div>
+#      <span style="font-size:12px;color:#444;">High (%.1f)</span>
+#    </div>', fill_min, fill_max)
+# 
+# rmse_table %>%
+#   gt(rowname_col = "Model") %>%
+#   data_color(
+#     columns = c(`1 field`, `5 fields`, `10 fields`),
+#     colors  = pal
+#   ) %>%
+#   fmt_number(columns = c(`1 field`, `5 fields`, `10 fields`), decimals = 1) %>%
+#   cols_label(
+#     `1 field`  = "1 field",
+#     `5 fields` = "5 fields",
+#     `10 fields`= "10 fields"
+#   ) %>%
+#   tab_options(
+#     table.font.size = px(13),
+#     data_row.padding = px(6),
+#     column_labels.font.weight = "bold"
+#   ) %>%
+#   tab_source_note(source_note = gt::html(legend_html))
 ```
@@
     ),
-    field_f = factor(`#field`, levels = c(1, 5, 10),
-                     labels = c("1 field", "5 fields", "10 fields"))
+    field_f = factor(`#field`, levels = c(1,3, 5, 10, 20),
+                     labels = c("1 field", "3 fields", "5 fields", "10 fields", "20 fields"))
   )
@@
 ```
+```{r}
 
-Predicted yield–N curves (@fig-yield-res-1 through @fig-yield-res-5) for five randomly selected analysis units clarify why methods differ.
+#| label: tbl-best_model_by_field_count
+#| tbl-cap: "Best-performing model by field count."
+#| tbl-cap-location: bottom
+#| echo: false
+#| message: false
+#| warning: false
+
+library(gt)
+library(dplyr)
+library(stringr)
+
+best_models <- rmse_summary %>%
+  mutate(field_n = as.numeric(str_extract(field_f, "\\d+"))) %>%
+  group_by(field_n) %>%
+  slice_min(order_by = mean, with_ties = FALSE) %>%
+  ungroup() %>%
+  arrange(mean, field_n) %>%
+  select(Model, field_f, mean)
+
+best_models %>%
+  gt() %>%
+  tab_header(
+    title = "Best models by #field",
+    subtitle = "Ordered by lowest mean RMSE"
+  ) %>%
+  fmt_number(columns = c(mean), decimals = 3) %>%
+  cols_label(
+    field_f = "# fields",
+    mean = "mean RMSE"
+  ) %>%
+  tab_options(table.font.size = "small")
+
+```
+
+Predicted yield–N curves (@fig-yield-res-1 through @fig-yield-res-3) for five randomly selected analysis units clarify why methods differ.
 SO_ANN produces agronomically plausible, monotone-concave responses with clear plateaus (ten-field panel), yielding interior optima and low EONR error.
@@
 ```{r}
+
 #| label: fig-yield-res-1
-#| fig-cap: "Yield response function (NO_ANN; 5 fields)."
-#| fig-cap-location: bottom
-#| fig-width: 12
-#| fig-height: 7
-#| message: false
-#| warning: false
+#| include: false
 
 source(here("codes", "R", "visualize_yield_response_function.R"))
-
-plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
-```
-```{r}
-#| label: fig-yield-res-2
-#| fig-cap: "Yield response function (DO_ANN; 5 fields)."
-#| fig-width: 12
-#| fig-height: 7
-
-source(here("codes", "R", "visualize_yield_response_function.R"))
-plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde
+#p1 <- plot_yield_response(model = "CF", n_fields = 1)
+p1 <- plot_yield_response("CF", 1, add_title = FALSE)
 
@@
 ```{r}
-#| label: fig-yield-res-3
-#| fig-cap: "Yield response function SO_ANN (SO_ANN; 10 fields)."
-#| fig-width: 12
-#| fig-height: 7
-#| 
+
+#| label: fig-yield-res-2
+#| include: false
+
 source(here("codes", "R", "visualize_yield_response_function.R"))
-plot_yield_response(model = "SO_ANN", n_fields = 10)
+#p2 <- plot_yield_response(model = "CF", n_fields = 3)
+p2 <- plot_yield_response("CF", 3, add_title = FALSE)
 
@@
 ```{r}
-#| label: fig-yield-res-4
-#| fig-cap: "Yield response function NO_RF (NO_RF; 5 fields)."
-#| fig-width: 12
-#| fig-height: 7
-#| 
+
+#| label: fig-yield-res-3
+#| include: false
+
 source(here("codes", "R", "visualize_yield_response_function.R"))
-plot_yield_response(model = "NO_RF", n_fields = 5)
+#p3 <- plot_yield_response(model = "CF", n_fields = 5)
+p3 <- plot_yield_response("CF", 5, add_title = FALSE)
 
@@
 ```{r}
+
+#| label: fig-yield-res-4
+#| include: false
+
+source(here("codes", "R", "visualize_yield_response_function.R"))
+#p4 <- plot_yield_response(model = "SO_ANN", n_fields = 10)
+p4 <- plot_yield_response("SO_ANN", 10, add_title = FALSE)
+
+```
+
+```{r}
+
 #| label: fig-yield-res-5
-#| fig-cap: "Yield response function SO_RF (SO_RF; 5 fields)."
-#| fig-width: 12
-#| fig-height: 7
-#| 
+#| include: false
+
 source(here("codes", "R", "visualize_yield_response_function.R"))
-plot_yield_response(model = "SO_RF", n_fields = 5) 
+#p5 <- plot_yield_response(model = "SO_ANN", n_fields = 20)
+p5 <- plot_yield_response("SO_ANN", 20, add_title = FALSE)
+
+```
+
+```{r}
+#| include: false
+
+library(patchwork)
+
+```
+
+```{r}
+# #| label: fig-yield-res-combined
+# #| fig-cap: "Yield response functions for CF and SO_ANN models across different field counts."
+# #| echo: false
+# #| fig-width: 10
+# #| fig-height: 8
+# 
+# (p1 | p2 | p3) /
+# (p4 | p5)
+```
+
+```{r}
+
+# #| label: fig-yrf
+# #| fig-cap: "Yield response function"
+# #| out.width: "100%"
+# #| fig-align: "center"
+# 
+# 
+# knitr::include_graphics(here::here("writing/Figures/YRF.pdf"))
+
+```
+
+```{r}
+#| label: fig-yield-res-combined
+#| fig-cap: "Yield response functions for CF and SO_ANN models across different field counts."
+#| echo: false
+#| fig-width: 10
+#| fig-height: 8
+
+library(patchwork)
+library(grid)
+
+# Add small text labels between rows
+title_CF <- grid::textGrob(
+  "Causal Forest (CF): 1, 3, and 5 fields",
+  gp = grid::gpar(fontsize = 13, fontface = "bold")
+)
+title_SO <- grid::textGrob(
+  "Single-Orthogonalized ANN (SO_ANN): 10 and 20 fields",
+  gp = grid::gpar(fontsize = 13, fontface = "bold")
+)
+
+# Combine with labeled rows
+(
+  wrap_elements(full = title_CF) /
+  (p1 | p2 | p3) /
+  wrap_elements(full = title_SO) /
+  (p4 | p5)
+) +
+  plot_annotation(
+    #title = "Yield response under different models and field counts",
+    theme = theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5))
+  )
+
 ```

### 2025-10-26 codes/R/visualize_yield_response_function.R
YRF modified

@@ -94,192 +94,247 @@
 
 
 
-#############$%^&&&&&&&&&&&&&&&&&&& add true yrf and mean match as well
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+#############$%^&&&&&&&&&&&&&&&&&&& Previous used ##############################
+# 
+# suppressPackageStartupMessages({
+#   library(here)
+#   library(readr)
+#   library(ggplot2)
+#   library(dplyr)
+#   library(tidyr)
+#   library(magrittr)  # %>%
+# })
+# 
+# #  folder label
+# # fields_label <- function(n_fields) {
+# #   if (n_fields == 1) "one_field"
+# #   else if (n_fields == 3) "three_fields"
+# #   else if (n_fields == 5) "five_fields"
+# #   else if (n_fields == 10) "ten_fields"
+# #   else if (n_fields == 20) "twenty_fields"
+# #   else paste0(n_fields, "_fields")
+# # }
+# 
+# fields_label <- function(n_fields, model = "SO_ANN") {
+#   # normalize input
+#   n_str <- tolower(as.character(n_fields))
+#   n_str <- stringr::str_replace_all(n_str, "_fields?$", "")
+# 
+#   # map word → numeric
+#   lookup <- c(
+#     "one" = 1, "1" = 1,
+#     "three" = 3, "3" = 3,
+#     "five" = 5, "5" = 5,
+#     "ten" = 10, "10" = 10,
+#     "twenty" = 20, "20" = 20
+#   )
+#   n_val <- lookup[[n_str]]
+#   if (is.null(n_val)) stop("Unknown n_fields value: ", n_fields, call. = FALSE)
+# 
+#   path1 <- paste0(n_val, "_fields")
+#   path2 <- switch(as.character(n_val),
+#                   "1"  = "one_field",
+#                   "3"  = "three_fields",
+#                   "5"  = "five_fields",
+#                   "10" = "ten_fields",
+#                   "20" = "twenty_fields")
+# 
+#   # --- check which folder actually exists (for any model) ---
[…]

### 2025-10-26 codes/R/visualized_results.qmd
modifications

@@
 
-models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF")
+models_to_compare <- c("NO_ANN", "DO_ANN", "SO_ANN", "NO_RF", "SO_RF", "CF")
 
@@
 plot_yield_response("NO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 3, global_models = models_to_compare)
 plot_yield_response("NO_ANN", 5, global_models = models_to_compare)
 plot_yield_response("NO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("NO_ANN", 20, global_models = models_to_compare)
 plot_yield_response("DO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 3, global_models = models_to_compare)
 plot_yield_response("DO_ANN", 5, global_models = models_to_compare)
 plot_yield_response("DO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("DO_ANN", 20, global_models = models_to_compare)
 plot_yield_response("SO_ANN", 1, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 3, global_models = models_to_compare)
 plot_yield_response("SO_ANN", 5, global_models = models_to_compare)
 plot_yield_response("SO_ANN", 10, global_models = models_to_compare)
+plot_yield_response("SO_ANN", 20, global_models = models_to_compare)
 plot_yield_response("NO_RF",  1, global_models = models_to_compare)
+plot_yield_response("NO_RF",  3, global_models = models_to_compare)
 plot_yield_response("NO_RF",  5, global_models = models_to_compare)
 plot_yield_response("NO_RF",  10, global_models = models_to_compare)
+plot_yield_response("NO_RF",  20, global_models = models_to_compare)
 plot_yield_response("SO_RF",  1, global_models = models_to_compare)
+plot_yield_response("SO_RF",  3, global_models = models_to_compare)
 plot_yield_response("SO_RF",  5, global_models = models_to_compare)
 plot_yield_response("SO_RF",  10, global_models = models_to_compare)
+plot_yield_response("SO_RF",  20, global_models = models_to_compare)
+plot_yield_response("CF", 1, global_models = models_to_compare)
+plot_yield_response("CF", 3, global_models = models_to_compare)
+plot_yield_response("CF", 5, global_models = models_to_compare)
+plot_yield_response("CF", 10, global_models = models_to_compare)
+plot_yield_response("CF", 20, global_models = models_to_compare)
+
 ```
@@
 
-# ---- Packages ----
 library(tidyverse)
@@
 
-# --------
+# ---- RMSE function ----
 rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))
 
-# number of fields
+# ---- number of fields parser ----
 parse_n_fields <- function(folder) {
@@
 
+  # numeric pattern
   num <- str_match(f, "(_|\\b)(\\d+)_?fields?\\b")[,3]
@@
 
-  words <- c("one"=1,"two"=2,"three"=3,"four"=4,"five"=5,
-             "ten"=10,"twenty"=20,"thirty"=30,"forty"=40)
-  word <- str_match(f, "(one|two|three|four|five|ten|twenty|thirty|forty)_?fields?")[,2]
+  # text pattern
+  words <- c("one"=1,"three"=3,"five"=5,"ten"=10,"twenty"=20)
+  word <- str_match(f, "(one|three|five|ten|twenty)_?fields?")[,2]
   if (!is.na(word)) return(words[[word]])
@@
 
-# --- model name mapping ---
+# ---- model name parser ----
 parse_model <- function(folder) {
@@
   case_when(
-    str_detect(f, "^cf_python")     ~ "CF_Python",
     str_detect(f, "^cf")            ~ "CF",
@@
 
-# read RDS, CSV
+# ---- read any file ----
 read_any <- function(f) {
@@
 
+# ---- read eonr file ----
 read_eonr_file <- function(f) {
@@
   if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
-    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF or CF_python csv
+    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))
   } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
@@
 
-# list files in dir 
+# ---- list valid files ----
 files_in_dir <- function(dir_path) {
@@
 
-# extract sim id from filename
+# ---- extract sim id ----
 extract_sim_id <- function(base) {
@@
 
+# ---- build one row ----
 row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
@@
 
-# --- RMSE table ----
+# ---- build RMSE table ----
 build_rmse_table <- function(base_dir = here("results"),
@@
     keep(~ str_detect(basename(.x),
-                      regex("^(cf|cf_python|rann|half_rann|simple_ann|no_rf|halfrrf)",
+                      regex("^(cf|cf_outcome|rann|half_rann|simple_ann|no_rf|halfrrf)",
                             ignore_case = TRUE)))
@@
     n_f    <- parse_n_fields(folder)
+
+    # only keep 1, 3, 5, 10, 20
+    if (!n_f %in% c(1, 3, 5, 10, 20)) return(NULL)
 
@@
 
+rmse_tbl_raw <- readRDS(here("results", "rmse_tbl_raw.rds"))
+
 # -- label mapping for models ---
@@
     "NO_RF"      = "NO_RF",
-    "CF"         = "CF",
-    "CF_Python"  = "CF_Python"
+    "CF"         = "CF"
   )

### 2025-10-28 Writing/manuscript_CF_continuous.qmd
Draft of chapter two

@@
     toc: true
-
+    number-sections: true
+    number-tables: true 
+  html:
+    toc: true
 execute:
@@
 Accuracy is assessed by the RMSE of the estimated economically optimal nitrogen rate (EONR) with respect to the true (simulated) EONR.
-Figures 1–4 summarize errors; Figure 5 panels illustrate representative yield–N response functions that help interpret the error patterns.
 
@@
 
-@fig-rmse-hist shows a clear ordering across methods.
-CF concentrates tightly at low RMSE for all training sizes, indicating both accuracy and stability.
-SO_ANN shifts steadily left as the number of fields increases, with visibly narrower spreads at 5 and 10 fields.
-In contrast, NO_ANN exhibits a heavy right tail with a single field but collapses toward the leading methods at 5–10 fields, showing its sensitivity to sample size.
-SO_RF improves with more fields yet remains less accurate than SO_ANN or CF.
-NO_RF has broad, right-shifted distributions and is a weak baseline.
-DO_ANN performs poorly across all sizes, with distributions centered at very high RMSE and little improvement from additional fields.
+Taken together, @fig-rmse-hist and @fig-avg-performance illustrate clear performance patterns across methods and training sizes.
+CF delivers low and tightly concentrated RMSE distributions even with limited data—consistently maintaining strong accuracy and stability for 1–5 fields (mean RMSE ≈ 14.7, 13.1, and 12.1, respectively), with only modest additional gains at 10 and 20 fields (11.9–11.8).
+SO_ANN initially lags behind CF (18.4 at one field), but its mean RMSE decreases steadily as more fields are added, surpassing CF performance at 10 and 20 fields (11.2 and 8.5, respectively).
+This pattern underscores the neural network’s superior ability to exploit richer spatial and agronomic heterogeneity when sufficient training data are available.
+
+By contrast, NO_ANN exhibits substantial instability in the single-field case, with a heavy right tail and much higher errors, but rapidly converges toward the leading methods by 5–10 fields (mean RMSE improving to around 14–13), highlighting its strong dependence on sample size.
+SO_RF improves modestly with more fields (mid-20s RMSE for 10–20 fields) but remains consistently less accurate than SO_ANN or CF.
+NO_RF maintains broad, right-shifted distributions and higher mean errors throughout (≈ 36–40), confirming its role as a weak baseline.
+Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting over-correction in the orthogonalization process.
 
@@
 ```
-
-Mean RMSE by model and training size (@fig-avg-performance) confirms these patterns.
-CF is consistently low (≈15 across sizes).
-SO_ANN improves monotonically as data increase and attains the best mean at 10 fields (≈11).
-NO_ANN is unusable with a single field (mean ≈82) but becomes competitive by 5–10 fields (≈14 and ≈13).
-SO_RF yields moderate means (≈22 at 10 fields), while NO_RF remains highest among the non-ANN baselines (≈36–40).
-DO_ANN is uniformly poor (≈90–93).
-The heatmap (Fig.
-3) prints the means in each cell and emphasizes the strong gains from orthogonalization (SO_ANN and CF) relative to their non-orthogonal counterparts.
 
@@
 
-Using NO_RF as a baseline, CF reduces mean RMSE by roughly 58–61 % across training sizes (@fig-improvement).
-SO_ANN delivers comparable or larger gains (≈54–69 %), with the largest improvement at 10 fields.
-SO_RF provides modest gains (≈38–42 %).
-NO_ANN is worse than the baseline with 1 field but improves substantially at 5–10 fields (≈63–65 % reduction).
-DO_ANN is consistently inferior to the baseline at all sizes.
+@fig-improvement reports each method’s percentage improvement in mean RMSE relative to the NO_RF baseline.
+CF consistently provides the strongest gains among the robust approaches when data are scarce—reducing error by ≈63% with a single field, and climbing to ≈67–68% once 5–20 fields are available.
+SO_ANN exhibits a similar improvement trend but grows even more dominant as more diverse data accumulate, achieving ≈53% improvement at 1 field, ≈63% by 5 fields, and reaching ≈68% at 10 fields and ≈77% at 20 fields, the highest improvement across all models.
+NO_ANN transitions from a large negative result when trained on just one field (substantially worse than baseline) to strong relative improvements exceeding 65% by 10–20 fields, reinforcing its reliance on adequate sample support.
+SO_RF offers moderate improvement (≈38–43%) that changes little with added fields.
+DO_ANN remains entirely uncompetitive across all sample sizes—showing worsening performance relative to NO_RF by more than 100–150%, indicating severe model misspecification.
+
+Together, these results emphasize that while CF is the most reliable option under limited data availability, SO_ANN ultimately yields the largest relative error reduction when multi-field datasets are accessible, highlighting its superior scalability with richer training information.
 
@@
 ```
+@tbl-best_model_by_field_count highlights the top-performing model for each training size.
+CF is the dominant method when only a few fields are available, delivering the lowest mean RMSE for 1–5 fields.
+As field diversity increases, SO_ANN begins to outperform CF, becoming the preferred approach at 10 fields and showing a substantial further accuracy gain at 20 fields.
+
 ```{r}
@@
   tab_header(
-    title = "Best models by #field",
+    title = "Best models by field count",
     subtitle = "Ordered by lowest mean RMSE"
@@
 
-Predicted yield–N curves (@fig-yield-res-1 through @fig-yield-res-3) for five randomly selected analysis units clarify why methods differ.
-SO_ANN produces agronomically plausible, monotone-concave responses with clear plateaus (ten-field panel), yielding interior optima and low EONR error.
-NO_ANN resembles SO_ANN once trained on 5–10 fields, but with a single field it is noticeably noisier, aligning with its heavy-tailed errors.
-Tree-based models (NO_RF and SO_RF) generate step-like, piecewise-constant responses; the implied EONR can only move at jump points, which coarsens the optimum and limits accuracy.
-DO_ANN often yields nearly linear increasing curves across the tested N range, pushing the optimum to the upper bound and explaining its uniformly high error.
-Although CF is not naturally expressed as per-unit yield curves, its error distribution indicates high stability across all sizes.
-
 ```{r}
@@
 ```
+
+@fig-yield-res-combined displays predicted and true yield response functions for the best-performing models at each training size.
+For the CF, shown for 1, 3, and 5 fields, the estimated curves closely track the overall shape of the true response but exhibit a slight upward bias in predicted yield levels.
+The general curvature, including diminishing returns at higher nitrogen rates, is captured consistently even under data-constrained conditions.
+As more fields are incorporated, the CF predictions become smoother and more aligned with the true function, reflecting improved generalization.
+
+For the SO_ANN, shown for 10 and 20 fields, the estimated response matches the true curve more closely in both slope and curvature, particularly at higher nitrogen levels where CF tended to flatten.
+At 20 fields, SO_ANN produces the most accurate approximation overall, with prediction and truth nearly coincident across the nitrogen range.
+These differences visually reinforce the earlier quantitative results: CF offers strong robustness with sparse multi-field data, while SO_ANN increasingly excels as richer spatial heterogeneity becomes available, enabling superior recovery of the underlying agronomic yield response function.
 
@@
 
+# Discussion
+
+This study evaluates alternative causal-ML strategies for estimating heterogeneous yield response to a continuous nitrogen treatment in an OFPE setting.
+CF consistently attains low and tightly concentrated EONR errors with only 1–5 training fields, indicating strong small-sample stability (see @fig-rmse-hist and @fig-avg-performance).
+SO_ANN improves monotonically with additional fields and ultimately dominates once ≥10 fields are available, achieving the lowest mean RMSE overall (e.g., 11.2 at 10 fields and 8.5 at 20 fields).
+These complementary strengths that CF is preferred under data-constrained conditions, whereas SO_ANN becomes the method of choice as multi-field datasets accumulate (see @tbl-best_model_by_field_count).
+
+Methodologically, the contrast between single-orthogonalization and double-orthogonalization is central.
+For continuous treatments, the double-orthogonal residualization of the treatment does not generally preserve the relevant heterogeneous response structure.
+Our simulations reflect this: DO_ANN performs poorly and does not benefit from larger samples.
+In contrast, single-orthogonalization—residualizing only the outcome to remove baseline variation from covariates $W$—retains the full interaction between $X$ and $N$, allowing the final learner to recover the nonlinearity and heterogeneity of the response function.
+The extended CF strikes a different balance: by expressing $N$ through a spline basis and orthogonalizing the basis components, it restores a linear-in-parameters score that CF can learn reliably.
+
 # Conclusion
 
-Across 500 replications, SO_ANN (with ≥10 fields) and CF deliver the best overall accuracy.
-Orthogonalization is critical: SO_ANN and CF are consistently superior to NO_ANN and NO_RF, especially in small-to-moderate samples.
-NO_ANN becomes competitive only once training data reach at least five fields.
-Tree-based baselines remain substantially less accurate, and the doubly-orthogonal ANN (DO_ANN) is inappropriate for this setting because it fails to recover the saturating response shape.
-These findings are robust across training sizes and are consistent.
+By combining orthogonalization with flexible final-stage learners, we demonstrate how to preserve the strengths of causal inference while accommodating continuous treatments and agronomic heterogeneity.
+
+Across 500 simulated test fields, two approaches consistently dominate: the extended causal forest and the single-orthogonalized neural network.
+CF is highly robust in small-sample environments, achieving strong accuracy even with training data from just one to five fields.
+As the number of training fields increases, SO_ANN increasingly improves as a result of spatial and agronomic diversity, ultimately delivering the best predictive accuracy of all methods once ten or more fields are available.
+In contrast, non-orthogonalized baselines require larger samples to become competitive and the doubly-orthogonalized ANN performs poorly in all conditions due to the incompatibility between double-orthogonalization and continuous treatments.
+Methodologically, our results highlight the importance of aligning the orthogonalization strategy with the structure of the treatment variable in causal ML applications.

### 2025-11-16 Proposal_presentation/Chapters_1&2.qmd
proposal presentation

---
title: "Chapter 1: A new model selection approach based on local economically optimal input rate"
author:
  - name: Taro Mieno$^*$
  - name: Mona Mousavi$^*$
  - name: David Bullock$^\dagger$
institute:
  - $^*$ University of Nebraska Lincoln
  - $^\dagger$ University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
header-includes: 
- \logo{\ifnum\thepage>1\includegraphics[width=1cm]{nebraska-n.jpg}\fi}
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

# Background and Research Objectives

## Background

+ Uniform input management strategies can lead to both economic losses and environmental inefficiencies.

+ Variable-rate input management as a solution (accounts for spatial heterogeneity in crop input demand.

+ On-farm precision experiments (OFPE) as a promising way (GPS technologies and variable-rate applicators).
  + Run on-farm agronomic trials and collect data
  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
  + Identify site-specific optimal nitrogen rates

## Motivation

+ Machine learning approaches being increasingly used for estimating site-specific yield response functions
  + identify the best model among several candidate ML approaches and also tuning hyper parameters via cross-validation as the key part of such process.

+ In all of the previous studies, the best model was selected based on their abilities to predict \textcolor{blue}{yield level} at the observation points (yield-based model selection).

+ However, the ultimate goal is not to predict yield levels, but to estimate yield response (causal marginal impact of nitrogen on crop yield).

## Research Objective

+ Proposes an alternative to conventional yield-based model selection. Specifically, we select models based on their ability to predict local EONR using spatial cross-validation.

+ Evaluate this approach, through Monte Carlo simulations, and compare its performance against the standard yield-based selection method.

# Method

Monte Carlo simulations with 500 iterations.

Steps for a single iteration:

  + Step 1: Generate OFPE experiment data
  + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
  + Step 3: Select the best model among the 5 candidate models based on
    + yield (conventional approach)
    + local-EONR (our approach)
  + Step 4: Evaluate the model selection approaches

The five candidate models:
linear model, linear spatial error (SE) model, random forest (RF), boosted regression forest (BRF), and causal forest (CF).

## Data generation

+ The data generated from a simulated corn production field (about 37.3 hectares) with spatially varying field characteristics.

+ One possible scenario of the spatial variability in the field characteristics is shown.

+ This study assumed the yield data were collected at the subplot level (total 1440 subplots).

## The simulated OFPE field:
```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "95%"
knitr::include_graphics("~/Desktop/g_field_map.png")
```

## True yield responses

The true yield is assumed to follow the quadratic plateau functional form as follows:

$$
y_{j, i}=f_{j, i}(N)= \begin{cases}\alpha_{j, i}+\beta_{j, i} N+\gamma_{j, i} N^2+\varepsilon_{j, i}, & N<\tau_{j, i} \\ \eta_{j,i}+\varepsilon_{j, i}, & N \geq \tau_{j, i}\end{cases}
$$

+ $j \in{1,\ 2,\ 3,...,1440}$: the subplot,
+ $i \in{1,\ ...,6}$: the cell within a subplot,
+ $y_{j,i}$: Corn yield,
+ $N$: Nitrogen rate,
+ $\alpha, \beta, \gamma$: parameters (soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$
+ $\tau_{j, i}$:  the $N$ rate at which the yield reaches a plateau.
+ $\eta_{j,i}$: the yield at $N = \tau_{j, i}$,
+ $\varepsilon_{j, i}$: spatial autocorrelation.

---

The true economically optimal N rate for each cell is then determined by:

$$
N_{j, i}^* \equiv \operatorname{argmax}_N\left[p \cdot f_{j, i}(N)-w \cdot N\right]
$$

+ $p$: the crop price, 
+ $w$ the nitrogen fertilizer price, 
+ $f_{j, i}(N)$: previously defined true yield response function for the specific cell.

---

## On-farm trial (nitrogen rate)

+ six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) followed a 6 × 6 Latin square design.

(Note: 8 blocks, each block including 36 plots ($8*36=288$ plots) to establish the Latin square design).

---

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics("g_N.png")
```

---

## Yield data

Based on the response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and error term $\varepsilon_{j,i}$, the yield $y_{j, i}$ for each cell $(j, i)$ was generated following the true response function.

* Note:
The harvesting equipment was assumed to be able to accurately collect yield data only at the subplot resolution. Therefore, the observed yield data were at the resolution of subplot level.

---

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics("/Users/monamousavi/Dropbox/Causal_climate/g_y.png")
```

## independent variables

+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and economically optimal N rates $N_{j, i}^*$ are not directly observable by producers or researchers.

+ This study creates some covariates (correlated with those true parameters) to serve as the observable field characteristics:

$$X_j = \left(\beta^{1,1}_j, \beta^{1,2}_j, \beta^{2,1}_j, \beta^{2,2}_j, \tau^{1,1}_j, \tau^{1,2}_j, \tau^{2,1}_j, \tau^{2,2}_j, \eta^{1,1}_j, \eta^{1,2}_j, \eta^{2,1}_j, \eta^{2,2}_j \right)$$

$$\beta = \beta^1 + \beta^2, \;\;
\beta^1 = \beta^{1,1} + \beta^{1,2}, \;\;\beta^2 = \beta^{2,1} + \beta^{2,2}$$ 
$$\tau = \tau^1 + \tau^2, \;\;
\tau^1 = \tau^{1,1} + \tau^{1,2}, \;\;\tau^2 = \tau^{2,1} + \tau^{2,2}$$ 
$$
\eta = \min(\eta^1, \eta^2), \;\;
\eta^1 = \min(\eta^{1,1}, \eta^{1,2}), \;\;
\eta^2 = \min(\eta^{2,1}, \eta^{2,2})
$$

# Identifying the true best model among the five candidate models

+ Each of the candidate models are trained using the whole experiment data and then variable rate EONR are estimated for the whole field.

+ The estimated variable rate N recommendations are evaluated against the true variable rate N to evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):

---

The profit loss of model $m$:

$$
\Delta\pi_m = \pi_{true} - \pi_m
$$

Where:

$$
\pi_{true} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(N^*_i, X_i) - P_N \cdot N^*_i\right]
$$

$$
\pi_{m^*} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(\widehat{N}^*_{m}(X_i), X_i) - P_N \cdot \widehat{N}^*_{m}(X_i)\right]
$$

## Model Selection

+ For both approaches (yield prediction accuracy, local EONR prediction accuracy) spatial cross-validation is employed.
  + The entire dataset is randomly split into a train and test dataset in a spatially clustered manner, which is repeated to create multiple folds.

## An example of the train-test split for each of the five folds

```{r}
knitr::include_graphics("g_split_example.png")
```

## Model selection based on yield level prediction

For each fold:

1. train the model using the train data of the fold
2. predict yield for the test data based on the trained model ($\hat{y}_{i,f}$)
3. calculate RMSE of yield prediction as 
$$
RMSE\text{-}Yield_{m,f} = \sqrt{\frac{\sum_{i=1}^{N_t}(\hat{y}_{i,f} - y_{i,f})^2}{N_t}}
$$

## Model selection based on local EONR

+ True EONRs are never observable in the real world

+ Using estimated “local uniform EONR” as the proxy for the true local EONR  to overcome this challenge

+ Steps for each fold and candidate models:
  1. train the GAM model using the **test** data 
  2. estimate uniform EONR for the **test** data ($\hat{N}^*_{f,gam}$)
  3. train the candidate model using the **train** data 
  4. estimate site-specific EONR for the **test** data based on the trained candidate model ($\hat{N}^*_{i, f,m}$). 
  5. Calcurate the average of the site-specific EONR ($\hat{N}^*_{f,m}$)
  6. Calcurate the RMSE of the estimated local EONR:

$$
RMSE\text{-}LE_{m} = \sqrt{\frac{\sum_{f=1}^{F} (\hat{N}^*_{f,m} - \hat{N}^*_{f,gam})^2}{F}}
$$

## Evaluation of model selection performance

+ Selected models (based on both approaches) are evaluated based on the following difference:

$$
\pi_{m^*} - \pi_{m^s}
$$

+ $\pi_{m^*}$: the highest profit that could be achieved if (EONR from) the true best model were used. 
+ $\pi_{m^s}$: the profit achieved by the selected model

# Results and Conclusions

## Simulation Results: Performance of the model selection approaches

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_ranking.png")
```

## Simulation Results: Accuracy of the model selection approaches

```{r}
#| echo: false
#| results: 'asis'
tbl <- readRDS("main_selection_accuracy_table.rds")
invisible(flextable::save_as_image(tbl, path = "table_accuracy.png"))
knitr::include_graphics("table_accuracy.png")

```

## Simulation Results: profit loss associated with the seleted models relative to the actual best models

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_pi_loss.png")
```

## Simulation Results: Accuracy of Local EONR estimation with GAM

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_gam_leonr.png")
```

## Conclusion

This study demonstrates the economic advantage of local EONR method in making more accurate and cost-effective input management decisions compared to the yield-based model selection.

## Chapter two:

### Causal methods for Site-specific yield response function estimation

# Background and Research Objectives

## Background

+ Efficient nitrogen management remains one of the most critical challenges in modern agricultural production.

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather).

+ An increasing amount of data from on-farm field trials has been accumulated over the years.

+ Estimating yield response functions using data pooled from multiple fields can provide significant agronomic and economic insights.

## Motivation

+ When on-farm trial data is pooled, the treatment variable becomes "continuous" rather than discrete.

+ Existing approaches—such as direct applications of random forests or neural networks—often tend to fail to capture heterogeneous treatment effects of a continuous variable.

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.

## Research Objective

+ Find a causal machine learning approach capable of effectively identifying heterogeneous treatment effects tailored for yield response function estimation using data from on-farm trials, where the treatment variable is continuous and randomized.

+ Evaluate the performance of the proposed method through Monte Carlo simulations, comparing it against commonly used existing approaches.

## Method

### Double-ML (binary treatment ($T$) case):

DML works fine in a \textcolor{blue}{binary} treatment ($T$) case (model assumes linearity in $T$):

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

---

$$
\begin{aligned} & Y_{res}=Y-E[Y \mid X, W]=Y-f(X, W) \\ & T_{res}=T-E[T \mid X, W]=T-h(X, W)\end{aligned}
$$

+ A simple regression of $Y_{res}$ on $T_{res}$ can recover $\theta(X)$:

$$
\begin{aligned}
& Y_{res}= \\
& \begin{aligned}
\theta(X) \cdot T + g(X, W) + v 
- & \underbrace{E\{[\theta(x) \cdot T + g(X, W) + v] \mid X, W\}}= \\
& \textcolor{blue}{=} \textcolor{blue}{\theta(X) \cdot E(T \mid X, W) + g(X, W)} \\
& \textcolor{blue}{=}  \textcolor{blue}{\theta(X) \cdot h(X, W) + g(X, W)}
\end{aligned} \\
& \theta(X) \cdot T + g(X, W) + v 
- \theta(X) \cdot h(X, W) - g(X, W)= \\
& \theta(X) \cdot [\underbrace{T - h(X, W)}_{T_{res}}] + v \\
& \Rightarrow Y_{res} = \theta(X) \cdot T_{res} + v
\end{aligned}
$$

---

### Double-ML (continuous treatment ($T$) case):

DML fails in a \textcolor{blue}{continuous} treatment ($T$) case:

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

---

$\begin{aligned}
& Y_{res}= \\
& \theta(X, T)+g(X, W)+v-E\{\theta(X, T)+g(X, W)+v \mid X, W\}= \\
& \theta(X, T)+g(X, W)+v-E(\theta(X, T) \mid X, W)-g(X, W) \\
& \Rightarrow Y_{res}=\theta(X, T)-E(\theta(X, T) \mid X, W)+v
\end{aligned}$

So, orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.

## Our proposed approaches

+ Single-orthogonalization
+ Extended CF

## Single-orthogonalization

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $Y-f(X,W)$ on $T$

This approach works because The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$.

## Extended CF

+ We propose the use of smoothing splines. In spline regression, a single variable can be decomposed into multiple basis functions in a linear-in-parameter manner.

```{=latex}
\begin{align} \label{eq-spline-cf}
y = \sum_{k=1}^K \theta_k(X)\cdot S_k(T)+ g(X, W) + v
\end{align}
```

+ $S_k(T)$: the value of the basis function $S_k(\cdot)$ evaluated at $T$.

## Models

$$
y = f(X, N) + g(X,W) + v
$$

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

## Scenarios

+ Estimation approaches:

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("f_est.png")
```

+ For each of 500 simulation rounds, one field is selected as the test field. The remaining fields are combined to construct the training set in five scenarios: 
  + number of fields from which data is combined: 1, 3, 5, 10 and 20

## Site-specific EONR estimation

For a given price of corn and nitrogen EONR is found by solving the following profit maximization problem:

$$
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
$$

+ $\widehat{f}$: yield response curve estimated by the model under evaluation

## Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

$$
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
$$

# Results and Conclusions

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("fig_rmse_hist.png")
```

## Simulation (preliminary) results

```{r}
#| fig-align: "center"
#| out-width: "45%"
knitr::include_graphics("tbl_best_model_by_field_count.png")
```

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("model_improvement.png")
```

## Conclusion

By combining orthogonalization with flexible final-stage learners, we demonstrate how to preserve the strengths of causal inference while accommodating continuous treatments and agronomic heterogeneity.

### 2025-11-17 Writing/manuscript_CF_continuous.qmd
comments on the manuscript added

@@
 ---
-title: "Estimation of the effect of a continuous treatment variable"
+title: "Estimation of the Effect of a Continuous Treatment Variable"
+author:
+  - name: Mona Mousavi
+    orcid: 
+    email: mmousavi2@huskers.unl.edu
+    affiliations:
+      - name: University of Nebraska Lincoln
+  - name: Taro Mieno
+    orcid: 0000-0002-0614-0771 
+    email: tmieno2@unl.edu
+    affiliations:
+      - name: University of Nebraska Lincoln
+  - name: Takashi Tanaka
+    orcid: 0000-0001-7116-6962
+    email: takashi@agro.au.dk
+    affiliations:
+      - name: Aarhus University
+  - name: David Bullock
+    orcid: 
+    email: dsbulloc@illinois.edu
+    affiliations:
+      - name: University of Illinois at Urbana Champaign
 format:
-  docx:
+  html:
     toc: true
     number-sections: true
-    number-tables: true 
-  html:
+    number-depth: 3
+  pdf:
+    toc: true
+    number-sections: true
+    pdf-engine: xelatex
+  docx:
     toc: true
@@
 Efficient nitrogen (N) management remains one of the most critical challenges in modern agricultural production.
-Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm.
+Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm (<span style = "color: blue;"> add citations here </span>).
 The principle of applying the right amount of nitrogen at the right place and time requires understanding the spatial and temporal variability in crop nitrogen needs.
@@
 
-On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management.
+On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management (<span style = "color: blue;"> add citations here.
+e.g., Dave's paper in agronomy and Lacoste's nature paper </span>).
 Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions.
@@
 
+::: {.column-margin}
+Right now, it is written as if prediction-oriented and causal ML have been used in parallel.
+But, @kakimoto2022causal is in response to the danger of applying simply applying prediction-oriented ML approaches.
+Modify the text to reflect that here.
+Cite [this](https://link.springer.com/article/10.1007/s11119-024-10136-x) to motivate the use of causal ML.
+:::
+<!--end of column-margin-->
+
 In the OFPE setting, causal machine learning methods have shown to be promising to recover the full yield response curve.
 @kakimoto2022causal showed that causal forest can estimate heterogeneous treatment effects of an input on yield more accurately than prediction-oriented machined approaches like random forest, boosted forest, and convolutional neural network.
-However, their application of CF is only limited to a single-year OFPE experiment from a single field.
-Causal forest was developped primarily to estimate
-the heterogeneous treatment effects of a dicrete treatment. In OFPE, there are typically five
-or six distinctive target experimental rates (see the first panel of @fig-distribution-N).
-In @kakimoto2022causal, CF estimates the effect of multiple treatments defined as the increase in nitrogen rate from the lowest rate to the other rates (so n − 1 tretments in n-rate trials).
-However,
-once data from multiple fields with different target rates are combined, applied nitrogen rates
-become rather continuous as seen in the second panel of @fig-distribution-N.
-Existing approaches—such as direct applications of random forests or neural networks and CF—often tend to fail to capture heterogeneous treatment effects of a continuous variable (@kakimoto2022causal).
+However, their application of traditional CF is only limited to a single-year OFPE experiment from a single field because it was developed primarily to estimate
+the heterogeneous treatment effects of a discrete treatment.
+In OFPE, there are typically five or six distinctive target experimental rates (see the first panel of @fig-distribution-N).
+In @kakimoto2022causal, CF estimates the effect of multiple treatments defined as the increase in nitrogen rate from the lowest rate to the other rates (so n − 1 treatments in n-rate trials).
+However, once data from multiple fields with different target rates are combined, applied nitrogen rates become rather continuous as seen in the second panel of @fig-distribution-N.
+While existing prediction-oriented ML approaches—such as direct applications of random forests or neural networks — can naturally handle continuous treatment variable, they tend to fail to capture heterogeneous treatment effects of a treat variable (@kakimoto2022causal).
+As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights.
+It creates a need for methods that can estimate heterogeneous treatment effects of a continuous treatment variable.
 
@@
 ```
-As multi-year, multi-field OFPE datasets accumulate, pooling data across fields has become both feasible and desirable—it can improve estimation of yield response functions and deliver stronger agronomic and economic insights.
-It creates a need for methods that handle continuous treatments and it becomes especially important to adapt causal methods to continuous treatments without losing their statistical strengths over other machine-learning approaches.
 
@@
 
+::: {.column-margin}
+<span style = "color: blue;"> We need to emphasize that we are proposing approaches that have not been tried anywhere: SO-NN and semi-parametric-CF.
+They are the main approaches.
+We then have other contrasting approaches to compare against.
+</span>
+:::
+<!--end of column-margin-->
+
 We find that the SO-ANN approach outperforms the non-orthogonalized model, which represents the common practice in yield response function estimation.
@@
 
+::: {.column-margin}
+<span style = "color: blue;"> SO-ANN not introduced yet. You need to talk more about what you do in the above paragraph (mostly SO-NN and semi-parametric CF). </span>
+:::
+<!--end of column-margin-->
+
 # Method
@@
 
-```{=tex}
-\begin{align*}
+$$
 Y=\theta(X) T+g(X, W)+v
-\end{align*}
-```
+$$
 
@@
 Orthogonalization estimates $f(X, W)=\mathbb{E}[Y \mid X, W]$ and $h(X, W)=\mathbb{E}[T \mid X, W]$, forms residuals $Y_{\mathrm{res}}=Y-f(X,W)$ and $T_{\mathrm{res}}=T-h(X, W)$, and then relates $Y_{\mathrm{res}}$ to $T_{\mathrm{res}}$.
-Because this score is Neyman-orthogonal and is typically implemented with cross-fitting, small errors in $f$ and $h$ do not induce first-order bias in $\theta(X)$.
+Because this score is Neyman-orthogonal and is typically implemented with cross-fitting, small errors in $f$ and $h$ do not induce first-order bias in $\theta(X)$ [(@chernozhukov2018double].
 
@@
 
-```{=tex}
+$$
 \begin{align*}
@@
 \end{align*}
-```
+$$
 
@@
 
-@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield and treated heterogeneity in a discrete way.
+@kakimoto2022causal applied this framework to estimate heterogeneous treatment effects of nitrogen on yield.
 However, when data are pooled across fields and applied nitrogen rates become continuous, the standard residualization of the treatment in DML does not extend cleanly to that setting.
@@
 \end{align*}
-```
+$$
 
 and the clean decomposition that underlies double-orthogonalization in the linear-in-$T$ case breaks down.
+
 Instead, we apply single-orthogonalization (SO).
@@
 
+::: {.column-margin}
+<span style = "color: blue;"> We will add justifications for this action later. </span>
+:::
+<!--end of column-margin-->
+
 To study and compare these ideas systematically, we adopt a single underlying structural model and then implement alternative orthogonalization schemes.
@@
 \end{align*}
-```
+$$
 where $P_{\text {Corn }}$ is the corn price, $P_{\text {N }}$ is the nitrogen price, and $\hat{f}$ is the estimated yield response.
@@
 
+::: {.column-margin}
+<span style = "color: blue;"> Explain in  more detail about why it this is innovative. Explain the concept of linearization with splines using the figure I generated in my old manuscript. </span>
+:::
+<!--end of column-margin-->
+
 To adapt double-orthogonalization to the model of interest (Model 3), we estimate the heterogeneous response $f(X,N)$ by expanding the continuous treatment $N$ in a set of spline basis functions and allowing their coefficients to vary with the heterogeneity-inducing covariates $X$:
 
-```{=tex}
+$$
 \begin{align*}
@@
 
-```{=tex}
+$$
+\begin{align*}
 \tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
 \end{align*}
-```
+$$
 
@@
 
-```{=tex}
+$$
+\begin{align*}
 \hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
 \end{align*}
-```
+$$
 
@@
 These simulations serve as a controlled environment for evaluating model performance in estimating yield response functions and recommending optimal nitrogen rates.
+
+<span style = "color: red;"> Add profit differential as the measure. Discuss them in the results as well. </span>
 
@@
 
-```{=latex}
+$$
 \begin{align} 
@@
 \end{align}
-```
+$$
 
@@
 The erorr term ($\varepsilon_{i,t}$) is generated using the same spatial process.
-This is repeated independenty 500 times to create 500 fields of the same size but different spatial distribution and values of the three parameters.
+This is repeated independently 500 times to create 500 fields of the same size but different spatial distribution and values of the three parameters.
 
@@
 ```{r fig-trial-design, fig.cap = "Example trial design"}
-readRDS(here::here("writing/Figures/g_td.rds"))
+g_td <- readRDS(here::here("writing/Figures/g_td.rds"))
 ```
@@
 NO_RF maintains broad, right-shifted distributions and higher mean errors throughout (≈ 36–40), confirming its role as a weak baseline.
-Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting over-correction in the orthogonalization process.
+Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting <span style = "color: red;"> over-correction </span>   in the orthogonalization process.
+
+::: {.column-margin}
+<span style = "color: blue;"> It is not over-correction. It is fundamentally messed up as the original functions are no longer retained. </span>
+:::
+<!--end of column-margin-->
 
@@
 
+# ggsave(
+#   filename = here::here("Proposal_presentation/fig_rmse_hist.png"),
+#   plot = p,
+#   width = 12,
+#   height = 7,
+#   bg = "white"
+# )
+
 ```
@@
 # ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)
+
+# ggsave(here::here("Proposal_presentation","model_improvement.png"), p_mean, width = 9, height = 5, dpi = 300)
 ```
@@
 
+::: {.column-margin}
+<span style = "color: blue;"> All this is probably redundant. What are you adding beyond the previous paragraphs? </span>
+:::
+<!--end of column-margin-->
+
 Together, these results emphasize that while CF is the most reliable option under limited data availability, SO_ANN ultimately yields the largest relative error reduction when multi-field datasets are accessible, highlighting its superior scalability with richer training information.
@@
 
-best_models %>%
+best_table <- best_models %>%
   gt() %>%
@@
 
+#gtsave(best_table, here::here("Proposal_presentation/tbl_best_model_by_field_count.png"))
+
 ```
@@
 ```
+
+<span style = "color: blue;"> I do not see the purpose of this much. Contrast CF, SO-NN, and the worst (all in the same panel) across the number of fields. </span>
 

### 2025-11-17 Writing/manuscript_CF_continuous.tex
comments on the manuscript added

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{anyfontsize}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimation of the Effect of a Continuous Treatment Variable},
  pdfauthor={Mona Mousavi},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Estimation of the Effect of a Continuous Treatment Variable}
\author{Mona Mousavi}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Efficient nitrogen (N) management remains one of the most critical
challenges in modern agricultural production. Nitrogen is essential for
achieving high crop yields, yet excessive or poorly timed applications
can lead to economic losses and environmental harm. The principle of
applying the right amount of nitrogen at the right place and time
requires understanding the spatial and temporal variability in crop
nitrogen needs. This variability arises from complex interactions among
soil nitrogen availability, soil and field characteristics, and weather
conditions. Accurately identifying the optimal nitrogen rate for each
site and season is therefore fundamental to improving both profitability
and sustainability in crop production systems.

On-farm field experiments have emerged as a powerful approach to
generate the data necessary for site-specific nitrogen management.
Through on-farm agronomic trials, researchers can directly observe yield
responses to varying nitrogen rates under real-world field conditions.
Analyzing these data enables estimation of yield response functions and
identification of site-specific economically optimal nitrogen rates.
Recent advances in data analytics---particularly the widespread adoption
of machine learning (ML) methods---have expanded the potential of
on-farm experimentation. Two major methodological directions have
emerged: (1) yield prediction--based methods, which recommend
site-specific input rates by predicting yield as a function of nitrogen
and environmental covariates (Krause et al. (2020); Barbosa et al.
(2020); Barbosa, Hovakimyan, and Martin (2020); Gardner, Mieno, and
Bullock (2021)), and (2) causal machine learning (causal ML) methods
(Kakimoto et al. (2022)), which estimate the underlying causal
relationship between nitrogen application and yield response.

In the OFPE setting, causal machine learning methods have shown to be
promising to recover the full yield response curve. Kakimoto et al.
(2022) showed that causal forest can estimate heterogeneous treatment
effects of an input on yield more accurately than prediction-oriented
machined approaches like random forest, boosted forest, and
convolutional neural network. However, their application of CF is only
limited to a single-year OFPE experiment from a single field. Causal
forest was developped primarily to estimate the heterogeneous treatment
effects of a dicrete treatment. In OFPE, there are typically five or six
distinctive target experimental rates (see the first panel of
Figure~\ref{fig-distribution-N}). In Kakimoto et al. (2022), CF
estimates the effect of multiple treatments defined as the increase in
nitrogen rate from the lowest rate to the other rates (so n − 1
tretments in n-rate trials). However, once data from multiple fields
with different target rates are combined, applied nitrogen rates become
rather continuous as seen in the second panel of
Figure~\ref{fig-distribution-N}. Existing approaches---such as direct
applications of random forests or neural networks and CF---often tend to
fail to capture heterogeneous treatment effects of a continuous variable
(Kakimoto et al. (2022)).

\begin{figure}

{\centering \includegraphics{Figures/g_N_hist_single_multiple.pdf}

}

\caption{\label{fig-distribution-N}Distribution of nitrogen rates from a
single OFPE and multiple OFPEs}

\end{figure}

As multi-year, multi-field OFPE datasets accumulate, pooling data across
fields has become both feasible and desirable---it can improve
estimation of yield response functions and deliver stronger agronomic
and economic insights. It creates a need for methods that handle
continuous treatments and it becomes especially important to adapt
causal methods to continuous treatments without losing their statistical
strengths over other machine-learning approaches.

The key objective of this article is to identify a causal machine
learning approach capable of estimating heterogeneous treatment effects
for yield response function estimation using data from on-farm trials,
where the treatment variable---nitrogen rate---is continuous and
randomized. We investigate two complementary approaches: a
Single-Orthogonalized Neural Network (SO-ANN), designed to flexibly
capture nonlinear treatment--response relationships, and an extended
Causal Forest (CF) that incorporates smoothing splines to model
continuous treatments. Both methods aim to preserve the strengths of
causal ML---nonparametric modeling of complex interactions without
restrictive assumptions. Their performance is evaluated through Monte
Carlo simulations that mimic typical OFPE conditions and crop yield
responses to nitrogen.

We find that the SO-ANN approach outperforms the non-orthogonalized
model, which represents the common practice in yield response function
estimation. In contrast, the Double-Orthogonalized approach performs
poorly, and its accuracy does not improve even with larger sample sizes,
consistent with theoretical expectations. Our extended CF method
successfully captures heterogeneous and nonlinear effects of a
continuous treatment and accurately estimates site-specific
profit-maximizing nitrogen rates with minimal bias underscoring the
promise of this approach. Although the methodological extension proposed
here is primarily motivated by the needs of OFPE, it holds potential for
a wide range of applications in applied economics where continuous
treatments and heterogeneous effects are of interest.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{models-and-estimation-procedures}{%
\subsection{Models and estimation
procedures}\label{models-and-estimation-procedures}}

The standard approach in the precision-agriculture literature for
estimating site-specific input response is to model the outcome level
directly as a function of the treatment/input and covariates---e.g.,
estimate \(Y=f(I,C)\) where \(I\) is the input (like nitrogen rate) and
\(C\) are field characteristics---and then derive optimal input
decisions from that estimated function. This is what is typically done
in the recent on-farm precision experimentation literature (e.g.,
Barbosa, Hovakimyan, and Martin (2020); Barbosa et al. (2020); Krause et
al. (2020); Gardner, Mieno, and Bullock (2021)). The problem is that
these methods optimize prediction of yield levels---not the causal
effect of changing the input. Good level prediction at the observed
input does not guarantee accurate recovery of the response function
needed for optimal input choice, because the fitted \(f(I,C)\) can
conflate input variation with confounding and misattribute the drivers
of yield variation.

To identify causal effects rather than just predict levels in the single
binary treatment case, researchers use double/debiased machine learning
(DML), which introduces orthogonalization so the treatment-effect
estimator is insensitive to small errors in the auxiliary fits
(Chernozhukov et al. (2018); Mackey, Syrgkanis, and Zadik (2018)).
Consider the structural model (Model 1):

\[
Y=\theta(X) T+g(X, W)+v
\]

where \(Y\) is the outcome, \(T \in\{0,1\}\) is the treatment, \(X\)
covariates that also modify the treatment effect \(\theta(X)\), \(W\)
additional predictors of \(Y\), and \(v\) is noise. Orthogonalization
estimates \(f(X, W)=\mathbb{E}[Y \mid X, W]\) and
\(h(X, W)=\mathbb{E}[T \mid X, W]\), forms residuals
\(Y_{\mathrm{res}}=Y-f(X,W)\) and \(T_{\mathrm{res}}=T-h(X, W)\), and
then relates \(Y_{\mathrm{res}}\) to \(T_{\mathrm{res}}\). Because this
score is Neyman-orthogonal and is typically implemented with
cross-fitting, small errors in \(f\) and \(h\) do not induce first-order
bias in \(\theta(X)\).

The identifying algebra in Model 1 follows from linearity in \(T\):

\[
\begin{align*}
\mathbb{E}[Y \mid X, W]=\theta(X) \mathbb{E}[T \mid X, W]+g(X, W)=\theta(X) h(X, W)+g(X, W)
\end{align*}
\]

so

\[
\begin{align*}
Y_{\mathrm{res}}=Y-\mathbb{E}[Y \mid X, W]=\theta(X)(T-h(X, W))+v=\theta(X) T_{\mathrm{res}}+v
\end{align*}
\]

Thus regressing \(Y_{\mathrm{res}}\) on \(T_{\mathrm{res}}\) recovers
\(\theta(X)\) up to noise.

Kakimoto et al. (2022) applied this framework to estimate heterogeneous
treatment effects of nitrogen on yield and treated heterogeneity in a
discrete way. However, when data are pooled across fields and applied
nitrogen rates become continuous, the standard residualization of the
treatment in DML does not extend cleanly to that setting.

Consider the continuous-treatment setting, where the structural model is

\[
\begin{align*}
Y=\theta(X, T)+g(X, W)+v
\end{align*}
\]

with \(T\) (e.g., nitrogen rate) continuous, and \(X\)
heterogeneity-inducing covariates that interact with \(T\) through the
flexible response function \(\theta(X, T)\). In this case
double-orthogonalization fails because \(\theta\) is a general
(nonlinear and interaction-rich) function of \(T\). Residualizing the
treatment by replacing \(T\) with \(T-\mathbb{E}[T \mid X, W]\) does not
produce a valid adjustment for the heterogeneous effect. Formally,

\[
\begin{align*}
\theta(X, T)-\mathbb{E}[\theta(X, T) \mid X, W] \neq \theta(X, T-\mathbb{E}[T \mid X, W])
\end{align*}
\]

and the clean decomposition that underlies double-orthogonalization in
the linear-in-\(T\) case breaks down. Instead, we apply
single-orthogonalization (SO). We first estimate
\(f(X, W)=\mathbb{E}[Y \mid X, W]\) and form the residual
\(Y_{\mathrm{res}}=Y-f(X, W)\), and then recover \(\theta(X, T)\) by
flexibly modeling \(Y_{\mathrm{res}}\) as a function of \((X,T)\). This
approach removes baseline variation due to \(g(X,W)\) while preserving
the full interaction structure between \(X\) and the continuous
treatment.

To study and compare these ideas systematically, we adopt a single
underlying structural model and then implement alternative
orthogonalization schemes.

The model of interest throughout is (Model 3)

\[
\begin{align*}
y = f(X, N) + g(W) + v
\end{align*}
\]

where \(y\) is crop yield, \(N\) is the continuous nitrogen rate, \(X\)
is a set of covariates that induce heterogeneity in the impact of \(N\)
on \(y\), and \(W\) is a set of covariates that affect \(y\) but do not
interact with \(N\). We implement three estimation
strategies---double-orthogonalized, single-orthogonalized, and
non-orthogonalized---that differ only in how (and whether) the outcome
and treatment are residualized before the final fit. Orthogonalization
means fitting a variable on its conditioning covariates and using the
residual (observed minus predicted) so that the subsequent learner
focuses on variation not explained by those covariates; all
residualizations are performed with random forests. In the
double-orthogonalized approach, both \(y\) and \(N\) are residualized
with respect to \((X,W)\) and the final learner is trained on those
residuals. In the single-orthogonalized approach, only \(y\) is
residualized and the raw \(N\) is retained. The non-orthogonalized
specification skips residualization entirely and fits the learner
directly to \((y,N,X,W)\). For each strategy we consider three
final-stage learners---neural networks (NN), random forests (RF), and
causal forests (CF)---to evaluate how the orthogonalization scheme and
learning algorithm interact in recovering heterogeneous nitrogen
response.

The table below shows all the cases we consider.

\includegraphics[width=4in,height=\textheight]{f_est.png}

\hypertarget{single-orthogonalized-nn-so-nn-and-so-rf}{%
\subsubsection{Single-orthogonalized NN (SO-NN) and
(SO-RF)}\label{single-orthogonalized-nn-so-nn-and-so-rf}}

In the single-orthogonalized framework only the baseline component
\(g(W)\) is partialed out before the final-stage estimation; the
continuous treatment \(N\) and heterogeneity-inducing covariates \(X\)
remain in their original form. Formally, starting from Model 3, we first
estimate \(g(W)=\mathbb{E}[y \mid W]\) using a cross-fitted random
forest and form the residualized outcome

\[
\begin{align*}
\tilde{y}=y-\widehat{g}(W)
\end{align*}
\]

The second stage fits a flexible model of \(\tilde{y}\) on \((X,N)\) to
recover the heterogeneous response function \(f(X,N)\). The distinction
between SO-NN and SO-RF is only in the choice of the final learner:
SO-NN uses a neural network, while SO-RF uses a random forest. In both
cases the learner estimates \(\hat{f}(X, N)\), which is then used to
predict yield under candidate nitrogen rates. Marginal profit is
computed over a grid of \(N\) values and the economically optimal
nitrogen rate is selected as the one that maximizes

\[
\begin{align*}
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
\end{align*}
\] where \(P_{\text {Corn }}\) is the corn price, \(P_{\text {N }}\) is
the nitrogen price, and \(\hat{f}\) is the estimated yield response.

\hypertarget{double-orthogonalized-cf-do-cf-and-ann-do-ann}{%
\subsubsection{Double-orthogonalized CF (DO-CF) and ANN
(DO-ANN)}\label{double-orthogonalized-cf-do-cf-and-ann-do-ann}}

\hypertarget{docf-continuous}{%
\paragraph{DO‑CF (continuous)}\label{docf-continuous}}

To adapt double-orthogonalization to the model of interest (Model 3), we
estimate the heterogeneous response \(f(X,N)\) by expanding the
continuous treatment \(N\) in a set of spline basis functions and
allowing their coefficients to vary with the heterogeneity-inducing
covariates \(X\):

\[
\begin{align*}
y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
\end{align*}
\]

Although double-orthogonalization generally fails when \(f(X,N)\) is an
arbitrary nonlinear function of \(N\), this spline-based formulation
restores the key linear-in-parameter structure that makes DO valid,
because the model is linear in the transformed treatment basis
\(\left\{S_k(N)\right\}\). Residualizing each basis component and the
outcome with respect to \((X,W)\) then yields a clean decomposition
analogous to the linear-in-\(T\) case. Let
\(\widetilde{S}_k(N)=S_k(N)-\mathbb{E}\left[S_k(N) \mid X, W\right]\).
Then

\[
\begin{align*}
\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
\end{align*}
\]

so causal forest can nonparametrically estimate each heterogeneous
weight \(\theta_k(X)\) by relating the residualized outcome to the
residualized spline bases. This Neyman-orthogonal (R-learner) style
decomposition preserves the ability to capture complex, heterogeneous,
and nonlinear treatment effects while inheriting the robustness of
double-orthogonalization: small errors in the nuisance fits for
\(g(X,W)\) or the conditional means of the bases do not induce
first-order bias in \(\theta_k(X)\). The estimated \(\theta_k(X)\) then
combine to recover the full heterogeneous response function and, by
differencing across candidate \(N\) values, site-specific treatment
effects and economically optimal nitrogen rates.

\hypertarget{doann}{%
\paragraph{DO‑ANN}\label{doann}}

In the DO-ANN approach, we apply double-orthogonalization to Model 3 by
residualizing both the outcome \(y\) and the continuous treatment \(N\)
with respect to \((X, W)\) using cross-fitted random forests:

\[
\begin{align*}
\tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
\end{align*}
\]

The final-stage learner is a neural network that models \(\tilde{y}\) as
a linear function of \(\tilde{N}\), with heterogeneity in the slope
captured by covariates \(X\):

\hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
\end{align*}

This structure preserves the linear-in-treatment-residual form required
for valid double-orthogonalization while allowing \(\hat{\theta}(X)\) to
vary flexibly across sites. The neural network receives \(X\) as input
and outputs the site-specific treatment effect.

After training, we reconstruct \(\hat{f}(X, N)\) over a grid of
candidate nitrogen rates by evaluating the model across values of \(N\),
and select the economically optimal nitrogen rate.

\hypertarget{non-orthogonalized-ann-no-nn-and-rf-no-rf}{%
\subsubsection{Non-orthogonalized ANN (NO-NN) and RF
(NO-RF)}\label{non-orthogonalized-ann-no-nn-and-rf-no-rf}}

In the non‑orthogonalized setting, neither the outcome variable nor the
treatment is residualized prior to estimation. Instead, the model
directly fits the observed yield on both the nitrogen rate and
covariates. We estimate the response function \(\hat{Y}=\hat{f}(X, N)\)
using the original values of \(Y\), \(N\) and \(X\), without any
orthogonalization. As in the single‑orthogonalized approaches, the
function \(\hat{f}(X, N)\) is estimated non‑parametrically. The
difference between NO‑NN and NO‑RF lies in the choice of final-stage
learner: NO‑NN uses a neural network, while NO‑RF relies on a random
forest. Once the yield response function is estimated, the EONR is
determined using the same profit-maximization procedure described in
earlier models.

\hypertarget{performance-measurement}{%
\subsection{Performance Measurement}\label{performance-measurement}}

For each simulation round, we evaluate model performance by calculating
the root mean squared error (RMSE) of EONR predictions and the
associated profit loss for each method. The RMSE of EONR predictions in
a given simulation round is computed as follows:

\begin{align*}
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
\end{align*}

To assess the predictive accuracy and economic relevance of these
modeling approaches, we design a set of simulation experiments that
reflect realistic agronomic conditions. These simulations serve as a
controlled environment for evaluating model performance in estimating
yield response functions and recommending optimal nitrogen rates.

\hypertarget{simulation-scenarios}{%
\subsection{Simulation Scenarios}\label{simulation-scenarios}}

To evaluate the performance of the proposed models we design Monte Carlo
simulations that mimic a typical on-farm precision experimentation
(OFPE) setting. These simulations examine the ability of each method to
accurately estimate the site-specific yield response to nitrogen
application and the EONR under varying data availability conditions.

Our simulation consists of 500 independently generated fields, each
representing a distinct agronomic environment with field-specific
parameters that shape the crop's response to nitrogen application. These
parameters vary systematically across fields to introduce heterogeneity
in yield responses. The true yield response in each field follows a
quadratic-plateau functional form, a widely used structure for modeling
nitrogen response curves in agronomic studies.

For each simulation round, one field is selected as the test field. The
remaining fields are used to construct the training set in three
scenarios: (1) one-field training, (2) three-field training, (3)
five-field training, (4) ten-field training and (5) twenty-field
training. These scenarios allow us to assess how model performance
changes with the size and diversity of training data. Within each round,
the model is trained on the designated number of training fields and
then used to estimate the yield response function and the site-specific
EONR for the test field.

The experimental design within each field consists of randomized
nitrogen treatment rates applied across plots. The nitrogen rates vary
between fields and are drawn to reflect realistic trial setups that span
a range of agronomic optima. For evaluation purposes, we use a grid of
candidate nitrogen rates to compute predicted yield and associated
marginal profit for each observation. The predicted EONR is the nitrogen
rate that maximizes predicted profit.

\hypertarget{data-generation}{%
\subsection{Data generation}\label{data-generation}}

First, a 92.17-acre rectangular field is created, which consists of
\(384\) of 60 feet \(\times\) 240 feet plots. each of which will be
assigned an N fertilizer application rate. Each plot is made up of four
(4-rows \(\times\) 1-column) 60 feet \(\times\) 60 feet ``subplots,''
which are the unit of analysis used in subsequent statistical analysis.
Each subplot consists of a 6-rows \(\times\) 6-columns grid of
thirty-six 10 feet \(\times\) 10 feet ``cells.'' Visual representation
of the field layout is presented in Figure~\ref{fig-field-layout}. The
data are generated at the cell-level first and then aggregated up to
analysis-unit level.

\begin{figure}

{\centering \includegraphics{Figures/g_layout.pdf}

}

\caption{\label{fig-field-layout}Visualization of the field layout and
components}

\end{figure}

Cell-level yields are generated following the quadratic-plateau
functional form as follows:

\begin{align} 
y_{i,j} = 
  \begin{cases}
  \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
  \end{cases}
\label{eq:yield-gen-sim}
\end{align}

In this formulation, yield increases as \(N_{i,j}\) increase until
\(N_{i,j}\) reaches \(\tau_{i,j}\), at which yield hits the plateau,
\(\eta_{i,j}\). The yield level at \(N_{i,j}= 0\) is \(\alpha_{i,j}\).
The rate at which yield increases with respect to \(N_{i,j}\) is
governed by \(\beta_{i,j}\) and \(\gamma_{i,j}\). For a given cell, the
quadratic-plateu function has five parameters, but the degree of freedom
is only three. That is, once any three of the five parameters are
determined, then the rest of the parameters are automatically
determined. Specifically, we first generated \(\alpha_{i,j}\),
\(\eta_{i,j}\), and \(\tau_{i,j}\) in a spatially correlated manner
following the Gaussian bluh bluh process (cite), and then
\(\beta_{i,j}\) and \(\gamma_{i,j}\) were derived. The erorr term
(\(\varepsilon_{i,t}\)) is generated using the same spatial process.
This is repeated independenty 500 times to create 500 fields of the same
size but different spatial distribution and values of the three
parameters.

Trial designs vary field by field based on the value of \(\tau_{i,j}\).
Specifically, the lowest nitrogen rate is set at the minimum of
\(\tau_{i,j}\) of the field minus 50, and the highest rate is set at the
maximum of \(\tau_{i,j}\) of the field plus 30. We then create a
sequence of equi-distance 5-level nitrogen rates. Once the unique
nitrogen levels are determined, a Latin square design is used as a trial
design (see Figure~\ref{fig-trial-design}) for an example trial design).
This ensures that nitrogen rate is orthogonal to any variable including
the error term, avoiding endogeneity problem as done in OFPE in real
world. Now, yield values at the cell level can be calculated by plugging
in the value of \(N_{i,j}\), \(\alpha_{i,j}\), \(\eta_{i,j}\),
\(\tau_{i,j}\), and \(\varepsilon_{i,j}\) into
Equation\textasciitilde{}\ref{eq:yield-gen-sim}.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-trial-design-1.pdf}

}

\caption{\label{fig-trial-design}Example trial design}

\end{figure}

\hypertarget{simulation-results}{%
\section{Simulation Results}\label{simulation-results}}

We compare six estimators---CF, SO\_ANN, DO\_ANN, NO\_ANN, SO\_RF, and
NO\_RF---over 500 Monte-Carlo replications and five training sizes (1,
3, 5, 10 and 20 fields). Each replication contains 1,440 evaluation
units. Accuracy is assessed by the RMSE of the estimated economically
optimal nitrogen rate (EONR) with respect to the true (simulated) EONR.

Taken together, Figure~\ref{fig-rmse-hist} and
Figure~\ref{fig-avg-performance} illustrate clear performance patterns
across methods and training sizes. CF delivers low and tightly
concentrated RMSE distributions even with limited data---consistently
maintaining strong accuracy and stability for 1--5 fields (mean RMSE ≈
14.7, 13.1, and 12.1, respectively), with only modest additional gains
at 10 and 20 fields (11.9--11.8). SO\_ANN initially lags behind CF (18.4
at one field), but its mean RMSE decreases steadily as more fields are
added, surpassing CF performance at 10 and 20 fields (11.2 and 8.5,
respectively). This pattern underscores the neural network's superior
ability to exploit richer spatial and agronomic heterogeneity when
sufficient training data are available.

By contrast, NO\_ANN exhibits substantial instability in the
single-field case, with a heavy right tail and much higher errors, but
rapidly converges toward the leading methods by 5--10 fields (mean RMSE
improving to around 14--13), highlighting its strong dependence on
sample size. SO\_RF improves modestly with more fields (mid-20s RMSE for
10--20 fields) but remains consistently less accurate than SO\_ANN or
CF. NO\_RF maintains broad, right-shifted distributions and higher mean
errors throughout (≈ 36--40), confirming its role as a weak baseline.
Finally, DO\_ANN performs poorly throughout, with both histograms and
means showing persistently high errors (≈ 90+) and negligible benefit
from additional data, reflecting over-correction in the
orthogonalization process.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-rmse-hist-1.pdf}

}

\caption{\label{fig-rmse-hist}RMSE histograms by model and number of
fields.}

\end{figure}

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-avg-performance-1.pdf}

}

\caption{\label{fig-avg-performance}Average performance by model and
number of fields.}

\end{figure}

Figure~\ref{fig-improvement} reports each method's percentage
improvement in mean RMSE relative to the NO\_RF baseline. CF
consistently provides the strongest gains among the robust approaches
when data are scarce---reducing error by ≈63\% with a single field, and
climbing to ≈67--68\% once 5--20 fields are available. SO\_ANN exhibits
a similar improvement trend but grows even more dominant as more diverse
data accumulate, achieving ≈53\% improvement at 1 field, ≈63\% by 5
fields, and reaching ≈68\% at 10 fields and ≈77\% at 20 fields, the
highest improvement across all models. NO\_ANN transitions from a large
negative result when trained on just one field (substantially worse than
baseline) to strong relative improvements exceeding 65\% by 10--20
fields, reinforcing its reliance on adequate sample support. SO\_RF
offers moderate improvement (≈38--43\%) that changes little with added
fields. DO\_ANN remains entirely uncompetitive across all sample
sizes---showing worsening performance relative to NO\_RF by more than
100--150\%, indicating severe model misspecification.

Together, these results emphasize that while CF is the most reliable
option under limited data availability, SO\_ANN ultimately yields the
largest relative error reduction when multi-field datasets are
accessible, highlighting its superior scalability with richer training
information.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-improvement-1.pdf}

}

\caption{\label{fig-improvement}Improvement relative to NO\_RF.}

\end{figure}

\textbf{?@tbl-best\_model\_by\_field\_count} highlights the
top-performing model for each training size. CF is the dominant method
when only a few fields are available, delivering the lowest mean RMSE
for 1--5 fields. As field diversity increases, SO\_ANN begins to
outperform CF, becoming the preferred approach at 10 fields and showing
a substantial further accuracy gain at 20 fields.

\begin{table}
\caption*{
{\large Best models by field count} \\ 
{\small Ordered by lowest mean RMSE}
} 
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcr}
\toprule
Model & \# fields & mean RMSE \\ 
\midrule\addlinespace[2.5pt]
SO\_ANN & 20 fields & 8.545 \\ 
SO\_ANN & 10 fields & 11.233 \\ 
CF & 5 fields & 12.115 \\ 
CF & 3 fields & 13.092 \\ 
CF & 1 field & 14.732 \\ 
\bottomrule
\end{tabular*}
\end{table}

Figure~\ref{fig-yield-res-combined} displays predicted and true yield
response functions for the best-performing models at each training size.
For the CF, shown for 1, 3, and 5 fields, the estimated curves closely
track the overall shape of the true response but exhibit a slight upward
bias in predicted yield levels. The general curvature, including
diminishing returns at higher nitrogen rates, is captured consistently
even under data-constrained conditions. As more fields are incorporated,
the CF predictions become smoother and more aligned with the true
function, reflecting improved generalization.

For the SO\_ANN, shown for 10 and 20 fields, the estimated response
matches the true curve more closely in both slope and curvature,
particularly at higher nitrogen levels where CF tended to flatten. At 20
fields, SO\_ANN produces the most accurate approximation overall, with
prediction and truth nearly coincident across the nitrogen range. These
differences visually reinforce the earlier quantitative results: CF
offers strong robustness with sparse multi-field data, while SO\_ANN
increasingly excels as richer spatial heterogeneity becomes available,
enabling superior recovery of the underlying agronomic yield response
function.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-yield-res-combined-1.pdf}

}

\caption{\label{fig-yield-res-combined}Yield response functions for CF
and SO\_ANN models across different field counts.}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This study evaluates alternative causal-ML strategies for estimating
heterogeneous yield response to a continuous nitrogen treatment in an
OFPE setting. CF consistently attains low and tightly concentrated EONR
errors with only 1--5 training fields, indicating strong small-sample
stability (see Figure~\ref{fig-rmse-hist} and
Figure~\ref{fig-avg-performance}). SO\_ANN improves monotonically with
additional fields and ultimately dominates once ≥10 fields are
available, achieving the lowest mean RMSE overall (e.g., 11.2 at 10
fields and 8.5 at 20 fields). These complementary strengths that CF is
preferred under data-constrained conditions, whereas SO\_ANN becomes the
method of choice as multi-field datasets accumulate (see
\textbf{?@tbl-best\_model\_by\_field\_count}).

Methodologically, the contrast between single-orthogonalization and
double-orthogonalization is central. For continuous treatments, the
double-orthogonal residualization of the treatment does not generally
preserve the relevant heterogeneous response structure. Our simulations
reflect this: DO\_ANN performs poorly and does not benefit from larger
samples. In contrast, single-orthogonalization---residualizing only the
outcome to remove baseline variation from covariates \(W\)---retains the
full interaction between \(X\) and \(N\), allowing the final learner to
recover the nonlinearity and heterogeneity of the response function. The
extended CF strikes a different balance: by expressing \(N\) through a
spline basis and orthogonalizing the basis components, it restores a
linear-in-parameters score that CF can learn reliably.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

By combining orthogonalization with flexible final-stage learners, we
demonstrate how to preserve the strengths of causal inference while
accommodating continuous treatments and agronomic heterogeneity.

Across 500 simulated test fields, two approaches consistently dominate:
the extended causal forest and the single-orthogonalized neural network.
CF is highly robust in small-sample environments, achieving strong
accuracy even with training data from just one to five fields. As the
number of training fields increases, SO\_ANN increasingly improves as a
result of spatial and agronomic diversity, ultimately delivering the
best predictive accuracy of all methods once ten or more fields are
available. In contrast, non-orthogonalized baselines require larger
samples to become competitive and the doubly-orthogonalized ANN performs
poorly in all conditions due to the incompatibility between
double-orthogonalization and continuous treatments. Methodologically,
our results highlight the importance of aligning the orthogonalization
strategy with the structure of the treatment variable in causal ML
applications.

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-barbosa2020risk}{}}%
Barbosa, Alexandre, Naira Hovakimyan, and Nicolas F Martin. 2020.
{``Risk-Averse Optimization of Crop Inputs Using a Deep Ensemble of
Convolutional Neural Networks.''} \emph{Computers and Electronics in
Agriculture} 178: 105785.

\leavevmode\vadjust pre{\hypertarget{ref-barbosa2020modeling}{}}%
Barbosa, Alexandre, Rodrigo Trevisan, Naira Hovakimyan, and Nicolas F
Martin. 2020. {``Modeling Yield Response to Crop Management Using
Convolutional Neural Networks.''} \emph{Computers and Electronics in
Agriculture} 170: 105197.

\leavevmode\vadjust pre{\hypertarget{ref-chernozhukov2018double}{}}%
Chernozhukov, Victor, Denis Chetverikov, Mert Demirer, Esther Duflo,
Christian Hansen, Whitney Newey, and James Robins. 2018.
{``Double/Debiased Machine Learning for Treatment and Structural
Parameters.''} Oxford University Press Oxford, UK.

\leavevmode\vadjust pre{\hypertarget{ref-gardner2021economic}{}}%
Gardner, Grant, Taro Mieno, and David S Bullock. 2021. {``An Economic
Evaluation of Site-Specific Input Application r x Maps: Evaluation
Framework and Case Study.''} \emph{Precision Agriculture} 22: 1304--16.

\leavevmode\vadjust pre{\hypertarget{ref-kakimoto2022causal}{}}%
Kakimoto, Shunkei, Taro Mieno, Takashi ST Tanaka, and David S Bullock.
2022. {``Causal Forest Approach for Site-Specific Input Management via
on-Farm Precision Experimentation.''} \emph{Computers and Electronics in
Agriculture} 199: 107164.

\leavevmode\vadjust pre{\hypertarget{ref-krause2020random}{}}%
Krause, M, Savanna Crossman, Todd DuMond, Rodman Lott, Jason Swede,
Scott Arliss, Ron Robbins, Daniel Ochs, and Michael A Gore. 2020.
{``Random Forest Regression for Optimizing Variable Planting Rates for
Corn and Soybean Using Topographical and Soil Data.''}

\leavevmode\vadjust pre{\hypertarget{ref-mackey2018orthogonal}{}}%
Mackey, Lester, Vasilis Syrgkanis, and Ilias Zadik. 2018. {``Orthogonal
Machine Learning: Power and Limitations.''} In \emph{International
Conference on Machine Learning}, 3375--83. PMLR.

\end{CSLReferences}

\end{document}

### 2025-11-17 codes/R/Add_mean_match_to_YRF.qmd
comments on the manuscript added

@@
 ```
+
+# CF
+
+```{r}
+# ==========================================================
+# Mean-match predicted yield ONLY for CF (change # fields)
+# ==========================================================
+suppressPackageStartupMessages({
+  library(here)
+  library(readr)
+  library(dplyr)
+  library(stringr)
+  library(purrr)
+})
+
+# Base directory
+base_dir <- here::here("results", "yield_response_function_for_one_iteration")
+target_dir <- file.path(base_dir, "YRF_CF_ten_fields")
+
+if (!dir.exists(target_dir)) {
+  stop("Target directory not found: ", target_dir, call. = FALSE)
+}
+
+# --- mean-match function (global shift per file) ---
+mean_match_file <- function(csv_path) {
+  df <- suppressMessages(readr::read_csv(csv_path, show_col_types = FALSE))
+
+  if (!all(c("pred_yield", "true_yield") %in% names(df))) {
+    message("Skipping (missing pred_yield/true_yield): ", csv_path)
+    return(invisible(FALSE))
+  }
+
+  delta <- mean(df$true_yield, na.rm = TRUE) - mean(df$pred_yield, na.rm = TRUE)
+
+  df <- df %>%
+    mutate(pred_yield_mm = pred_yield + delta) %>%
+    relocate(pred_yield_mm, .after = pred_yield)
+
+  readr::write_csv(df, csv_path)
+  message(sprintf("Updated: %s  (delta = %.6f)", csv_path, delta))
+  invisible(TRUE)
+}
+
+# --- find CF yield_response_*.csv files and apply mean-match ---
+csvs <- list.files(target_dir, pattern = "^yield_response_.*\\.csv$", full.names = TRUE)
+if (!length(csvs)) {
+  message("No yield_response_*.csv in: ", target_dir)
+} else {
+  walk(csvs, mean_match_file)
+}
+
+```

### 2025-11-17 codes/R/CF_YRF.qmd
comments on the manuscript added

@@
 ```{r}
-cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_10fields.rds"))
+cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_1fields.rds"))
 ```
@@
 train_data <- cf_results$train_data
-repsponse_data <- cf_results$repsponse_data
+response_data <- cf_results$response_data
 
@@
   # pick a few ids
-  check_id <- test_data[order(Nk), aunit_id] %>%
-    .[seq(1, nrow(test_data), length = 5)] %>%
-    .[-1]
+  # check_id <- test_data[order(Nk), aunit_id] %>%
+  #   .[seq(1, nrow(test_data), length = 2)] %>%
+  #   .[-1]
+   # Make CF use the same aunit_id as SO_ANN visualization
+so_ann_csv <- here::here(
+  "results", "yield_response_function_for_one_iteration",
+  "YRF_SO_ANN_1_fields",   
+  "yield_response_1.csv"
+)
+
+if (file.exists(so_ann_csv)) {
+  so_df <- readr::read_csv(so_ann_csv, show_col_types = FALSE)
+  # choose the same ID that SO_ANN used (the first unique aunit_id in its CSV)
+  shared_id <- unique(so_df$aunit_id)[1]
+
+  # ensure it exists in CF test_data
+  if (shared_id %in% test_data$aunit_id) {
+    check_id <- shared_id
+  } else {
+    # fallback: use the first available CF id
+    check_id <- unique(test_data$aunit_id)[1]
+  }
+} else {
+  # fallback if file missing
+  check_id <- unique(test_data$aunit_id)[1]
+}
+
+message("Using aunit_id = ", check_id, " to match SO_ANN")
 
@@
   # plot
+
   ggplot() +
@@
   xlab("Nitrogen Application Rate (lb/acre)") +
-    ggtitle(paste("Yield Response Function -", label))
+  ggtitle(paste("CF -", label))
+
 }
 
-file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
+# file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
+# file_paths <- here::here(
+#   "results/yield_response_function_for_one_iteration",
+#   paste0("YRF_CF_", file_labels, ".rds")
+# )
+
+field_counts <- c(1, 3, 5, 10, 20)
+
 file_paths <- here::here(
-  "Results/yield_response_function_for_one_iteration",
-  paste0("YRF_CF_", file_labels, ".rds")
+  "results/yield_response_function_for_one_iteration",
+  sprintf("YRF_CF_%dfields.rds", field_counts)
 )
+
+file_labels <- sprintf("%d fields", field_counts)
 
@@
 #   all_plots,
-#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds")
+#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds")
 # )
@@
 
-all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF.rds"))
+all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds"))
 

### 2025-11-17 codes/R/GRF_CF_Loop.qmd
comments on the manuscript added

--- 2025-09-30 codes/R/CF_Loop.qmd
@@
 ---
-title: "CF Loop"
+title: "GRF CF Loop"
 format:
@@
   MKL_NUM_THREADS = "1",
-  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
+  VECLIB_MAXIMUM_THREADS = "1" 
 )
-
-#--- python functions ---#
-use_condaenv("cf_conda311", required = TRUE)
-
-# use_condaenv("cf_conda311",
-#              conda = "/Users/monamousavi/miniforge3/bin/conda",
-#              required = TRUE)
-
-py_config()
-py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
 
@@
 
-#--- python functions ---#
-source_python(here::here("codes/Python/run_CF_c.py"))
 ```
@@
 ## load data
-
 ```{r}
@@
 # how many fields to train on?
-n_fields <- 20
+n_fields <- 5
 
@@
 output_dir <- here(
-  "results",
+  "results/GRF_CF",
   sprintf("CF_outcome_%d_fields", n_fields)
@@
 # ─── LOAD SPLITS ────────────────────────────────────────────────────────────────
-
 train_test_splits <- read_csv(splits_file)
@@
 # CF loop
-
 ```{r}
@@
 ```
+
+```{r}
+
+# ── EONR subfolder ──────────────────────────────────────────────────────────────
+eonr_dir <- file.path(output_dir, "eonr")
+if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)
+
+## ── Loop over test sims ───────────────────────────────────────────────────────
+#for (i in seq_len(1))
+#for (i in seq_len(nrow(train_test_splits)))
+for (i in seq_len(1)) {
+  test_sim_id <- train_test_splits$test_id[i]
+  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))
+
+  if (file.exists(out_fn)) {
+    message("Skipping sim ", test_sim_id, " (already saved)")
+    next
+  }
+
+  tryCatch({
+    ## 1) train/test IDs
+    split_row     <- train_test_splits[i, ]
+    train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
+    sim_id_ls     <- c(train_sim_ids, test_sim_id)
+
+    ## 2) train & test data
+    data <- raw_sim_data$reg_data[[1]] %>%
+      .[sim %in% sim_id_ls, ] %>%
+      .[, .(sim, data)] %>%
+      tidyr::unnest(cols = c(data)) %>%
+      as.data.table()
+
+    train_data <- data[sim != test_sim_id, ]
+    test_data  <- raw_sim_data$reg_data[[1]] %>%
+      .[sim == test_sim_id, ] %>%
+      .[, .(sim, data)] %>%
+      tidyr::unnest(cols = c(data)) %>%
+      as.data.table()
+
+    ## 3) CF variables
+    x_vars <- c("Nk", "plateau", "b0")
+    T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
+    Y      <- train_data[, yield]
+    X      <- as.matrix(train_data[, ..x_vars])
+    W      <- X
+    X_test <- as.matrix(test_data[, ..x_vars])
+
+    ## 4) cf in GRF
+    cf_fit <- causal_forest(
+      X = X,
+      Y = Y,
+      W = T_info$T_sp,     # spline
+      num.trees = 2000,
+      min.node.size = 10,
+      seed = 3747823
+    )
+
+    # Estimate treatment effects for test data
+    te_hat_cf <- predict(cf_fit, newdata = X_test)$predictions
+
+    te_info <- data.table(aunit_id = test_data$aunit_id, est = te_hat_cf)
+
+    ## 5) Response curves
+    T_seq <- train_data[, seq(min(N), max(N), length = 200)]
+    response_data <- find_response_semi(T_seq, T_info, te_info)
+
+    ## 6) Site-specific EONR
+    pCorn <- 6.25 / 25.4
+    pN    <- 1 / 0.453592
+
+    ss_eonr <- response_data %>%
+      .[, profit := est * pCorn - pN * T] %>%
+      .[, .SD[which.max(profit), ], by = aunit_id] %>%
+      .[, .(aunit_id, opt_N_hat = T)]
+
+    true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
+      .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
+      .[, opt_N := pmin(Nk, opt_N)] %>%
+      .[, opt_N := pmax(0,  opt_N)] %>%
+      .[, .(aunit_id, opt_N)]
+
+    combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
+    corr_val <- combined_eonr[, cor(opt_N, opt_N_hat)]
+
+    ## 7) SAVE EONR
+    e <- as.data.table(combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
+    e[, sim := as.integer(test_sim_id)]
+    data.table::setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))
+
+    fn_eonr   <- file.path(eonr_dir, sprintf("sim_%03d_eonr.rds", as.integer(test_sim_id)))
+    tmp_eonr  <- paste0(fn_eonr, ".tmp")
+    saveRDS(e, tmp_eonr, compress = "xz")
+    file.rename(tmp_eonr, fn_eonr)
+
+    csv_path <- file.path(eonr_dir, "eonr_progress.csv")
+    data.table::fwrite(e, csv_path, append = file.exists(csv_path))
+
+    ## 8) Save results object (optional)
+    # res <- list(
+    #   cf_fit        = cf_fit,
+    #   te_info       = te_info,
+    #   T_seq         = T_seq,
+    #   response_data = response_data,
+    #   test_data     = test_data,
+    #   train_data    = train_data,
+    #   combined_eonr = combined_eonr,
+    #   corr          = corr_val
+    # )
+    # tmp_fn <- paste0(out_fn, ".tmp")
+    # saveRDS(res, tmp_fn, compress = "xz")
+    # file.rename(tmp_fn, out_fn)
+
+    ## 9) Log progress
+    prog <- data.frame(
+      test_sim_id = test_sim_id,
+      corr        = corr_val,
+      n_train     = nrow(train_data),
+      saved_file  = out_fn,
+      eonr_file   = fn_eonr,
+      ts          = as.character(Sys.time())
+    )
+    write.table(
+      prog,
+      file      = progress_csv,
+      sep       = ",",
+      row.names = FALSE,
+      col.names = !file.exists(progress_csv),
+      append    = TRUE
+    )
+
+    message("Finished sim ", test_sim_id,
+            " | corr = ", round(corr_val, 3))
+
+    # Free RAM
+    rm(cf_fit, te_info, response_data, train_data, test_data, e)
+    if (i %% 10 == 0) gc()
+  },
+  error = function(e) {
+    message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
+    err_row <- data.frame(
+      test_sim_id = test_sim_id,
+      error       = conditionMessage(e),
+      ts          = as.character(Sys.time())
+    )
+    write.table(
+      err_row,
+      file      = file.path(output_dir, "errors.csv"),
+      sep       = ",",
+      row.names = FALSE,
+      col.names = !file.exists(file.path(output_dir, "errors.csv")),
+      append    = TRUE
+    )
+  })
+}
+
+```

### 2025-11-17 codes/Python/Functions/CF_savedT.py

@@ -54,7 +54,7 @@ def train_cf_savedT_model_from_splits(data_2nd_stage, evall_N_seq, split_csv_pat
         Y = train_df["yield"].to_numpy()
         X = train_df[["Nk", "plateau", "b0"]].to_numpy()
         W = X.copy()
-        T = train_df[["T_1_tilde", "T_2_tilde", "T_3_tilde"]].to_numpy()
+        T = train_df[["T_1", "T_2", "T_3"]].to_numpy()
 
         # --- Causal Forest (orthogonalized, tuned) ---
         model_y = GradientBoostingRegressor(n_estimators=100, max_depth=3, min_samples_leaf=20)

### 2025-11-17 codes/Python/Functions/CF_spline.py

@@ -1,0 +1,130 @@
+import os
+import sys
+import numpy as np
+import pandas as pd
+from tqdm import tqdm
+from sklearn.preprocessing import StandardScaler
+from patsy import dmatrix  # for spline basis like mgcv::s()
+
+# make parent directory visible for import
+sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), "..")))
+from run_CF_c import run_CF_c_py
+import torch
+
+
+# === MODEL: Causal Forest with spline-transformed treatment (normalized per field) === #
+
+def train_CF_spline_from_splits(data_2nd_stage, evall_N_seq, split_csv_path, folder_name, device):
+    # --- Resolve project root automatically ---
+# Finds the "CF_Continuous" folder no matter where the script is launched from
+    project_root = os.path.abspath(
+    os.path.join(os.path.dirname(__file__), "..", "..", "..")
+)
+    results_dir = os.path.join(project_root, "results")
+    os.makedirs(results_dir, exist_ok=True)
+
+# --- Create subfolder for model output ---
+    output_folder = os.path.join(results_dir, folder_name)
+    os.makedirs(output_folder, exist_ok=True)
+
+
+    splits_df = pd.read_csv(split_csv_path)
+    p_corn = 6.25 / 25.4  # $/kg
+    p_N = 1 / 0.453592    # $/kg
+
+    # For quick test only 1 iteration:
+    #row = splits_df[splits_df['test_id'] == 1].iloc[0:1].copy()
+    #for _, row in row.iterrows():
+    for _, row in tqdm(splits_df.iterrows(), total=len(splits_df)):
+        test_id = row["test_id"]
+        train_ids = row[[c for c in row.index if c.startswith("train_")]].dropna().values
+
+        # ---- Split data ----
+        train_df = data_2nd_stage[data_2nd_stage["sim"].isin(train_ids)].reset_index(drop=True)
+        test_df = data_2nd_stage[data_2nd_stage["sim"] == test_id].reset_index(drop=True)
+
+        # ---- Normalize N per field to [0,1] ----
+        mm = train_df.groupby("sim")["N"].agg(Nmin="min", Nmax="max").reset_index()
+        train_df = train_df.merge(mm, on="sim", how="left")
+        test_df = test_df.merge(mm, on="sim", how="left")  # may be missing if test sim not in train set
+
[…]

### 2025-11-17 codes/Python/Functions/CF_test.py

@@ -1,0 +1,193 @@
+import os
+import sys
+import numpy as np
+import pandas as pd
+from tqdm import tqdm
+from patsy import dmatrix, build_design_matrices
+from sklearn.preprocessing import StandardScaler
+
+# make parent directory visible for import
+sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), "..")))
+from run_CF_c import run_CF_c_py
+
+# ==============================================================
+# CONFIGURATION
+# ==============================================================
+k_splines = 4
+spline_formula = f"cr(N, df={k_splines}) - 1"
+p_corn = 6.25 / 25.4  # $/kg
+p_N = 1 / 0.453592    # $/kg
+
+# ==============================================================
+# Prepare spline basis (same as mgcv s(N, k=4))
+# ==============================================================
+from patsy import cr
+def prepare_T_mat_py(train_df, test_df):
+    N_train = train_df["N"].to_numpy()
+    N_min, N_max = float(np.min(N_train)), float(np.max(N_train))
+
+    T_train_df = dmatrix(spline_formula, {"N": N_train}, return_type="dataframe")
+    design_info = T_train_df.design_info
+
+    N_test_clip = np.clip(test_df["N"].to_numpy(), N_min, N_max)
+    T_test_mat = build_design_matrices([design_info], {"N": N_test_clip})[0]
+    T_test_df = pd.DataFrame(np.asarray(T_test_mat), columns=T_train_df.columns)
+
+    col_names = [f"T_{i+1}" for i in range(T_train_df.shape[1])]
+    T_train_df.columns = col_names
+    T_test_df.columns = col_names
+
+    return T_train_df, T_test_df, design_info, (N_min, N_max)
+
+# ==============================================================
+# TRAIN + TEST LOOP (Mimics R “CF Loop”)
+# ==============================================================
+def train_CF_R_like(data_2nd_stage, split_csv_path, folder_name):
+    project_root = os.path.abspath(os.path.join(os.path.dirname(__file__), "..", "..", ".."))
+    results_dir = os.path.join(project_root, "results")
+    os.makedirs(results_dir, exist_ok=True)
+
+    output_folder = os.path.join(results_dir, folder_name)
+    os.makedirs(output_folder, exist_ok=True)
+    progress_csv = os.path.join(output_folder, "progress.csv")
+
+    splits_df = pd.read_csv(split_csv_path)
[…]

### 2025-11-17 codes/Python/Functions/main_simulation_run.py

@@ -5,6 +5,8 @@ from half_RRF import run_model as run_halfrrf
 from NO_RF import run_model as run_NO_RF
 from CF_python import run_model as run_CF_python
 from CF_savedT import run_model as run_CF_savedT
+from CF_spline import run_model as run_CF_spline
+from CF_test import run_model as run_test
 from Generate_train_test_splits import data_2nd_stage, evall_N_seq, device
 
 if __name__ == "__main__":
@@ -58,10 +60,23 @@ if __name__ == "__main__":
 
  
 # === CF using saved spline columns (T_1, T_2, T_3, …) ===
-    run_CF_savedT("CF_savedT", 1, data_2nd_stage, evall_N_seq)
+    #run_CF_savedT("CF_savedT", 1, data_2nd_stage, evall_N_seq)
     #run_CF_savedT("CF_savedT", 3, data_2nd_stage, evall_N_seq)
     #run_CF_savedT("CF_savedT", 5, data_2nd_stage, evall_N_seq)
    #run_CF_savedT("CF_savedT", 10, data_2nd_stage, evall_N_seq)
     #run_CF_savedT("CF_savedT", 20, data_2nd_stage, evall_N_seq)
 
-  
+
+# === CF spline ===
+   #run_CF_spline("CF_spline", 1, data_2nd_stage, evall_N_seq, device)
+   #run_CF_spline("CF_spline", 3, data_2nd_stage, evall_N_seq, device)
+   #run_CF_spline("CF_spline", 5, data_2nd_stage, evall_N_seq, device)
+   #run_CF_spline("CF_spline", 10, data_2nd_stage, evall_N_seq, device)
+   #run_CF_spline("CF_spline", 20, data_2nd_stage, evall_N_seq, device)
+
+  # === TEST (pygam-based spline computation, mgcv equivalent) ===
+   #run_test("test", 1, data_2nd_stage, evall_N_seq, device)
+   # run_test("test", 3, data_2nd_stage, evall_N_seq, device)
+   # run_test("test", 5, data_2nd_stage, evall_N_seq, device)
+  # run_test("test", 10, data_2nd_stage, evall_N_seq, device)
+  run_test("test", 20, data_2nd_stage, evall_N_seq, device)

### 2025-11-17 codes/Python/cf_loop_python.py

@@ -1,0 +1,328 @@
+#!/usr/bin/env python3
+"""
+CF Loop (Python port)
+
+This script replicates the R "CF_Loop" workflow in Python using:
+- pandas / numpy for data handling
+- sklearn SplineTransformer to build a fixed spline basis for N (fit on train only)
+- econml.dml.CausalForestDML for treatment effect estimation
+
+It expects:
+- Train/test split CSV at data/train_test_split/train_test_splits_{n_fields}fields.csv
+- A flat table of observations with columns at least:
+  sim, aunit_id, N, yield, Nk, plateau, b0, b1, b2
+  available at either data/reg_data.parquet OR data/reg_data.csv
+
+Outputs per test sim:
+- results/CF_Loop_p_{n_fields}_fields/sim_{test_sim_id}_eonr.csv
+- results/CF_Loop_p_{n_fields}_fields/progress.csv (appended)
+- results/CF_Loop_p_{n_fields}_fields/errors.csv (if any)
+
+Notes:
+- Thread env vars are limited similar to the R script to avoid BLAS over‑subscription.
+- The spline basis is *fit on the current TRAIN set* and then applied to both train and test.
+- If you already have a saved basis you want to reuse across runs, persist the fitted
+  SplineTransformer with joblib and load it instead of re-fitting (marked below).
+"""
+
+from __future__ import annotations
+import os
+import sys
+import time
+from pathlib import Path
+from dataclasses import dataclass
+from typing import List, Tuple
+
+# --- Limit math library threads (macOS/BLAS friendliness) ---
+os.environ.setdefault("OMP_NUM_THREADS", "1")
+os.environ.setdefault("OPENBLAS_NUM_THREADS", "1")
+os.environ.setdefault("MKL_NUM_THREADS", "1")
+os.environ.setdefault("VECLIB_MAXIMUM_THREADS", "1")
+os.environ.setdefault("KMP_AFFINITY", "disabled")
+os.environ.setdefault("KMP_DUPLICATE_LIB_OK", "TRUE")
+
+import numpy as np
+import pandas as pd
+
+from sklearn.preprocessing import SplineTransformer
+from sklearn.ensemble import GradientBoostingRegressor
+from sklearn.multioutput import MultiOutputRegressor
+
+try:
+    from econml.dml import CausalForestDML
+except Exception as e:  # helpful import message
+    raise SystemExit(
[…]

### 2025-11-17 codes/R/Utility.R

@@ -17,14 +17,16 @@ get_te <- function(trained_model, test_data, x_vars, id_var = "aunit_id") {
   return(list(te_hat = data.table(te_hat), id_data = test_data[, ..id_var]))
 }
 
-
+#______________________________________________
+#---- If I do not save splines use it:-------- 
+#______________________________________________
 find_response_semi <- function(T_seq, T_info, te_info) {
   eval_T <- data.table(T = T_seq) %>%
     setnames("T", T_info$T_var_name) %>%
     predict(T_info$gam_setup, newdata = ., type = "lpmatrix") %>%
     .[, -1] %>%
     data.table()
-  
+
   curv_data <- as.matrix(te_info$te_hat) %*% t(as.matrix(eval_T)) %>%
     data.table() %>%
     setnames(names(.), as.character(T_seq)) %>%
@@ -32,12 +34,44 @@ find_response_semi <- function(T_seq, T_info, te_info) {
     melt(id.var = "id") %>%
     setnames(c("variable", "value"), c("T", "est")) %>%
     .[, T := as.numeric(as.character(T))]
-  
+
   id_data <- te_info$id_data[, id := 1:.N]
   final_data <- id_data[curv_data, on = "id"][, id := NULL]
   return(final_data)
 }
 
+# otherwise use it:
+
+# find_response_semi <- function(T_seq, T_vars, train_data, te_info) {
+#   # T_vars: vector of spline column names, e.g. c("T_1","T_2","T_3")
+#   # train_data: training data that already has those columns
+#   # te_info: output from get_te()
+#   
+#   # extract spline basis matrix from training data
+#   T_mat <- as.matrix(train_data[, ..T_vars])
+#   
+#   # create spline basis interpolation function for each column
+#   interp_list <- lapply(1:ncol(T_mat), function(j) {
+#     approxfun(train_data$N, T_mat[, j], rule = 2)
+#   })
+#   
+#   # evaluate the basis at T_seq values
+#   eval_T <- sapply(interp_list, function(f) f(T_seq))
+#   colnames(eval_T) <- T_vars
+#   
+#   # estimate predicted yield responses
+#   curv_data <- as.matrix(te_info$te_hat) %*% t(eval_T)
+#   curv_data <- data.table(curv_data)
+#   setnames(curv_data, as.character(T_seq))
+#   
+#   curv_data <- curv_data[, id := 1:.N] %>%
[…]

### 2025-11-18 Writing/manuscript_CF_continuous.qmd
Working on comments

@@
 
+new
+
+On-farm precision experiments (OFPE) have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management (@bullock2019data; @lacoste2022farm).
+Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions, enabling estimation of yield response functions and the identification of economically optimal nitrogen rates.
+In recent years, prediction-oriented machine learning (ML) methods have been widely adopted in this context to model yield as a function of nitrogen and environmental covariates (@krause2020random; @barbosa2020modeling; @barbosa2020risk; @gardner2021economic).
+However, recent evidence suggests that prediction-oriented ML approaches, despite often achieving high yield-prediction accuracy, do not necessarily recover the underlying yield response to nitrogen.
+@tanaka2024can show that fertilizer rate recommendations derived from predictive ML models are highly sensitive to algorithm choice and covariate selection, leading to substantial variability in estimated optimal rates.
+These findings indicate a critical difference between predicting yield levels and estimating the causal yield response that determines economically optimal nitrogen rates, motivating the use of methods that explicitly estimate causal effects of nitrogen on yield.
+
 ::: {.column-margin}
@@
 \begin{align*}
-y = f(X, N) + g(W) + v
+y = f(X, N) + g(X,W) + v
 \end{align*}
@@
 
-In the single-orthogonalized framework only the baseline component $g(W)$ is partialed out before the final-stage estimation; the continuous treatment $N$ and heterogeneity-inducing covariates $X$ remain in their original form.
-Formally, starting from Model 3, we first estimate $g(W)=\mathbb{E}[y \mid W]$ using a cross-fitted random forest and form the residualized outcome
+In the single-orthogonalized framework only the baseline component $g(X,W)$ is partialed out before the final-stage estimation; the continuous treatment $N$ and heterogeneity-inducing covariates $X$ remain in their original form.
+Formally, starting from Model 3, we first estimate $g(X,W)=\mathbb{E}[y \mid X,W]$ using a cross-fitted random forest and form the residualized outcome
 
@@
 \begin{align*}
-\tilde{y}=y-\widehat{g}(W)
+\tilde{y}=y-\widehat{g}(X,W)
 \end{align*}
@@
 
-To adapt double-orthogonalization to the model of interest (Model 3), we estimate the heterogeneous response $f(X,N)$ by expanding the continuous treatment $N$ in a set of spline basis functions and allowing their coefficients to vary with the heterogeneity-inducing covariates $X$:
+To adapt double-orthogonalization to the model of interest (Model 3), we estimate the heterogeneous response $f(X,N)$ by expanding the continuous treatment $N$ in a set of spline basis functions and allowing their coefficients to vary with the heterogeneity-inducing covariates $X$.
+
+In spline regression, a single variable can be decomposed into multiple basis functions in a linear-in-parameter manner.
+For example consider the following general model, where the non-linear impact of $T$ is represented by $f(T)$.
+
+$$
+y = f(T) + v 
+$$
+
+Non-linearity in $f(T)$ can be captured by a linear combination of basis functions as follows:
+
+$$
+f(T) = \alpha_1 \cdot S_1(T) + \alpha_2 \cdot S_2(T) \dots, + \alpha_K \cdot S_K(T) = \sum_{k=1}^K \alpha_k \cdot S_k(T)
+$$
+
+where $\alpha_k$ is the coefficient to be estimated and $S_k(T)$ is the value of the basis function $S_k(\cdot)$ evaluated at $T$.
+By changing the value of $\alpha_k$, $\sum_{k=1}^K \alpha_k S_k(T)$ can flexibly approximate various non-linear functions that cannot be captured by commonly used functional forms.
+In spline regression, the weights (or coefficients) are obtained by regression with the spline bases included as covariates.
+
+```{r fig-spline-examples, fig.cap="An example of spline regression fit of the yield response curved and weighted basis functions"}
+knitr::include_graphics(here::here("writing/Figures/g_tsp_fit.pdf"))
+```
+
+@fig-spline-examples shows the fitted line (red line), which is the sum of the six weighted (multiplied by coefficients) thin-plate splines (dotted lines).
+As can be seen, the fitted line closely traces the the true yield response function (blue line).
+
+Since this modeling approach is linear-in-parameter, it can be readily used inside the CF framework.
+Instead of a constant as the weight (coefficient) for each spline basis function, the weight is made a function of observables, $X$, as follows:
 
@@
 \begin{align*}
-y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
+y=\sum_{k=1}^K f_k(X) S_k(N)+g(X, W)+v
 \end{align*}
@@
 Residualizing each basis component and the outcome with respect to $(X,W)$ then yields a clean decomposition analogous to the linear-in-$T$ case.
-Let $\widetilde{S}_k(N)=S_k(N)-\mathbb{E}\left[S_k(N) \mid X, W\right]$.
-Then
+
+Let $\widetilde{S}_k(N)=S_k(N)-\mathbb{E}\left[S_k(N) \mid X, W\right]$. Then
 
@@
 \begin{align*}
-\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
+\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K f_k(X) \widetilde{S}_k(N)+v,
 \end{align*}
@@
 
-so causal forest can nonparametrically estimate each heterogeneous weight $\theta_k(X)$  by relating the residualized outcome to the residualized spline bases.
-This Neyman-orthogonal (R-learner) style decomposition preserves the ability to capture complex, heterogeneous, and nonlinear treatment effects while inheriting the robustness of double-orthogonalization: small errors in the nuisance fits for $g(X,W)$ or the conditional means of the bases do not induce first-order bias in $\theta_k(X)$.
-The estimated $\theta_k(X)$ then combine to recover the full heterogeneous response function and, by differencing across candidate $N$ values, site-specific treatment effects and economically optimal nitrogen rates.
+so causal forest can nonparametrically estimate each heterogeneous weight $f_k(X)$  by relating the residualized outcome to the residualized spline bases.
+This Neyman-orthogonal (R-learner) style decomposition preserves the ability to capture complex, heterogeneous, and nonlinear treatment effects while inheriting the robustness of double-orthogonalization: small errors in the nuisance fits for $g(X,W)$ or the conditional means of the bases do not induce first-order bias in $f_k(X)$.
+The estimated $f_k(X)$ then combine to recover the full heterogeneous response function and, by differencing across candidate $N$ values, site-specific treatment effects and economically optimal nitrogen rates.
 

### 2025-11-18 writing/meeting_Sep_2nd.qmd
Rename folder to writing (correct lowercase)

---
title: "Meeting Sep 2nd"
format:
  html: default
execute:
  echo: false
bibliography: references.bib
---

# paper citation

In OFPE with continuous nitrogen rates pooled across fields, much of the literature either predicts levels and optimizes (@barbosa2020risk; @barbosa2020modeling; @krause2020random; @gardner2021economic) or treats the rate as discrete treatment levels (@kakimoto2022causal).

@krause2020random:

- Objective:

They checked to see if  RF regression can be used to optimize variable planting rates for corn and soybean and then used this information to design site-specific variable rate planting strategies.

- Data:

On-farm field trials conducted in 27 sites across New York from 2014–2018 (57 site-years: 32 corn, 25 soybean).
Design: Each site planted with four planting rates (corn: 66,718–103,784 seeds/ha; soybean: 197,684–494,210 seeds/ha).

They get:
yield maps 
as-applied planting rate maps
soil data 
weather data

- Method:

RF:
They fit models for each site-year 
Predictors: planting rate, soil/topography, weather.
Then they predicted yield at each rate per grid cell, then assigned the rate that maximized yield

- Conclusion:

Random forests performed well at predicting yield  but did not performed good when applied to untested site-years.

@barbosa2020risk:

- Objective:

They wanted to predict corn yields using deep learning and then use those predictions to decide how much nitrogen and seed to apply in each part of the field.
They also add a risk-averse concept (income does not drop below a certain level).

- Data:

OFPE from 9 cornfields in Illinois, Ohio, Nebraska, and Kansas (2017 season).
Design: each field split into ~200 experimental strips, each with randomized N and seeding rates.

Variables:
N rate and seeding rate.
Field characteristics
Output: yield data.

- Method:

They used a CNN (as an ensemble) to predict a distribution (mean + variance = yield with uncertainty).
Then found N and seed maps that maximize expected profit while controlling risk.
They tested prediction accuracy with yield RMSE and reported improvements in expected profit compared to constant farmer rates.

- Conclusion:

Their risk-aware CNN ensemble can slightly raise expected farm profit and reducing excess N and seed use, but it still relies on yield prediction.

@barbosa2020modeling:

- Objective:

They test whether CNN can better capture spatial patterns in data to model crop yield response to nitrogen and seeding rates, compared to standard ML models.

- Data:

DIFM project, 2017.
9 corn fields in Illinois, Ohio, Nebraska, Kansas.
Design: OFPE with randomized nitrogen and seed rates.

Variables:
nitrogen rate, seed rate.
elevation, shallow soil electroconductivity, bare soil reflectance (satellite image).
Output: yield

- Method:

They trained CNN (4 variants) to predict yield
They Compared CNNs against baseline models: linear regression, random forest, support vector machine, fully connected NN and then used RMSE of yield prediction.

- Conclusion:

CNNs predict crop yields more accurately than other models by using spatial field patterns.

* An example of the new ML-prediction approach with OFPE data: focusing on yield prediction accuracy instead of validating input response and EONR.

@gardner2021economic

- Objective:

To develop a method to test if commercial site-specific prescription (Rx) maps for seed and nitrogen rates are actually economically optimal for farmers.

- Data:

cornfield, Ohio, 2018
OFPE design: 276 experimental plots, randomized across 16 treatment combos (4 N rates × 4 seed rates).
as-applied N and seed rates, yield data.

- Method:

They estimate yield response functions in zones using SCAM, then Compute profit functions for each zone and find EONR and EOSR for each zone by maximizing the profit function. And then compare the consultant’s Rx recommendations against the estimated optimal rates in terms of yield, input use, and profit.

- Conclusion:

The consultant’s Rx map was not economically optimal; it over-applied N in some zones and under-applied seed in others, causing profit losses, but the OFPE-based method showed more profitable input rates.

* It’s prediction-based because the method fits yield functions to the observed treatment points and then derives EONR from that fitted function, instead of generating and validating yield responses at unobserved N/seed values for each site.

# True yield response function

# === Quadratic-Plateau response

yield_opt= gen_yield_QP(b0, b1, b2, NK, opt_N)

gen_yield_QP <- function(b0, b1, b2, Nk, N) {
    yield <- (N < Nk) * (b0 + b1 * N + b2 * N^2) + (N >= Nk) * (b0 + b1 * Nk + b2 * Nk^2)
    return(yield)}

### 2025-11-18 writing/manuscript_CF_continuous.qmd
Rename folder to writing (correct lowercase)
addressing comments
Working on comments (still)

--- 2025-11-18 Writing/manuscript_CF_continuous.qmd
@@
 
-On-farm field experiments have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management (<span style = "color: blue;"> add citations here.
-e.g., Dave's paper in agronomy and Lacoste's nature paper </span>).
-Through on-farm agronomic trials, researchers can directly observe yield responses to varying nitrogen rates under real-world field conditions.
-Analyzing these data enables estimation of yield response functions and identification of site-specific economically optimal nitrogen rates.
-Recent advances in data analytics—particularly the widespread adoption of machine learning (ML) methods—have expanded the potential of on-farm experimentation.
-Two major methodological directions have emerged: (1) yield prediction–based methods, which recommend site-specific input rates by predicting yield as a function of nitrogen and environmental covariates (@krause2020random; @barbosa2020modeling; @barbosa2020risk; @gardner2021economic), and (2) causal machine learning (causal ML) methods (@kakimoto2022causal), which estimate the underlying causal relationship between nitrogen application and yield response.
-
-new
-
 On-farm precision experiments (OFPE) have emerged as a powerful approach to generate the data necessary for site-specific nitrogen management (@bullock2019data; @lacoste2022farm).
@@
 These findings indicate a critical difference between predicting yield levels and estimating the causal yield response that determines economically optimal nitrogen rates, motivating the use of methods that explicitly estimate causal effects of nitrogen on yield.
-
-::: {.column-margin}
-Right now, it is written as if prediction-oriented and causal ML have been used in parallel.
-But, @kakimoto2022causal is in response to the danger of applying simply applying prediction-oriented ML approaches.
-Modify the text to reflect that here.
-Cite [this](https://link.springer.com/article/10.1007/s11119-024-10136-x) to motivate the use of causal ML.
-:::
-<!--end of column-margin-->
 

### 2025-11-19 presentation/proposal/proposal.qmd
presentation for proposal defense added

--- 2025-11-16 Proposal_presentation/Chapters_1&2.qmd
@@
 ---
-title: "Chapter 1: A new model selection approach based on local economically optimal input rate"
+title: "Proposal Defense"
 author:
-  - name: Taro Mieno$^*$
   - name: Mona Mousavi$^*$
-  - name: David Bullock$^\dagger$
 institute:
   - $^*$ University of Nebraska Lincoln
-  - $^\dagger$ University of Illinois Urbana Champaign 
 date: 06/12/2025
 format:
+  revealjs: default
   beamer:
@@
 header-includes: 
-- \logo{\ifnum\thepage>1\includegraphics[width=1cm]{nebraska-n.jpg}\fi}
+logo: nebraska-n.jpg
 ---
@@
 
+# Chapter 1: A new model selection approach based on local economically optimal input rate
+
 # Background and Research Objectives
@@
 
-+ Variable-rate input management as a solution (accounts for spatial heterogeneity in crop input demand.
++ Variable-rate input management as a solution (accounts for spatial heterogeneity in crop input demand).
 
@@
 #| out-width: "95%"
-knitr::include_graphics("~/Desktop/g_field_map.png")
+knitr::include_graphics(
+  here::here("presentation/proposal/g_field_map.png")
+)
 ```
@@
 #| out-width: "100%"
-knitr::include_graphics("g_N.png")
+knitr::include_graphics(
+  here::here("presentation/proposal/g_N.png")
+)
 ```
@@
 #| out-width: "100%"
-knitr::include_graphics("/Users/monamousavi/Dropbox/Causal_climate/g_y.png")
+knitr::include_graphics(
+  here::here("presentation/proposal/g_y.png")
+)
 ```
@@
   4. estimate site-specific EONR for the **test** data based on the trained candidate model ($\hat{N}^*_{i, f,m}$). 
-  5. Calcurate the average of the site-specific EONR ($\hat{N}^*_{f,m}$)
-  6. Calcurate the RMSE of the estimated local EONR:
+  5. Calculate the average of the site-specific EONR ($\hat{N}^*_{f,m}$)
+  6. Calculate the RMSE of the estimated local EONR:
 
@@
 By combining orthogonalization with flexible final-stage learners, we demonstrate how to preserve the strengths of causal inference while accommodating continuous treatments and agronomic heterogeneity.
+
+# Chapter 3: Economic assessment of biochar as a means to enhance soil water holding capacity
+
+## Background
+
++ Improving soil health is considered to be an important means of enhancing crop production resilience to climate change
++ Soil water holding capacity (WHC) is an important property of soil. Higher soil WHC leads to
+  + withstanding droughts better
+  + less irrigation (conserving water resources)
++ Biochar is actively researched as a supplement to enhance soil WHC of production fields, but is expensive
+
+## Research question and objective
+
+:::{.callout-note title="Research Question"}
++ Is biochar economically viable?
++ If not, how much cost reduction in biochar needs to be achieved to make it economically viable?
+:::
+
+## Method
+
+1. Simulate corn production using APSIM (a crop growth model) at different values of soil WHC for
+   1. historical weather conditions 
+   2. weather scenarios under climate change
+2. Calculate profit
+3. Identify the value of soil WHC enhancement
+
+## Expected Outcomes

### 2025-11-19 writing/manuscript_CF_continuous.qmd
presentation for proposal defense added

@@
 
-The key objective of this article is to identify a causal machine learning approach capable of estimating heterogeneous treatment effects for yield response function estimation using data from on-farm trials, where the treatment variable—nitrogen rate—is continuous and randomized.
-We investigate two complementary approaches: a Single-Orthogonalized Neural Network (SO-ANN), designed to flexibly capture nonlinear treatment–response relationships, and an extended Causal Forest (CF) that incorporates smoothing splines to model continuous treatments.
-Both methods aim to preserve the strengths of causal ML—nonparametric modeling of complex interactions without restrictive assumptions.
-Their performance is evaluated through Monte Carlo simulations that mimic typical OFPE conditions and crop yield responses to nitrogen.
-
-::: {.column-margin}
-<span style = "color: blue;"> We need to emphasize that we are proposing approaches that have not been tried anywhere: SO-NN and semi-parametric-CF.
-They are the main approaches.
-We then have other contrasting approaches to compare against.
-</span>
-:::
-<!--end of column-margin-->
+The objective of this study is to develop and evaluate new causal machine-learning approaches capable of estimating heterogeneous treatment effects for yield response function estimation using data from on-farm trials, where the treatment variable—nitrogen rate—is continuous and randomized.
+We introduce two methods: a Single-Orthogonalized Neural Network (SO-NN), which isolates the treatment–response component by removing variation in yield explained solely by field conditions, and a semi-parametric Causal Forest (CF) that incorporates smoothing splines to accommodate continuous treatments.
+Both methods preserve key strengths of causal machine learning by flexibly modeling complex interactions without imposing restrictive functional assumptions.
+To assess their performance, we compare them with several existing approaches, such as direct applications of random forests or neural networks, using a Monte Carlo simulation framework.
 
@@
 Although the methodological extension proposed here is primarily motivated by the needs of OFPE, it holds potential for a wide range of applications in applied economics where continuous treatments and heterogeneous effects are of interest.
-
-::: {.column-margin}
-<span style = "color: blue;"> SO-ANN not introduced yet. You need to talk more about what you do in the above paragraph (mostly SO-NN and semi-parametric CF). </span>
-:::
-<!--end of column-margin-->
 
@@
 #### DO‑CF (continuous)
-
-::: {.column-margin}
-<span style = "color: blue;"> Explain in  more detail about why it this is innovative. Explain the concept of linearization with splines using the figure I generated in my old manuscript. </span>
-:::
-<!--end of column-margin-->
 
@@
 $$
-\begin{align} 
+\begin{aligned}
 y_{i,j} =
@@
     \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
-  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
+    \eta_{i,j} + \varepsilon_{i,j}, & N \ge \tau_{i,j}
   \end{cases}
+\end{aligned}
 \label{eq:yield-gen-sim}
-\end{align}
 $$

### 2025-11-19 writing/manuscript_CF_continuous.qmd
Working on comments

@@
 
-To assess the predictive accuracy and economic relevance of these modeling approaches, we design a set of simulation experiments that reflect realistic agronomic conditions.
-These simulations serve as a controlled environment for evaluating model performance in estimating yield response functions and recommending optimal nitrogen rates.
-
-<span style = "color: red;"> Add profit differential as the measure. Discuss them in the results as well. </span>
+To assess the economic relevance of these modeling approaches, we use profit loss relative to the true site-specific EONR.
+The average per-acre profit from true site-specific EONR (denoted as $\pi_{true}$) is
+
+$$
+\begin{align}
+\pi_{true}  = \frac{1}{N}\cdot \sum_{i=1}^{N} [P_c \cdot f(N^*_i, X_i) - P_N \cdot N^*_i]
+\end{align}
+$$
+Where $P_c$ and $P_N$ represent the prices of corn yield and nitrogen, respectively.
+
+The average profit per acre associate with the site-specific N prescription by each model $m$ (denoted as $\pi$) is defined as
+
+$$
+\begin{align}
+\pi_m  = \frac{1}{N}\cdot \sum_{i=1}^{N} [P_c \cdot f(\widehat{N}^*_{m}(X_i), X_i) - P_N \cdot \widehat{N}^*_{m}(X_i)]
+\end{align}
+$$
+
+Then the profit loss of model $m$ relative to the true maximum profit achievable (denoted as $\pi-loss_{m}$) is then defined as
+
+$$
+\begin{align}
+\Delta\pi_m  = \pi_{true} - \pi_m
+\end{align}
+$$
 
@@
 ```
-@tbl-best_model_by_field_count highlights the top-performing model for each training size.
-CF is the dominant method when only a few fields are available, delivering the lowest mean RMSE for 1–5 fields.
-As field diversity increases, SO_ANN begins to outperform CF, becoming the preferred approach at 10 fields and showing a substantial further accuracy gain at 20 fields.
 
@@
 
+@fig-profit-loss shows the best-performing model at each training-data size based on average profit loss relative to the true economically optimal nitrogen rate.
+When only 1–5 fields are available, the semi-parametric-CF achieves the lowest profit loss.
+As the number of training fields increases, the SO_NN becomes the top performer, attaining substantially lower profit loss at 10 and 20 fields.
+This pattern shows that CF works best when data are limited, while SO-NN performs better once more fields are available.
+
+```{r}
+#| label: fig-profit-loss
+#| fig-cap: "Profit loss"
+knitr::include_graphics("Figures/best_models.png")
+```
+
 # Discussion

### 2025-11-25 presentation/proposal/proposal.qmd
Edit proposal

@@
 author:
-  - name: Mona Mousavi$^*$
+  - name: Mona Mousavi
 institute:
-  - $^*$ University of Nebraska Lincoln
+  - University of Nebraska Lincoln
 date: 06/12/2025
 format:
-  revealjs: default
+  revealjs:
+    theme: [default, ./../custom.scss]
+    fontsize: 1.4em
+    callout-icon: false
+    scrollable: true
+    fig-dpi: 400
   beamer:
@@
     fontsize: "10pt"
-header-includes: 
 logo: nebraska-n.jpg
@@
 
-# Background and Research Objectives
-
 ## Background
@@
   + Run on-farm agronomic trials and collect data
-  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
+  + <span style = "color: blue;">Estimate how yield responds to nitrogen based on observed characteristics</span>
   + Identify site-specific optimal nitrogen rates
@@
 
-+ In all of the previous studies, the best model was selected based on their abilities to predict \textcolor{blue}{yield level} at the observation points (yield-based model selection).
++ In all of the previous studies, the best model was selected based on their abilities to predict <span style = "color: blue;">yield level</span> at the observation points (yield-based model selection).
 
@@
 
-# Method
+## Method: Overview
 
@@
 
-Steps for a single iteration:
+### Steps for a single iteration:
 
@@
   + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
-  + Step 3: Select the best model among the 5 candidate models based on
-    + yield (conventional approach)
+  + Step 3: Select the best model among the 5 candidate models based on their abilities to predict
+    + yield level (conventional approach)
     + local-EONR (our approach)
@@
 
-The five candidate models:
-linear model, linear spatial error (SE) model, random forest (RF), boosted regression forest (BRF), and causal forest (CF).
+### The five candidate models:
 
-## Data generation
++ Linear model 
++ Linear spatial error (SE) model 
++ Random forest (RF) 
++ Boosted regression forest (BRF) 
++ Causal forest (CF)
+
+## Method: Data generation
 
@@
 
-+ One possible scenario of the spatial variability in the field characteristics is shown.
++ One possible scenario of the spatial variability in the field characteristics is shown below.
 
@@
 
-## The simulated OFPE field:
 ```{r}
@@
 
-## True yield responses
+## Method: True yield responses
 
@@
 
----
+## Method: On-farm trial (nitrogen rate)
 
-The true economically optimal N rate for each cell is then determined by:
-
-$$
-N_{j, i}^* \equiv \operatorname{argmax}_N\left[p \cdot f_{j, i}(N)-w \cdot N\right]
-$$
-
-+ $p$: the crop price, 
-+ $w$ the nitrogen fertilizer price, 
-+ $f_{j, i}(N)$: previously defined true yield response function for the specific cell.
-
----
-
-## On-farm trial (nitrogen rate)
-
-+ six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) followed a 6 × 6 Latin square design.
+Six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) following a 6 × 6 Latin square design.
 
 (Note: 8 blocks, each block including 36 plots ($8*36=288$ plots) to establish the Latin square design).
-
----
 
@@
 
----
-
-## Yield data
+## Method: Yield data
 
@@
 The harvesting equipment was assumed to be able to accurately collect yield data only at the subplot resolution. Therefore, the observed yield data were at the resolution of subplot level.
-
----
 
@@
 
-## independent variables
+## Method: Independent variables
 
-+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and economically optimal N rates $N_{j, i}^*$ are not directly observable by producers or researchers.
++ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ are not directly observable by producers or researchers.
 
@@
 
-$$\beta = \beta^1 + \beta^2, \;\;
-\beta^1 = \beta^{1,1} + \beta^{1,2}, \;\;\beta^2 = \beta^{2,1} + \beta^{2,2}$$ 
-$$\tau = \tau^1 + \tau^2, \;\;
-\tau^1 = \tau^{1,1} + \tau^{1,2}, \;\;\tau^2 = \tau^{2,1} + \tau^{2,2}$$ 
+$$\beta = \beta^{1,1} + \beta^{1,2} + \beta^{2,1} + \beta^{2,2}$$ 
+$$\tau = \tau^{1,1} + \tau^{1,2} + \tau^{2,1} + \tau^{2,2}$$ 
 $$
-\eta = \min(\eta^1, \eta^2), \;\;
-\eta^1 = \min(\eta^{1,1}, \eta^{1,2}), \;\;
-\eta^2 = \min(\eta^{2,1}, \eta^{2,2})
+\eta = \min(\min(\eta^{1,1}, \eta^{1,2}), \min(\eta^{2,1}, \eta^{2,2}))
 $$
 
-# Identifying the true best model among the five candidate models
+## Identifying the true best model among the five candidate models
 
@@
 + The estimated variable rate N recommendations are evaluated against the true variable rate N to evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):
-
----
 
@@
 + The entire dataset is randomly split into a train and test dataset in a spatially clustered manner, which is repeated to create multiple folds.
-
-## An example of the train-test split for each of the five folds
 
@@
 
-# Results and Conclusions
-
 ## Simulation Results: Performance of the model selection approaches
@@
 
-## Chapter two:
-
-### Causal methods for Site-specific yield response function estimation
-
-# Background and Research Objectives
+# Chapter 2: Causal methods for Site-specific yield response function estimation
 
@@
 
-+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.
++ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response <span style = "color: blue;">curve</span>.
 
@@
 
-## Method
+## Method: Double-ML (binary treatment ($T$) case):
 
-### Double-ML (binary treatment ($T$) case):
-
-DML works fine in a \textcolor{blue}{binary} treatment ($T$) case (model assumes linearity in $T$):
+DML works fine in a <span style = "color: blue;">binary</span> treatment ($T$) case (model assumes linearity in $T$):
 
@@
 \begin{aligned}
-& Y_{res}= \\
+& Y_{res} = \\
 & \begin{aligned}
-\theta(X) \cdot T + g(X, W) + v 
-- & \underbrace{E\{[\theta(x) \cdot T + g(X, W) + v] \mid X, W\}}= \\
-& \textcolor{blue}{=} \textcolor{blue}{\theta(X) \cdot E(T \mid X, W) + g(X, W)} \\
-& \textcolor{blue}{=}  \textcolor{blue}{\theta(X) \cdot h(X, W) + g(X, W)}
-\end{aligned} \\
-& \theta(X) \cdot T + g(X, W) + v 
-- \theta(X) \cdot h(X, W) - g(X, W)= \\
-& \theta(X) \cdot [\underbrace{T - h(X, W)}_{T_{res}}] + v \\
-& \Rightarrow Y_{res} = \theta(X) \cdot T_{res} + v
+& {[\theta(X)\cdot T + g(X,W) + v] - [E\{ \theta(X)\cdot T + g(X,W) + v \mid X,W\}]= } \\
+& {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot E(T\mid X,W) + g(X,W)]=} \\
+&  {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot h(X,W) + g(X,W)]=} \\
+&  \theta(X)\cdot [T - h(X,W)] + v =  \\
+& \theta(X)\cdot T_{res} + v
+\end{aligned}
+\\[0.5em]
+& \Rightarrow\quad Y_{res} = \theta(X)\cdot T_{res} + v
 \end{aligned}
@@
 
----
+## Method: Double-ML (continuous treatment ($T$) case):
 
-### Double-ML (continuous treatment ($T$) case):
-
-DML fails in a \textcolor{blue}{continuous} treatment ($T$) case:
+DML fails in a <span style = "color: blue;">continuous</span> treatment ($T$) case:
 
@@
 + $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed
-
----
 
@@
 $$
-RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
+RMSE_{EONR} = \sqrt{\frac{\sum_{i=1}^{n} (\widehat{EONR}_{i} - {EONR}_{i})^2}{n}}
 $$

### 2025-11-27 presentation/proposal/proposal.qmd
Edit

@@
 
-## Method: Data generation
+## Method: (Step 1) Data generation
 
@@
 + One possible scenario of the spatial variability in the field characteristics is shown below.
-
-+ This study assumed the yield data were collected at the subplot level (total 1440 subplots).
 
@@
 
-## Identifying the true best model among the five candidate models
+## (Step 2) Identifying the true best model among the five candidate models
 
-+ Each of the candidate models are trained using the whole experiment data and then variable rate EONR are estimated for the whole field.
++ Train Each of the candidate models using the whole experiment data and then estimate variable rate EONR for the whole field.
 
-+ The estimated variable rate N recommendations are evaluated against the true variable rate N to evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):
++ Evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):
 
@@
 
-## Model Selection
+## (Step 3) Model Selection
 
@@
 knitr::include_graphics("g_ranking.png")
-```
-
-## Simulation Results: Accuracy of the model selection approaches
-
-```{r}
-#| echo: false
-#| results: 'asis'
-tbl <- readRDS("main_selection_accuracy_table.rds")
-invisible(flextable::save_as_image(tbl, path = "table_accuracy.png"))
-knitr::include_graphics("table_accuracy.png")
-
 ```

### 2025-11-28 writing/manuscript_CF_continuous.qmd
Changing figure label

@@
 Efficient nitrogen (N) management remains one of the most critical challenges in modern agricultural production.
-Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm (<span style = "color: blue;"> add citations here </span>).
+Nitrogen is essential for achieving high crop yields, yet excessive or poorly timed applications can lead to economic losses and environmental harm (@fassa2022site).
 The principle of applying the right amount of nitrogen at the right place and time requires understanding the spatial and temporal variability in crop nitrogen needs.
@@
 ```{r fig-trial-design, fig.cap = "Example trial design"}
-g_td <- readRDS(here::here("writing/Figures/g_td.rds"))
+readRDS(here::here("writing/Figures/g_td.rds"))
 ```
@@
 SO_ANN initially lags behind CF (18.4 at one field), but its mean RMSE decreases steadily as more fields are added, surpassing CF performance at 10 and 20 fields (11.2 and 8.5, respectively).
-This pattern underscores the neural network’s superior ability to exploit richer spatial and agronomic heterogeneity when sufficient training data are available.
+This pattern hilightes the neural network’s ability to exploit richer spatial and agronomic heterogeneity when sufficient training data are available.
 
@@
 NO_RF maintains broad, right-shifted distributions and higher mean errors throughout (≈ 36–40), confirming its role as a weak baseline.
-Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, reflecting <span style = "color: red;"> over-correction </span>   in the orthogonalization process.
-
-::: {.column-margin}
-<span style = "color: blue;"> It is not over-correction. It is fundamentally messed up as the original functions are no longer retained. </span>
-:::
-<!--end of column-margin-->
+Finally, DO_ANN performs poorly throughout, with both histograms and means showing persistently high errors (≈ 90+) and negligible benefit from additional data, showing that the orthogonalization step breaks the original relationships in the data instead of preserving them.
 
@@
   map <- c(
-    "Simple_ANN" = "No_ANN",
-    "RANN"       = "DO_ANN",
-    "HALFRANN"   = "SO_ANN",
-    "HALFRRF"    = "SO_RF",
+    "NO_ANN" = "NO_NN",
+    "DO_ANN" = "DO_NN",
+    "SO_ANN" = "SO_NN",
     "NO_RF"  = "NO_RF",
+    "SO_RF"  = "SO_RF",
     "CF"     = "CF"
   )
+
   models_upper <- toupper(models)
   out <- map[models_upper]
+
   ifelse(is.na(out), models, out)
@@
 # ggsave(
-#   filename = here::here("Proposal_presentation/fig_rmse_hist.png"),
+#   filename = here::here("presentation/proposal/fig_rmse_hist.png"),
 #   plot = p,
@@
 
+model_names <- function(models) {
+  map <- c(
+    "CF"     = "CF",
+    "DO_ANN" = "DO_NN",
+    "NO_ANN" = "NO_NN",
+    "SO_ANN" = "SO_NN",
+    "NO_RF"  = "NO_RF",
+    "SO_RF"  = "SO_RF"
+  )
+
+  models_upper <- toupper(models)
+  out <- map[models_upper]
+  ifelse(is.na(out), models, out)
+}
+
 rmse_df <- rmse_tbl_raw %>%
@@
   dplyr::mutate(
-    # Field labels (auto from unique values)
+    Model_lab = model_names(Model),
     field_f = factor(
@@
       levels = sort(unique(`#field`)),
-      labels = paste0(sort(unique(`#field`)),
-                      ifelse(sort(unique(`#field`)) == 1, " field", " fields"))
+      labels = paste0(
+        sort(unique(`#field`)),
+        ifelse(sort(unique(`#field`)) == 1, " field", " fields")
+      )
     )
@@
 rmse_summary <- rmse_df %>%
-  dplyr::group_by(Model, field_f) %>%
+  dplyr::group_by(Model_lab, field_f) %>%
   dplyr::summarise(
@@
   )
-# 
-p_mean <- ggplot(rmse_summary, aes(x = field_f, y = mean, group = Model, color = Model)) +
+
+# 
+p_mean <- ggplot(rmse_summary, aes(x = field_f, y = mean,
+                                   group = Model_lab, color = Model_lab)) +
   geom_line(linewidth = 0.7, alpha = 0.8) +
   geom_point(size = 3) +
-  labs(x = "# fields", y = "Mean RMSE", color = "Model") +
+  labs(
+    x = "# fields",
+    y = "Mean RMSE",
+    color = "Model"
+  ) +
   theme_bw() +
@@
     axis.title = element_text(size = 13),
-    strip.text = element_text(face = "bold")
+    strip.text = element_text(face = "bold"),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12)
   )
@@
 
-# ggsave(here::here("Proposal_presentation","model_improvement.png"), p_mean, width = 9, height = 5, dpi = 300)
+# ggsave(here::here("presentation", "proposal", "model_improvement.png"), p_mean, width = 9, height = 5, dpi = 300)
 ```
@@
 
-@fig-improvement reports each method’s percentage improvement in mean RMSE relative to the NO_RF baseline.
-CF consistently provides the strongest gains among the robust approaches when data are scarce—reducing error by ≈63% with a single field, and climbing to ≈67–68% once 5–20 fields are available.
-SO_ANN exhibits a similar improvement trend but grows even more dominant as more diverse data accumulate, achieving ≈53% improvement at 1 field, ≈63% by 5 fields, and reaching ≈68% at 10 fields and ≈77% at 20 fields, the highest improvement across all models.
-NO_ANN transitions from a large negative result when trained on just one field (substantially worse than baseline) to strong relative improvements exceeding 65% by 10–20 fields, reinforcing its reliance on adequate sample support.
-SO_RF offers moderate improvement (≈38–43%) that changes little with added fields.
-DO_ANN remains entirely uncompetitive across all sample sizes—showing worsening performance relative to NO_RF by more than 100–150%, indicating severe model misspecification.
-
-::: {.column-margin}
-<span style = "color: blue;"> All this is probably redundant. What are you adding beyond the previous paragraphs? </span>
-:::
-<!--end of column-margin-->
-
-Together, these results emphasize that while CF is the most reliable option under limited data availability, SO_ANN ultimately yields the largest relative error reduction when multi-field datasets are accessible, highlighting its superior scalability with richer training information.
+@fig-improvement compares the performance of CF and SO_ANN relative to the worst-performing model, DO_ANN, across different numbers of fields.
+Both methods deliver large improvements in RMSE at all sample sizes.
+With only one field, CF and SO_ANN reduce error by roughly 80–84%.
+As more fields become available, both approaches strengthen further: CF’s improvement rises to about 87% by 5–10 fields, while SO_ANN continues to gain with increasing data, reaching nearly 91% at 20 fields.
 
@@
 # compare to which model?
-baseline <- "NO_RF"
+baseline <- "DO_ANN"
 
@@
 
+models_to_keep <- c("SO_ANN", "CF")
+
+avg_by_filtered <- avg_by %>%
+  dplyr::filter(Model %in% models_to_keep | Model == baseline)
+
 # baseline per field 
-baseline_by_field <- avg_by %>%
+baseline_by_field <- avg_by_filtered  %>%
   dplyr::filter(Model == baseline) %>%
@@
 # join and compute % improvement
-rel <- avg_by %>%
+rel <- avg_by_filtered  %>%
   dplyr::left_join(baseline_by_field, by = "field_f") %>%
@@
 
-<span style = "color: blue;"> I do not see the purpose of this much. Contrast CF, SO-NN, and the worst (all in the same panel) across the number of fields. </span>
+@fig-profit-loss shows the best-performing model at each training-data size based on average profit loss relative to the true economically optimal nitrogen rate.
+When only 1–5 fields are available, the semi-parametric-CF achieves the lowest profit loss.
+As the number of training fields increases, the SO_NN becomes the top performer, attaining substantially lower profit loss at 10 and 20 fields.
+This pattern shows that CF works best when data are limited, while SO-NN performs better once more fields are available.
+
+```{r}
+#| label: fig-profit-loss
+#| fig-cap: "Profit loss"
+knitr::include_graphics("Figures/best_models.png")
+```
 
@@
 
-@fig-profit-loss shows the best-performing model at each training-data size based on average profit loss relative to the true economically optimal nitrogen rate.
-When only 1–5 fields are available, the semi-parametric-CF achieves the lowest profit loss.
-As the number of training fields increases, the SO_NN becomes the top performer, attaining substantially lower profit loss at 10 and 20 fields.
-This pattern shows that CF works best when data are limited, while SO-NN performs better once more fields are available.
-
-```{r}
-#| label: fig-profit-loss
-#| fig-cap: "Profit loss"
-knitr::include_graphics("Figures/best_models.png")
-```
-
 # Discussion
@@
 SO_ANN improves monotonically with additional fields and ultimately dominates once ≥10 fields are available, achieving the lowest mean RMSE overall (e.g., 11.2 at 10 fields and 8.5 at 20 fields).
-These complementary strengths that CF is preferred under data-constrained conditions, whereas SO_ANN becomes the method of choice as multi-field datasets accumulate (see @tbl-best_model_by_field_count).
+These complementary strengths that CF is preferred under data-constrained conditions, whereas SO_ANN becomes the method of choice as multi-field datasets accumulate.
 
